### 2016-03-30 manuscript/political_bias_prediction.tex
Added draft for ECML.

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\setcounter{tocdepth}{3}
\usepackage{graphicx}
% For citations
\usepackage{natbib}
\usepackage{amsmath,amsfonts,amscd,amssymb}
\usepackage{dsfont}
\renewcommand{\vec}[1]{\mathbf{#1}}
\usepackage{hyperref}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Corr}{Corr}
\newcommand{\R}{\mathds{R}}

\usepackage{url}
\urldef{\mailsa}\path|{felix.biessmann,|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Automating Political Bias Prediction}

% a short form should be given in case it is too long for the running head
\titlerunning{Political Bias Prediction}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{
%Felix Bie\ss{}mann%
\thanks{}
%
\authorrunning{Political Bias Prediction}}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
%\institute{
%\mailsa\\
%\mailsb\\
%\mailsc\\
%\url{http://www.springer.com/lncs}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Political Bias Prediction}
\tocauthor{}
\maketitle

\begin{abstract} 
Every day media generate large amounts of text. Getting an unbiased view on what media report on requires an unbiased and heterogenous sample of media content. Assistive technology for estimating the political bias of texts can be helpful in this context. In this study a simple statistical learning approach is presented that can be used to predict political bias from text. Standard text features extracted from political texts are used to train classifiers that predict political bias in terms of a) party affiliation and b) in terms of political views. Results indicate that political bias prediction is possible with well above chance accuracy. The results are further validated by relating discriminative text features to the political orientation of a party. Moreover the results demonstrate that sentiment appears to be a very discriminative feature: Political power of an author strongly correlates with positive sentiment of a text. To highlight some potential use cases a web application shows how the model can be used for texts for which the political bias is not clear such as news articles.
\end{abstract}

\section{Introduction}
\label{sec:intro}
%
Modern media generate a large amount of content at an ever increasing rate. Keeping an unbiased view on what media report on requires to understand the political bias of texts. In many cases it is obvious which political bias an author has. In other cases some expertise is required to judge the political bias of a text. 
%
When dealing with large amounts of text however there are simply not enough experts to examine all possible sources and publications. Assistive technology can help in this context to try and obtain a more unbiased sample of information.

Ideally one would choose for each topic a sample of reports from the entire political spectrum in order to form an unbiased opinion. But ordering and assigning media content with respect to the political spectrum at scale requires automated prediction of political bias. The aim of this study is to provide some empirical evidence indicating that political bias prediction is possible with well above chance accuracy.

%-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate

\section{Data Sets and Feature Extraction}\label{sec:data}
%
We used publicly available data sets of german political texts and standard libraries for processing the text. The following sections describe the details of data acquisition and feature extraction.

\subsection{Data}
We obtained annotated political text data from sources a) the discussions and speeches held in the german parliament ({\em Bundestag}) and b) all manifesto texts from parties running for election in the german parliament in the current 18th and the last, 17th, legislation period.

\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy of political bias. The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}, we used an open source API to that data to query the data in cleaned and structured format\footnote{\url{https://github.com/bundestag}}. This API returns for each speech the speaker and its party affiliation. In total we obtained 22784 speeches for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.

\paragraph{Party manifesto data}
For party manifestos we also use an openly accessible API provided by the Wissenschaftszentrum Berlin (WZB). The API is released as part of the {\em Manifestoproject} \cite{manifesto}. The data released in this project comprises the complete manifestos for each party that ran for election enriched with annotations by political experts. Each sentence (in some cases also parts of sentences) are labels with one of 56 political labels. Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare, ...}. The set of labels was developed by political scientists at the WZB and released for public use. We obtained all manifestos of parties that were running for election in this and the last legislative period. In total this resulted in 29451 political statements that had two types of labels, for one we stored the party affiliation of each political statement. This party affiliation label was used to evaluate the party evaluation classifiers trained on the parliament speeches. For this purpose we constrained the data to only those parties that were elected into the parliament. Next to the party affiliation we were mainly interested in the political view labels annotated by human experts. For the analyses based on political view labels we considered all parties, also those that did not make it into the parliament.

\subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
For each semantic unit, in the case of parliament discussions this were the speeches\footnote{Often speeches were interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.} in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 political view labels. Strings of each semantic unit were tokenized and transformed into bag-of-word vectors as implemented in scikit-learn \cite{scikit-learn}. The general idea of bag-of-words vectors is to simply count occurences of words (or word sequences, also called {\em n-grams}) for each data point (usually documents, here parliament speeches). So the text of each speech is transformed into a vector $\vec{x}\in\mathds{R}^d$, were $d$ is the size of our dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary. We tried several options for vectorizing the speeches, including term-frequency-inverse-document-frequency normalisation using n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words. All of these hyperparameters were subjected to hyperparameter optimization as explained in \autoref{sec:crossvalidation}.

\section{Classification Model and Training Procedure}

We used a multinomial logistic regression model in order to classify bag-of-words feature vectors $\vec{x}\in\R^d$, where $d$ is the size of the bag-of-words dictionary. Let $y\in\{1,2,\dots,K\}$ be the true party label, where $K$ is the total number of parties in the parliament and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ be the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then 
\begin{eqnarray}\label{eq:logreg_multiclass}
p(y=k|\vec{x},\vec{W}) = &\frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}} \qquad \textrm{with }  z_k=&\vec{w}_k^{\top}\vec{x} \\\nonumber
\end{eqnarray}
%
We estimated $\vec{W}$ using quasi-newton gradient descent as packaged in the scikit-learn toolbox. The optimization function was obtained by adding a penalization term to the negative log-likelihood of the multinomial logistic regression objective and the optimization hence found the $\vec{W}$ that minimized
\begin{equation}\label{eq:objective}
L(\vec{W}, \vec{x}, \gamma) = - \log{\frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}}}+ \gamma \| \vec{W} \|_{F}
\end{equation}
Where $\|~\|_F$ denotes the Frobenius Norm and $\gamma$ is a regularization parameter controlling the complexity of the model. 
 We optimized the regularization parameter on a log-scaled grid from $10^{-4,\dots,4}$ and the regularization constant was adopted to reflect asymmetric class frequency distributions. The performance of the model was optimized using the classification accuracy, but we also report all other standard measures, precision ($TP / (FP + TP$), recall ($TP / (TP + FN)$) and f1-score ($2\times (Prec. \times Rec) / (Prec + Rec.)$).

Next to the four class problem of categorizing all four parties of the german parliament we also considered the binary classification task of only distinguishing texts between wether they were read by members of the government or members of the opposition. Results of this simpler two-class problem were obtained by exactly the same methods as in the four class case.

\subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
The model pipeline contained a number of {\em hyperparameters} that we did not want to tune by hand. These parameters are called hyperparameters becase we cannot optimize these like we optimize the parameters of our classifier, by writing down an objective function and optimizing this linear function by doing gradient descent. Thus we resort to the standard 'trial-and-error' approach in machine learning and perform cross-validation on the hyperparameters. In order to not overfit to our data and obtain good generalization performance to other texts we used two approaches. The first one is the usual procedure of 'nested-crossvalidation', which performs two stages of cross-validation, an {\em outer} cross-validation loop to evaluate the model correctly on held-out data, and an {\em inner} cross-validation loop within each training data fold of the outer cross-validation. In practice this means that in each step of the outer loop we set aside a part of our data set and then try out all parameters using the inner cross-validation on the not-held-out data, and afterwards make a prediction on the held-out data of the outer cross-validation. For each set of preprocessing (hyper-) parameters and regularisation parameter of the classifier, we performed the entire data extraction, vectorization and classifier training workflow. Next to this rather standard approach, we also tested the classifiers on the text of the party programs (see \autoref{sec:data}) which were never included in any of the training procedures.

\subsection{Sentiment analysis}\label{sec:sentiment_analysis_methods}
We used a publicly available key word list to extract sentiments from the text subjected to classification \cite{remquahey2010}. The sentiment index used for attributing positive or negative sentiment to a text was computed  as the Pearson correlation coefficient between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which was constructed by simply taking the values from the sentiment dictionary.

\subsection{Analysing bag-of-words features}
In order to get  better picture of which features contribute to the classification we computed correlation coefficients between each word and each party. Often linear models are interpreted by just looking at the weights of the model. Note that due to the fact that the features are obtained from real data and thus correlated we cannot inspect the model by interpreting the coefficients of $\vec{W}$; for an in-depth discussion of this problem see e.g. \cite{Haufe2013}. The words corresponding to the top positive and negative correlations are shown in section \autoref{sec:word_party_correlations}.

\small
\subsection*{Acknowledgements}

%
\small{
\bibliographystyle{plain}
\bibliography{political_bias_prediction} 
}

\end{document}

### 2016-03-30 bundestag.py

@@ -220,7 +220,8 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     predictedTest = test_clf.predict(X_test)
     # dump report on training held out data with CV
     report = "*** Training Set (CVd) ***\n" + metrics.classification_report(y_test, predictedTest)
-    report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(test_clf.steps[-1][1].classes_)+'\n'
+    labelsStr = [str(x) for x in test_clf.steps[-1][1].classes_]
+    report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(y_test, predictedTest).tolist(): report += str(line)+"\n" 
     # train again on entire training set
     final_clf = text_clf.set_params(**best_clf.get_params()).fit(trainData, trainLabels)
@@ -229,7 +230,8 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     # test on evaluation data ste
     predictedEval = final_clf.predict(evalData)
     report += "*** Evaluation Set ***\n" + metrics.classification_report(evalLabels, predictedEval)
-    report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(final_clf.steps[-1][1].classes_)+'\n'
+    labelsStr = [str(x) for x in final_clf.steps[-1][1].classes_]
+    report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(evalLabels, predictedEval).tolist(): report += str(line)+"\n" 
     # dump report
     open(OUT_DIR+'/report-'+saveId+'.txt','wb').write(report)

### 2016-03-30 manuscript/political_bias_prediction.tex
More text.

@@
 
+\section{Results}
+
+The following sections provide an overview of the results for both types of political bias prediction tasks, party affiliation and political view.
+Some interpretations of discriminative features are given one section will highlight a finding on the correlation of political power and positive sentiment of political texts.
+
+\subsection{Predicting political party affiliation}
+
+% Confusion matrix
+%\begin{table}[t]\label{tab:conf_mat_four_class}
+%\begin{center}
+%\begin{tabular}{lrrrr}
+%&	1402&518&216&179\\
+%&510&1939&295&306\\
+%&185&235&1422&766\\
+%&202&271&672&2692\\
+%\end{tabular}
+%\end{center}
+%\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lrrrrrrrr}
+& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\hline \hline
+      Linke     &  0.61    &  0.61  &    0.61   &   2315 &  0.81   &   0.76   &   0.79     &  104\\
+     Gr\"une   &    0.65    &  0.64   &   0.64 &     3050 & 0.57  &     0.21  &    0.31    &   110\\
+        SPD     &  0.55    &  0.55  &   0.55 &    2608 & 0.10  &    0.50   &   0.16    &    16\\
+        CDU/CSU    &   0.68    &  0.70  &    0.69  &    3837 & 0.06   &   0.08  &    0.07   &     38 \\
+\hline
+avg / total    &   0.63  &    0.63  &    0.63 &    11810 & 0.57   &   0.42    &  0.45  &     268\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_four_classes}
+Classification performance on the four party prediction problem; while classification on held-out data from parliament speeches was on average higher than on party programs, also on the completely out-of-training-sample data the model obtained reasonable accuracies in some cases.
+}
+\end{table}
+
+\subsection{Predicting political views}
+
+\section{Speech sentiment correlates with political power}\label{sec:sentiment_result}
+We analysed the sentiment of each speech of a political party and plot the results in \autoref{fig:party_sentiments}.
+While the methodology used for estimating the sentiment (see \autoref{sec:sentiment_analysis_methods}), it can yield some insight into the emotional content of a speech.
+High values indicate more pronounced usage of positive words, whereas negative values indicate more pronounced usage of words associated with negative emotional content.
+We can quantify political power as the number of seats a party has in the parliament.
+Our analyses show a strong trend supporting the hypothesis that the more political power a party has, the more positive sentiment their speech has, and vice versa, the less seats a party has, the more negative content their speeches contain.
+This pattern is evident from the barplot in \autoref{fig:party_sentiments} and supported by a high correlation of 0.86 between number of parliament seats and sentiment index.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=8cm]{images/party-sentiments-18.pdf} 
+%
+\end{center}
+\caption{
+\label{fig:party_sentiments}
+Speech sentiments computed for speeches of each party (parties are ordered according to political power measured in number of seats in the parliament from left to right).
+The more political power a party has, the more positive the speech content is, the fewer seats a party has the more words associated with negative emotions are used in a speech.
+}
+\end{figure}
+
+\section{Correlations between words and parties}\label{sec:word_party_correlations}
+
+\autoref{fig:party_word_correlations} shows the correlations of individual words and each party.
+We plotted the top 20 words for each sign, i.e.
+the 20 words with the strongest positive and negative correlations for each party, respectively.
+While the correlations do indicate that there are -- despite extensive and careful cleaning procedures -- still words with strong correlations in the top words, we conjecture that at least some of these words do have some discriminatory power, even if they do not refer to political views.
+Below we list some examples, but the most prominent is probably: the opposition often addresses the government in their speeches, the govermental parties do that less often.
+This is reflected in the frequency of the usage of words that refer to members of the government or ministers.
+But among the often used and avoided words we also find words that refer to actual political content.
+Thus we hypothesize that to some extent these correlation coefficients can give hints as to what the bag-of-words histogram representation into which we transformed the speeches are related to the actual content of the speeches.
+In the following we give some examples of words that appear to be preferentially used or avoided by each respective party.
+Even though interpretations of these quantitative results are somewhat invalid, as they neglect the context in which these words were mentioned, we find some interesting patterns.
+
+\begin{figure}
+\begin{center}
+%\includegraphics[width=3.5cm]{party_word_correlations-linke.pdf} 
+%\includegraphics[width=3.1cm]{party_word_correlations-gruene.pdf} 
+%\includegraphics[width=3.45cm]{party_word_correlations-spd.pdf} 
+%\includegraphics[width=3.65cm]{party_word_correlations-cdu.pdf}
+%
+\end{center}
+\caption{
+\label{fig:party_word_correlations}
+Correlations between words and party for parliament speeches. }
+\end{figure}
+
 \small

### 2016-03-30 manuscript/political_bias_prediction.tex
More results in tables

@@
 \hline \hline
-      Linke     &  0.61    &  0.61  &    0.61   &   2315 &  0.81   &   0.76   &   0.79     &  104\\
-     Gr\"une   &    0.65    &  0.64   &   0.64 &     3050 & 0.57  &     0.21  &    0.31    &   110\\
-        SPD     &  0.55    &  0.55  &   0.55 &    2608 & 0.10  &    0.50   &   0.16    &    16\\
-        CDU/CSU    &   0.68    &  0.70  &    0.69  &    3837 & 0.06   &   0.08  &    0.07   &     38 \\
+       cducsu   &    0.62  &    0.81  &    0.70  &     706&0.26    &  0.58   &   0.36    &  2030\\
+        fdp    &   0.70   &   0.37  &    0.49    &   331&0.38   &   0.28    &  0.33   &   2319\\
+     gruene &      0.59  &    0.40   &   0.48   &    298&0.47  &    0.20   &   0.28 &     3747\\
+      linke    &   0.71   &   0.61  &    0.65    &   338&0.30   &   0.47   &   0.37    &  1701\\
+        spd   &    0.60   &   0.69  &    0.65   &    606&0.26   &   0.16   &   0.20  &    2278\\
 \hline
-avg / total    &   0.63  &    0.63  &    0.63 &    11810 & 0.57   &   0.42    &  0.45  &     268\\
+avg / total &      0.64   &   0.63   &   0.62    &  2279 &0.35    &  0.31 &     0.30   &  12075
 %
@@
 \label{tab:results_four_classes}
-Classification performance on the four party prediction problem; while classification on held-out data from parliament speeches was on average higher than on party programs, also on the completely out-of-training-sample data the model obtained reasonable accuracies in some cases.
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively. Chance level is at 0.2.  
+}
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lrrrrrrrr}
+& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\hline \hline
+government    &   0.83   &   0.84&      0.84     & 1037 & 0.49  &    0.59 &     0.54  &    4349\\
+ opposition     &  0.86  &    0.86   &   0.86    &  1242 & 0.74 &     0.66  &    0.70   &   7726\\
+\hline
+avg / total   &    0.85 &     0.85 &     0.85  &    2279 & 0.65  &     0.63  &     0.64  &    12075\\
+
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_binary}
+Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches to government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
 }
@@
 \subsection{Predicting political views}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lrrrr}
+    &         precision    &recall &  f1-score  & N\\
+\hline\hline
+test set    &  0.48   &   0.47  &    0.47  &    2651\\
+evaluation set    &  0.47    &  0.46 &     0.46 &     2946\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_avg_political_view}
+Classification performance on the 56 class prediction problem of political views, see \autoref{sec:data}; chance performance would be around 0.02.
+}
+\end{table}
 

### 2016-03-31 manuscript/political_bias_prediction.tex
yet more text

@@
 The aim of this study is to provide some empirical evidence indicating that political bias prediction is possible with well above chance accuracy.
+\\
 
-%-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate
+Automated predictions of political bias can be problematic for ethical reasons.
+One trusts political experts, as they will take responsibility for what they say about a text, and they can explain their decisions.
+This is a key difference to many statistical learning approaches.
+Not only is the responsibility question not entirely clear, it can also be difficult to interpret some of the decisions.
+In order to validate the predictions of the models we devise two strategies that allow for better interpretations of the models.
+First we are looking at univariate measures of how discriminative the features are that the model uses.
+This somewhat allows to relate the political texts to the kind of information that political experts use, too.
+Second we use sentiment analysis to investigate whether this aspect of language also has discriminatory power.
+\\
+
+In the following \autoref{sec:data} gives an overview of data acquisition and preprocessing, \autoref{sec:model} presents the model, in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
 
@@
 
-\section{Classification Model and Training Procedure}
+\section{Classification Model and Training Procedure}\label{sec:model}
 
@@
 
-Next to the four class problem of categorizing all four parties of the german parliament we also considered the binary classification task of only distinguishing texts between wether they were read by members of the government or members of the opposition.
-Results of this simpler two-class problem were obtained by exactly the same methods as in the four class case.
+We considered three different classification problems: 1) classification of party affiliation (5 class problem for the 17th legislation period, 4 class problem for the 18th legislation period), 2) classification of government membership (binary problem) and 3) classification of political views as expressed by the political codes of the manifestoproject (56 class problem).
+For each of first two problems, party affiliation and government membership prediction, we trained classifiers on the parliament speeches.
+For the third problem we trained only on the manifesto data as only for those the political view labels were available.
 
@@
 The model pipeline contained a number of {\em hyperparameters} that we did not want to tune by hand. 
-These parameters are called hyperparameters becase we cannot optimize these like we optimize the parameters of our classifier, by writing down an objective function and optimizing this linear function by doing gradient descent.
-Thus we resort to the standard 'trial-and-error' approach in machine learning and perform cross-validation on the hyperparameters.
-In order to not overfit to our data and obtain good generalization performance to other texts we used two approaches.
-The first one is the usual procedure of 'nested-crossvalidation', which performs two stages of cross-validation, an {\em outer} cross-validation loop to evaluate the model correctly on held-out data, and an {\em inner} cross-validation loop within each training data fold of the outer cross-validation.
-In practice this means that in each step of the outer loop we set aside a part of our data set and then try out all parameters using the inner cross-validation on the not-held-out data, and afterwards make a prediction on the held-out data of the outer cross-validation.
-For each set of preprocessing (hyper-) parameters and regularisation parameter of the classifier, we performed the entire data extraction, vectorization and classifier training workflow.
-Next to this rather standard approach, we also tested the classifiers on the text of the party programs (see \autoref{sec:data}) which were never included in any of the training procedures.
+We first split the training data into a training data set and a test data set for optimizing the hyperparameters, the test data for hyperparameter optimization was 10\% of the entire training data.
+Hyperparameters were optimized using grid search: For each setting of hyperparameters the entire pipeline was trained and evaluated on the test set.
+For the best setting of all hyperparameters, we trained the pipeline again on all training data and evaluated it on the evaluation set.
+Evaluation data was never used for training the models.
+For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a legislative period.
+Evaluation data were the texts from party manifestos.
+For the political view prediction setting we first split all labeled manifesto sentences in both legislative periods into a training/test and evaluation set of 90\% (train/test) and 10\% (evaluation), then we split the train/test part again with a 90\%/10\% ratio for training/testing of the hyperparameters.
 
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 
@@
 
-\section{Speech sentiment correlates with political power}\label{sec:sentiment_result}
+\subsection{Speech sentiment correlates with political power}\label{sec:sentiment_result}
 We analysed the sentiment of each speech of a political party and plot the results in \autoref{fig:party_sentiments}.
@@
 
-\section{Correlations between words and parties}\label{sec:word_party_correlations}
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lrr}
+   Sentiment vs. &          Gov. Member    &  Seats\\
+\hline\hline
+17th period    &  0.84 & 0.70\\
+18th period    &  0.98 & 0.89\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_avg_political_view}
+Correlation coefficient between average sentiment of political speeches of a party in the german bundestag with two measures of their political power, a) their membership in the government and b) with the number of seats that party occupies in the parliament.
+Shown are correlations for the 17th legislative period and the (ongoing) 18th legislative period.
+}
+\end{table}
+
+\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
 
@@
 
+\section{Conclusions and Outlook}\label{sec:conclusion}
+
+%-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate
+
 \small

### 2016-03-31 bundestag.py

@@ -95,7 +95,8 @@ def get_sentiments(legis=17):
     sortedParties = [p +" (%d)"%bundestagSeats[legis][p]+ "\n" + '[Government]' if p in gov else p+" (%d)"%bundestagSeats[legis][p]+"\n[Opposition]"for p in sortedParties]
     import pylab
     pylab.figure(figsize=(7,4))
-    pylab.boxplot(sortedSentiments,labels=sortedParties)
+    #pylab.boxplot(sortedSentiments,labels=sortedParties)
+    pylab.bar(sortedSentiments,labels=sortedParties)
     pylab.grid('on')
     font = {'family' : 'normal', 'size'   : 10}
     pylab.rc('font', **font)
@@ -158,6 +159,7 @@ def list_top_words(legis=17,topwhat=20):
         pylab.barh(arange(len(words)),wordcors,color=colors[party])
         pylab.yticks(arange(len(words)),words)
         pylab.ylim(-.2,len(words))
+        pylab.xticks([-.2,0,.2])
         pylab.title(party)
         pylab.xlabel('Correlation')
         font = {'family' : 'normal', 'size'   : 16}

### 2016-03-31 manuscript/political_bias_prediction.tex
More tables and some text.

@@
 \newcommand{\R}{\mathds{R}}
-
+\usepackage{multicol}
+\usepackage{multirow}
+\usepackage{pbox}
 \usepackage{url}
@@
 % Confusion matrix
-%\begin{table}[t]\label{tab:conf_mat_four_class}
-%\begin{center}
-%\begin{tabular}{lrrrr}
-%&	1402&518&216&179\\
-%&510&1939&295&306\\
-%&185&235&1422&766\\
-%&202&271&672&2692\\
-%\end{tabular}
-%\end{center}
-%\end{table}
+\begin{table}[t]\label{tab:conf_mat_four_class}
+\begin{tabular}{lccccccc}
+&&& \multicolumn{5}{c}{Predicted Class}\\
+&&& cducsu & fdp& gruene& linke& spd\\
+\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True Class}}}& &cducsu &1186 &289& 178& 198& 179\\
+&&fdp &882& 658& 236& 329& 214\\
+&&gruene &1174& 404& 764& 941& 464\\
+&&linke &388& 92& 214& 806& 201\\
+&&spd &999& 268& 240& 398& 373\\
+ \multicolumn{8}{c}{\bf 17th Bundestag}\\
+\end{tabular}
+\quad
+\begin{tabular}{lcccccc}
+&&& \multicolumn{4}{c}{Predicted Class}\\
+&&& cducsu & gruene& linke& spd\\
+\multirow{4}{*}{\rotatebox{90}{\pbox{3cm}{\centering True Class}}}&&cducsu&1912& 156& 331& 584\\
+&&gruene&2092& 827& 1311& 1444\\
+&&linke&596& 186& 1216& 557\\
+&&spd&1284& 226& 563& 916\\
+ \multicolumn{7}{c}{\bf 18th Bundestag}\\
+\end{tabular}
+\caption{\label{tab:confusion} Confusion matrices for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+In particular the green party suffers from this effect: when trained on the speeches in the 18th Bundestag, 2092 statements out of the green party's manifesto were classified as belonging to the cducsu party.
+This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
+\end{table}
 
@@
 \begin{center}
-\begin{tabular}{lrrrrrrrr}
+\begin{tabular}{lcccccccc}
 & \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
@@
 \caption{
-\label{tab:results_four_classes}
+\label{tab:results_17}
 Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively. Chance level is at 0.2.  
@@
 \begin{center}
-\begin{tabular}{lrrrrrrrr}
+\begin{tabular}{lcccccccc}
+& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\hline \hline
+    cducsu    &   0.66   &   0.82   &   0.73    &   456 & 0.32  &    0.64  &    0.43    &  2983\\
+     gruene   &    0.68    &  0.54   &   0.60    &   173   &0.59   &   0.15   &   0.23   &   5674\\
+      linke     &  0.77  &    0.58    &  0.66    &   173 & 0.36   &   0.48   &   0.41   &   2555\\
+        spd     &  0.60  &    0.54   &   0.57    &   330 & 0.26 &     0.31   &   0.28     & 2989\\
+\hline
+avg / total    &   0.66  &    0.66  &    0.65   &   1132&  0.42 &     0.34  &    0.32&     14201\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_18}
+Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively. 
+}
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lcccccccc}
 & \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
@@
 \caption{
-\label{tab:results_binary}
-Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches to government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
+\label{tab:results_binary_17}
+Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches into government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
+}
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lcccccccc}
+& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\hline \hline
+government   &    0.88  &    0.95    &  0.92   &    786   &0.52   &   0.66 &     0.58  &    5972\\
+ opposition    &   0.86    &  0.71   &   0.78   &    346 & 0.69  &    0.56    &  0.62   &   8229\\
+\hline
+avg / total    &   0.88    &  0.88     & 0.87  &    1132 &  0.62   &   0.60    &  0.60 &    14201\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_binary_18}
+Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
 }

### 2016-03-31 bundestag.py

@@ -96,7 +96,7 @@ def get_sentiments(legis=17):
     import pylab
     pylab.figure(figsize=(7,4))
     #pylab.boxplot(sortedSentiments,labels=sortedParties)
-    pylab.bar(sortedSentiments,labels=sortedParties)
+    pylab.bar(arange(len(sortedSentiments)),[mean(s) for s in sortedSentiments],labels=sortedParties)
     pylab.grid('on')
     font = {'family' : 'normal', 'size'   : 10}
     pylab.rc('font', **font)

### 2016-03-31 manuscript/political_bias_prediction.tex
More images and tables

@@
 Every day media generate large amounts of text.
-Getting an unbiased view on what media report on requires an unbiased and heterogenous sample of media content.
+An unbiased view on what media report on requires an understanding of the political bias of media content.
 Assistive technology for estimating the political bias of texts can be helpful in this context.
-In this study a simple statistical learning approach is presented that can be used to predict political bias from text.
-Standard text features extracted from political texts are used to train classifiers that predict political bias in terms of a) party affiliation and b) in terms of political views.
-Results indicate that political bias prediction is possible with well above chance accuracy.
-The results are further validated by relating discriminative text features to the political orientation of a party.
-Moreover the results demonstrate that sentiment appears to be a very discriminative feature: Political power of an author strongly correlates with positive sentiment of a text.
+This study proposes a simple statistical learning approach to predict political bias from text.
+Standard text features extracted from speeches and manifestos of political parties are used to predict political bias in terms of political party affiliation and in terms of political views.
+Results indicate that political bias can be predicted with well above chance accuracy.
+Mistakes of the model can be interpreted with respect to changes of policies of political actors.
+Two approaches are presented to make the results more interpretable: a) discriminative text features are related to the political orientation of a party and b) sentiment features of texts are correlated with a measure of political power.
+Political power appears to be strongly correlated with positive sentiment of a text.
 To highlight some potential use cases a web application shows how the model can be used for texts for which the political bias is not clear such as news articles.
@@
 We used a publicly available key word list to extract sentiments from the text subjected to classification \cite{remquahey2010}.
-The sentiment index used for attributing positive or negative sentiment to a text was computed  as the Pearson correlation coefficient between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which was constructed by simply taking the values from the sentiment dictionary.
+The sentiment index used for attributing positive or negative sentiment to a text was computed  as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which was constructed by simply taking the values from the sentiment dictionary.
 
@@
 Often linear models are interpreted by just looking at the weights of the model.
-Note that due to the fact that the features are obtained from real data and thus correlated we cannot inspect the model by interpreting the coefficients of $\vec{W}$; for an in-depth discussion of this problem see e.g.
+Since the text features are obtained from real data and thus correlated we cannot inspect the model by interpreting the coefficients of $\vec{W}$, see e.g.
 \cite{Haufe2013}.
+We thus only use univariate correlation to investigate how discriminative features are.
 The words corresponding to the top positive and negative correlations are shown in section \autoref{sec:word_party_correlations}.
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
-\begin{tabular}{lccccccc}
-&&& \multicolumn{5}{c}{Predicted Class}\\
+\begin{tabular}{lcc|ccccc}
+&&& \multicolumn{5}{c}{Predicted}\\
 &&& cducsu & fdp& gruene& linke& spd\\
-\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True Class}}}& &cducsu &1186 &289& 178& 198& 179\\
+\hline
+\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}}& &cducsu &1186 &289& 178& 198& 179\\
 &&fdp &882& 658& 236& 329& 214\\
@@
 \quad
-\begin{tabular}{lcccccc}
-&&& \multicolumn{4}{c}{Predicted Class}\\
+\begin{tabular}{lcc|cccc}
+&&& \multicolumn{4}{c}{Predicted}\\
 &&& cducsu & gruene& linke& spd\\
-\multirow{4}{*}{\rotatebox{90}{\pbox{3cm}{\centering True Class}}}&&cducsu&1912& 156& 331& 584\\
+\hline
+\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&1912& 156& 331& 584\\
 &&gruene&2092& 827& 1311& 1444\\
@@
 \label{tab:results_17}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively. Chance level is at 0.2.  
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively.
 }
@@
 \label{tab:results_avg_political_view}
-Classification performance on the 56 class prediction problem of political views, see \autoref{sec:data}; chance performance would be around 0.02.
+Classification performance on the 56 class prediction problem of political views, see \autoref{sec:data}.
 }
@@
 \begin{center}
-\includegraphics[width=8cm]{images/party-sentiments-18.pdf} 
+\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=6cm]{images/party-sentiments-18.pdf} 
 %
@@
 \label{fig:party_sentiments}
-Speech sentiments computed for speeches of each party (parties are ordered according to political power measured in number of seats in the parliament from left to right).
-The more political power a party has, the more positive the speech content is, the fewer seats a party has the more words associated with negative emotions are used in a speech.
+Speech sentiments computed for speeches of each party; parties are ordered according to political power measured in number of seats in the parliament from left to right.
+The more political power a party has, the more positive the speech content is.
+Note that while the SPD (red) was in the opposition in the 17th Bundestag it became part of the government in the 18th Bundestag: their seats in the parliament increased and the average sentiment of their speeches switched sign from negative to overall positive sentiment.
 }
@@
 \begin{center}
-\begin{tabular}{lrr}
+\begin{tabular}{lcc}
    Sentiment vs. &          Gov. Member    &  Seats\\
 \hline\hline
-17th period    &  0.84 & 0.70\\
-18th period    &  0.98 & 0.89\\
+17th Bundestag    &  0.84 & 0.70\\
+18th Bundestag   &  0.98 & 0.89\\
 %

### 2016-03-31 bundestag.py

@@ -92,13 +92,17 @@ def get_sentiments(legis=17):
     sortedParties = sorted(sentiments,key=bundestagSeats[legis].get)
     sortedSentiments = [sentiments[d] for d in sortedParties]
     gov = bundestagGovernment[legis]['government']
-    sortedParties = [p +" (%d)"%bundestagSeats[legis][p]+ "\n" + '[Government]' if p in gov else p+" (%d)"%bundestagSeats[legis][p]+"\n[Opposition]"for p in sortedParties]
+    sortedPartiesLabels = [p +" (%d)"%bundestagSeats[legis][p]+ "\n" + '[Government]' if p in gov else p+" (%d)"%bundestagSeats[legis][p]+"\n[Opposition]"for p in sortedParties]
     import pylab
     pylab.figure(figsize=(7,4))
     #pylab.boxplot(sortedSentiments,labels=sortedParties)
-    pylab.bar(arange(len(sortedSentiments)),[mean(s) for s in sortedSentiments],labels=sortedParties)
+    col = {'gruene':'green','linke':'purple','spd':'red','cducsu':'black','fdp':'yellow'}
+    pylab.bar(0.1+arange(len(sortedSentiments)),[mean(s) for s in sortedSentiments],color=[col[p] for p in sortedParties])
     pylab.grid('on')
-    font = {'family' : 'normal', 'size'   : 10}
+    pylab.ylabel('Average Text Sentiment')
+    pylab.xticks(0.5 + arange(len(sortedSentiments)), sortedPartiesLabels)
+    pylab.xlim([0,len(sortedSentiments)+.01])
+    font = {'family' : 'normal', 'size'   : 12}
     pylab.rc('font', **font)
     pylab.savefig(OUT_DIR+'/party-sentiments-%d.pdf'%legis,bbox_inches='tight')
     # correlation between sentiment and government membership

### 2016-04-01 manuscript/political_bias_prediction.tex
More text

@@
 Assistive technology can help in this context to try and obtain a more unbiased sample of information.
+\\
 
 Ideally one would choose for each topic a sample of reports from the entire political spectrum in order to form an unbiased opinion.
-But ordering and assigning media content with respect to the political spectrum at scale requires automated prediction of political bias.
-The aim of this study is to provide some empirical evidence indicating that political bias prediction is possible with well above chance accuracy.
+But ordering media content with respect to the political spectrum at scale requires automated prediction of political bias.
+And determining a political spectrum that goes beyond the two dimensional left-right scheme is difficult in itself.
+The aim of this study is to provide some empirical evidence indicating that leveraging some valuable open data sources political bias prediction is possible with above chance accuracy.
 \\
@@
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-For each semantic unit, in the case of parliament discussions this were the speeches\footnote{Often speeches were interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.} in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 political view labels.
+First each data set was segmented into semantic units; in the case of parliament discussions this were the speeches, in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 political view labels.
+Parliament speeches were often interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.
 Strings of each semantic unit were tokenized and transformed into bag-of-word vectors as implemented in scikit-learn \cite{scikit-learn}.
-The general idea of bag-of-words vectors is to simply count occurences of words (or word sequences, also called {\em n-grams}) for each data point (usually documents, here parliament speeches).
-So the text of each speech is transformed into a vector $\vec{x}\in\mathds{R}^d$, were $d$ is the size of our dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
+The general idea of bag-of-words vectors is to simply count occurences of words (or word sequences, also called {\em n-grams}) for each data point.
+A data point is usually a documents, here it is the semantic units of parliament speeches and manifesto sentences, respectively.
+The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of our dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
 We tried several options for vectorizing the speeches, including term-frequency-inverse-document-frequency normalisation using n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
@@
 \section{Classification Model and Training Procedure}\label{sec:model}
-
-We used a multinomial logistic regression model in order to classify bag-of-words feature vectors $\vec{x}\in\R^d$, where $d$ is the size of the bag-of-words dictionary.
-Let $y\in\{1,2,\dots,K\}$ be the true party label, where $K$ is the total number of parties in the parliament and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ be the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then
+We used a multinomial logistic regression model in order to classify bag-of-words feature vectors.
+Let $y\in\{1,2,\dots,K\}$ be the true  label (party affiliation or political view, see \autoref{sec:data}), where $K$ is the total number of labels and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ is the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then
 \begin{eqnarray}\label{eq:logreg_multiclass}
@@
 %
-We estimated $\vec{W}$ using quasi-newton gradient descent as packaged in the scikit-learn toolbox.
+We estimated $\vec{W}$ using quasi-newton gradient descent.
 The optimization function was obtained by adding a penalization term to the negative log-likelihood of the multinomial logistic regression objective and the optimization hence found the $\vec{W}$ that minimized
@@
 
-We considered three different classification problems: 1) classification of party affiliation (5 class problem for the 17th legislation period, 4 class problem for the 18th legislation period), 2) classification of government membership (binary problem) and 3) classification of political views as expressed by the political codes of the manifestoproject (56 class problem).
+We considered three different classification problems: 
+\begin{enumerate}
+\item {\bf Classification of party affiliation} (5 class problem for the 17th legislation period, 4 class problem for the 18th legislation period)
+\item {\bf Classification of government membership} (binary problem)
+\item {\bf Classification of political views} as expressed by the political codes of the manifestoproject (56 class problem)
+\end{enumerate}
+
 For each of first two problems, party affiliation and government membership prediction, we trained classifiers on the parliament speeches.
@@
 The model pipeline contained a number of {\em hyperparameters} that we did not want to tune by hand. 
-We first split the training data into a training data set and a test data set for optimizing the hyperparameters, the test data for hyperparameter optimization was 10\% of the entire training data.
-Hyperparameters were optimized using grid search: For each setting of hyperparameters the entire pipeline was trained and evaluated on the test set.
-For the best setting of all hyperparameters, we trained the pipeline again on all training data and evaluated it on the evaluation set.
-Evaluation data was never used for training the models.
+We first split the training data into a training data set that was used for optimization of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data; next to the in-domain test data, we also used out-of domain data wherever possible in order to evaluate how well the model would generalize to text domains that the model has not seen during training.
+The test data for hyperparameter optimization was 10\% of the entire training data.
+Hyperparameters were optimized using grid search and 3-fold cross-validation within the training set only: Within the training set a cross-validation split was made to obtain train/test data for the grid search and for each setting of hyperparameters the entire pipeline was trained and evaluated -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimization.
+For the best setting of all hyperparameters, we trained the pipeline again on all training data and evaluated it on both evaluation data sets.
 For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a legislative period.
-Evaluation data were the texts from party manifestos.
-For the political view prediction setting we first split all labeled manifesto sentences in both legislative periods into a training/test and evaluation set of 90\% (train/test) and 10\% (evaluation), then we split the train/test part again with a 90\%/10\% ratio for training/testing of the hyperparameters.
+Out-of-domain evaluation data were the texts from party manifestos.
+For the political view prediction setting we first split all labeled manifesto sentences in both legislative periods into a training/test and evaluation set of 90\% (train/test) and 10\% (evaluation), then we split the train/test part again with a 90\%/10\% ratio for training/evaluating of the hyperparameters.
 
@@
 \subsection{Predicting political party affiliation}
-
-% Confusion matrix
-\begin{table}[t]\label{tab:conf_mat_four_class}
-\begin{tabular}{lcc|ccccc}
-&&& \multicolumn{5}{c}{Predicted}\\
-&&& cducsu & fdp& gruene& linke& spd\\
-\hline
-\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}}& &cducsu &1186 &289& 178& 198& 179\\
-&&fdp &882& 658& 236& 329& 214\\
-&&gruene &1174& 404& 764& 941& 464\\
-&&linke &388& 92& 214& 806& 201\\
-&&spd &999& 268& 240& 398& 373\\
- \multicolumn{8}{c}{\bf 17th Bundestag}\\
-\end{tabular}
-\quad
-\begin{tabular}{lcc|cccc}
-&&& \multicolumn{4}{c}{Predicted}\\
-&&& cducsu & gruene& linke& spd\\
-\hline
-\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&1912& 156& 331& 584\\
-&&gruene&2092& 827& 1311& 1444\\
-&&linke&596& 186& 1216& 557\\
-&&spd&1284& 226& 563& 916\\
- \multicolumn{7}{c}{\bf 18th Bundestag}\\
-\end{tabular}
-\caption{\label{tab:confusion} Confusion matrices for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
-In particular the green party suffers from this effect: when trained on the speeches in the 18th Bundestag, 2092 statements out of the green party's manifesto were classified as belonging to the cducsu party.
-This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
-\end{table}
+The results for the political party affiliation prediction on held-out parliament data and on evaluation data are shown in \autoref{tab:results_17} for the 17th Bundestag and in \autoref{tab:results_18} for the 18th Bundestag, respectively.
+Shown are the evaluation results for in-domain data (held-out parliament speech texts) as well as the out-of-domain data results obtained from testing the models on the texts extracted from party manifestos.
+When looking at predicting party affiliation from text data from the same domain that was used for training precision and recall values of above 0.6 are obtained.
+For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4.
+\\
 
@@
 }
+\end{table}
+
+In order to investigate the errors the models made we extracted confusion matrices for the predictions on the out-of-domain evaluation data, the manifesto data.
+So the model, trained on party speeches in the parliament was tested on the texts parties published before the elections in order to get elected.
+One example in \autoref{tab:confusion} shows that these mistakes the model makes can give valuable insights into the changes in policy a party decides to make.
+The green party was promoting policies for renewable energy and agains nuclear energy in their manifestos prior to both legislative periods.
+When training a classifier on what representatives in the parliament say however, the statements of the green party are more often predicted to be from the conservative government than from the party that originally promoted these green ideas.
+This effect is also reflected in the precision values on out-of-domain data: The green party has the highest precision values by a large margin in both legislation periods, indicating that there is a relatively high true-positive rate.
+A similar effect was can be observed with the social democratic party (SPD): Evaluating the parliament classifier on manifestos it seems that a significant amount of speeches get classified as conservative government while being from the SPD.
+\\
+
+Next to the party affiliation we also predicted whether or not a text is from a party that belonged to a given governing coalition of the Bundestag.
+In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
+While the in-domain evaluation precision and recall values are much better in this case, reaching values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
+
+% Confusion matrix
+\begin{table}[t]\label{tab:conf_mat_four_class}
+\begin{tabular}{lcc|ccccc}
+&&& \multicolumn{5}{c}{Predicted}\\
+&&& cducsu & fdp& gruene& linke& spd\\
+\hline
+\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}}& &cducsu &1186 &289& 178& 198& 179\\
+&&fdp &882& 658& 236& 329& 214\\
+&&gruene &1174& 404& 764& 941& 464\\
+&&linke &388& 92& 214& 806& 201\\
+&&spd &999& 268& 240& 398& 373\\
+ \multicolumn{8}{c}{\bf 17th Bundestag}\\
+\end{tabular}
+\quad
+\begin{tabular}{lcc|cccc}
+&&& \multicolumn{4}{c}{Predicted}\\
+&&& cducsu & gruene& linke& spd\\
+\hline
+\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&1912& 156& 331& 584\\
+&&gruene&2092& 827& 1311& 1444\\
+&&linke&596& 186& 1216& 557\\
+&&spd&1284& 226& 563& 916\\
+ \multicolumn{7}{c}{\bf 18th Bundestag}\\
+\end{tabular}
+\vspace{1em}
+\caption{\label{tab:confusion} Confusion matrices for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+In particular the green party suffers from this effect: when trained on the speeches in the 18th Bundestag, 2092 statements out of the green party's manifesto were classified as belonging to the cducsu party.
+This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
 \end{table}
@@
 \section{Conclusions and Outlook}\label{sec:conclusion}
-
+Note that not in all settings we had this type of data, and also for the cases where we can get out-of-domain evaluation data, there will be quite some non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
 %-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate

### 2016-04-02 manuscript/political_bias_prediction.tex
More text

@@
 \begin{tabular}{lcc|ccccc}
+ \multicolumn{8}{c}{\bf 17th Bundestag}\\
+ \\
 &&& \multicolumn{5}{c}{Predicted}\\
@@
 &&spd &999& 268& 240& 398& 373\\
- \multicolumn{8}{c}{\bf 17th Bundestag}\\
 \end{tabular}
@@
 \begin{tabular}{lcc|cccc}
+ \multicolumn{7}{c}{\bf 18th Bundestag}\\
+ \vspace{1em}\\
 &&& \multicolumn{4}{c}{Predicted}\\
@@
 &&spd&1284& 226& 563& 916\\
- \multicolumn{7}{c}{\bf 18th Bundestag}\\
 \end{tabular}
@@
 \subsection{Speech sentiment correlates with political power}\label{sec:sentiment_result}
-We analysed the sentiment of each speech of a political party and plot the results in \autoref{fig:party_sentiments}.
-While the methodology used for estimating the sentiment (see \autoref{sec:sentiment_analysis_methods}), it can yield some insight into the emotional content of a speech.
+In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
+The average sentiment of each political party is shown in \autoref{fig:party_sentiments} for the two legislative periods under investigation.
+While the methodology used for estimating the sentiment (see \autoref{sec:sentiment_analysis_methods}) is simple and limited to a one dimensional spectrum of sentiment, it can yield some insight into the emotional content of a speech.
 High values indicate more pronounced usage of positive words, whereas negative values indicate more pronounced usage of words associated with negative emotional content.
-We can quantify political power as the number of seats a party has in the parliament.
-Our analyses show a strong trend supporting the hypothesis that the more political power a party has, the more positive sentiment their speech has, and vice versa, the less seats a party has, the more negative content their speeches contain.
-This pattern is evident from the barplot in \autoref{fig:party_sentiments} and supported by a high correlation of 0.86 between number of parliament seats and sentiment index.
+
+The results show an interesting relationship between political power and sentiment.
+Political power was evaluated in two ways: a) in terms of the number of seats a party has and b) in terms of membership of the government.
+Correlating either of these two proxies of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
+This pattern is evident from the barplot in \autoref{fig:party_sentiments} and in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+
+Note that there is one party, the social democrats (SPD), which have a large number of seats and were in the opposition in the 17th Bundestag and member of the government in the 18th Bundestag: The average sentiment of this party switched sign from negative to positive when they joined the government, indicating that positive sentiment is a strong indicator of government membership of affiliation when analysing a text.
 
@@
 \caption{
-\label{tab:results_avg_political_view}
+\label{tab:sentiments}
 Correlation coefficient between average sentiment of political speeches of a party in the german bundestag with two measures of their political power, a) their membership in the government and b) with the number of seats that party occupies in the parliament.
@@
 
-\section{Conclusions and Outlook}\label{sec:conclusion}
-Note that not in all settings we had this type of data, and also for the cases where we can get out-of-domain evaluation data, there will be quite some non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
+\subsection{An example web application}
+We implemented an example web application that downloads regularly all articles from major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}}, applies some simple topic modelling to them and assigns articles to a topic and predicts the political view of an article using the manifesto code classifier trained on all party manifestos.
+Within each topic it is then possible to get an order (from left to right) overview of the opinions on that topic.
+An example of one topic that emerged on March 31st is shown in \autoref{fig:fipi}.
+The demo is live and running\footnote{\url{http://to-be-included-in-camera-ready-version.com}} and the code is open sourced an available on git-hub\footnote{\url{http://to-be-included-in-camera-ready-version.com}}.
+\begin{figure}
+\begin{center}
+\includegraphics[width=10cm]{images/fipi-screenshot}
+%
+\end{center}
+\caption{
+\label{fig:fipi}
+A screen shot of an example web application using the political view prediction combined with topic modelling to provide a heterogeneous overview of a topic. }
+\end{figure}
+
+\section{Conclusions, Limitations and Outlook}\label{sec:conclusion}
+The goal of this study was to present a simple approach for automating political bias prediction and to quantify how well this is possible.
+The results of these experiments show that automated political bias prediction is possible with above chance accuracy.
+Moreover there are some interesting aspects of these analyses that could be interesting for social science researchers, too.
+
+For one the misclassifications of a model can be related to the changes in policy of a party.
+This can be helpful to quantitatively back up such a change in policy with more than just anecdotal evidence from single texts.
+Second analysing the word-party correlations shows some discriminative words that are also used by the classifier to determine political bias.
+This insight can be related to the political views of a party and allows for validation of the models by human experts.
+Third when correlating the sentiment of a speech with proxies of political power there is a strong and significant positive correlation between political power and positive sentiment.
+This supports the (probably trivial) intuition that if we read a text that has a positive opinion on what the party in power does, the author of the text is probably affiliated with those in power.
+While such an insight of something rather obvious in itself might not be very useful, the technology behind it is: Sentiment analysis, and in particular such a simple approach as used here, can be easily automated and scaled up to massive amounts of text data.
+Having established such a way of relating this aspect of media content to the political bias allows to automate political bias direction using the proxy of sentiments, which are a rather domain-independent measure.
+
+One of the main limitations of an automated political bias prediction system is the availability of training data.
+Every training data set that is publicly available has an inherent bias as it is taken from a different domain.
+This study tried to quantify the impact of this, but not in all settings we had out-of-domain data to quantify generalization performance across text domains.
+For the cases where we had evaluation data from two domains there was a pronounced shift in performance between in-domain evaluation set and out-of-domain evaluation set.
+This was to be expected, given that the texts are written for a different audience and in completely different styles.
+This results in quite some non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
+Another important issue was that the semantic units in the case of the out-of-domain data were of much shorter length (often just parts of a sentence) and hence much noisier than the in-domain evaluation data.
+\\
+
+Taken together the results show that when training on one domain and evaluating on an entirely different domain one can still obtain above chance performance, even though there is a significant drop in performance.
+The generalization performance across domains will be an important topic of future research.
+The aim of this study was mainly to assess the impact of this effect, but if the goal is to minimize the generalization performance the simplest solution, next to more complicated approaches from semisupervised learning, non-stationary methods or covariate shift approaches, is to simply include data from different -- possibly very heterogeneous -- domains in the training data set.
+A popular approach is to integrate data from social networks such as Facebook, as advocated for instance in \cite{Arzheimer2016}.
+\\
+
+All data sets used in this study were publicly available, all code for experiments and the live web application can be found online \cite{fipi, fipidemo}.
 %-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate
 
-\small
 \subsection*{Acknowledgements}
-
+This study would not have been possible without the open data initiatives behind the code and projects that led to the publicly available data sets and APIs described in \autoref{sec:data}.
 %

### 2016-04-02 manuscript/political_bias_prediction.tex
more text, more images.

@@
 
-\subsection{Analysing bag-of-words features}
+\subsection{Analysing bag-of-words features}\label{sec:correlations_methods}
 In order to get  better picture of which features contribute to the classification we computed correlation coefficients between each word and each party.
@@
 
-\autoref{fig:party_word_correlations} shows the correlations of individual words and each party.
-We plotted the top 20 words for each sign, i.e.
-the 20 words with the strongest positive and negative correlations for each party, respectively.
-While the correlations do indicate that there are -- despite extensive and careful cleaning procedures -- still words with strong correlations in the top words, we conjecture that at least some of these words do have some discriminatory power, even if they do not refer to political views.
-Below we list some examples, but the most prominent is probably: the opposition often addresses the government in their speeches, the govermental parties do that less often.
-This is reflected in the frequency of the usage of words that refer to members of the government or ministers.
-But among the often used and avoided words we also find words that refer to actual political content.
-Thus we hypothesize that to some extent these correlation coefficients can give hints as to what the bag-of-words histogram representation into which we transformed the speeches are related to the actual content of the speeches.
+In order to enable better interpretation of the results, we investigated the univariate dependency between single text features and the party affiliation labels of texts, see also \autoref{sec:correlations_methods}.
+The highest and lowest correlations between individual words and each party are shown in \autoref{fig:party_word_correlations}.
+While the correlations do indicate that there are -- despite extensive and careful cleaning procedures -- still some unspecific words in the top words, we conjecture that at least some of these words can be useful for a better interpretation of the model and the results, even if not every word can be easily associated with political views.
+
 In the following we give some examples of words that appear to be preferentially used or avoided by each respective party.
-Even though interpretations of these quantitative results are somewhat invalid, as they neglect the context in which these words were mentioned, we find some interesting patterns.
+Even though interpretations of these quantitative results are somewhat problematic, as they neglect the context in which these words were mentioned, we find some interesting patterns.
+
+\paragraph{\bf Green party (Gruene)}
+
+\paragraph{\bf Conservative party (CDU/CSU)}
+
+\paragraph{\bf Social democratic party (SPD)}
+
+\paragraph{\bf Left party (linke)}
 
@@
 \begin{center}
-%\includegraphics[width=3.5cm]{party_word_correlations-linke.pdf} 
-%\includegraphics[width=3.1cm]{party_word_correlations-gruene.pdf} 
-%\includegraphics[width=3.45cm]{party_word_correlations-spd.pdf} 
-%\includegraphics[width=3.65cm]{party_word_correlations-cdu.pdf}
+\includegraphics[width=2.7cm]{images/party_word_correlations-linke-17.pdf} 
+\includegraphics[width=3cm]{images/party_word_correlations-gruene-17.pdf} 
+\includegraphics[width=2.8cm]{images/party_word_correlations-spd-17.pdf} 
+\includegraphics[width=2.8cm]{images/party_word_correlations-cducsu-17.pdf}
 %

### 2016-04-02 bundestag.py

@@ -94,7 +94,7 @@ def get_sentiments(legis=17):
     gov = bundestagGovernment[legis]['government']
     sortedPartiesLabels = [p +" (%d)"%bundestagSeats[legis][p]+ "\n" + '[Government]' if p in gov else p+" (%d)"%bundestagSeats[legis][p]+"\n[Opposition]"for p in sortedParties]
     import pylab
-    pylab.figure(figsize=(7,4))
+    pylab.figure(figsize=(1.5*len(sortedParties),4))
     #pylab.boxplot(sortedSentiments,labels=sortedParties)
     col = {'gruene':'green','linke':'purple','spd':'red','cducsu':'black','fdp':'yellow'}
     pylab.bar(0.1+arange(len(sortedSentiments)),[mean(s) for s in sortedSentiments],color=[col[p] for p in sortedParties])

### 2016-04-02 manuscript/political_bias_prediction.tex
more text, polished figures, pre-final

@@
 \begin{center}
-\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=6cm]{images/party-sentiments-18.pdf} 
+\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf} 
 %
@@
 
-In order to enable better interpretation of the results, we investigated the univariate dependency between single text features and the party affiliation labels of texts, see also \autoref{sec:correlations_methods}.
-The highest and lowest correlations between individual words and each party are shown in \autoref{fig:party_word_correlations}.
+While interpretability of linear models is often propagated as one of their main advantages, doing so naively without modelling the noise covariances can lead to wrong conclusions, see e.g.
+\cite{Zien2009, Haufe2013} -- essentially interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; as this assumption is violated by our data simple univariate correlation measures are used to interpret the features used in the linear predictors, see also \autoref{sec:correlations_methods}.
+The 10 highest and lowest correlations between individual words and each party are shown in \autoref{fig:party_word_correlations}.
 While the correlations do indicate that there are -- despite extensive and careful cleaning procedures -- still some unspecific words in the top words, we conjecture that at least some of these words can be useful for a better interpretation of the model and the results, even if not every word can be easily associated with political views.
@@
 
+\paragraph{\bf Left party (linke)}
+The left party mostly criticises measures that affect social welfare negatively, such as the {\em Hartz IV} program.
+Main actors that are blamed for decisions of the conservative governments by the left party are big companies ({\em konzerne}).
+Rarely the party addresses concerns related to security ({\em sicherheit}).
+
 \paragraph{\bf Green party (Gruene)}
+The green party heavily criticised the secret negotiations about the TiSA agreement\footnote{\url{https://en.wikipedia.org/wiki/Trade_in_Services_Agreement}} and insists in formal inquiries that the representatives of the green party put forward in this matter ({\em fragen, anfragen}).
+They also often put forward critical questions related to army projects ({\em R\"ustungsprojekte, Wehrbericht}) or the military development in east europe ({\em  Jalta}\footnote{Referring to the \url{https://en.wikipedia.org/wiki/Yalta_Conference}}).
+
+\paragraph{\bf Social democratic party (SPD)}
+The social democrats often use words related to rights of the working class, as reflected by the heavy use of the {\em International Labour Organisation} (ILO) or rights of employes ({\em Arbeitnehmerrechte}).
+They rarely talk about competition ({\em Wettbewerb}) or climate change ({\em klimapolitik}).
 
 \paragraph{\bf Conservative party (CDU/CSU)}
-
-\paragraph{\bf Social democratic party (SPD)}
-
-\paragraph{\bf Left party (linke)}
+The conservative christian party often uses words related to a pro-economy attitute, such as competitiveness ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
+The latter refers to the ongoing debates about whether or not the governments should be allowed to collect data and thus restrict fundamental civil rights in order to better secure the population from terrorist attacks.
+However the party rarely uses the word war ({\em krieg}).
 
@@
 \begin{center}
-\includegraphics[width=2.7cm]{images/party_word_correlations-linke-17.pdf} 
-\includegraphics[width=3cm]{images/party_word_correlations-gruene-17.pdf} 
-\includegraphics[width=2.8cm]{images/party_word_correlations-spd-17.pdf} 
-\includegraphics[width=2.8cm]{images/party_word_correlations-cducsu-17.pdf}
+\includegraphics[width=2.8cm]{images/party_word_correlations-linke-18.pdf} 
+\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-18.pdf} 
+\includegraphics[width=3cm]{images/party_word_correlations-spd-18.pdf} 
+\includegraphics[width=3cm]{images/party_word_correlations-cducsu-18.pdf}
 %
@@
 \label{fig:party_word_correlations}
-Correlations between words and party for parliament speeches. }
+Correlations between words and party affiliation label for parliament speeches can help interpreting the features used by a predictive model.
+Shown are the top 10 positively and negatively correlated text features for the current Bundestag.
+For interpretations see \autoref{sec:word_party_correlations}.}
 \end{figure}
@@
 \subsection{An example web application}
-We implemented an example web application that downloads regularly all articles from major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}}, applies some simple topic modelling to them and assigns articles to a topic and predicts the political view of an article using the manifesto code classifier trained on all party manifestos.
+To show an example use case of the above approach a web application was implemented that downloads regularly all articles from major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}}, applies some simple topic modelling to them and assigns articles to a topic and predicts the political view of an article using the manifesto code classifier trained on all party manifestos.
 Within each topic it is then possible to get an order (from left to right) overview of the opinions on that topic.
+The left right index is computed according to the outputs of the manifestocodes and deterministic left-right mapping defined by human experts \cite{leftright}.
 An example of one topic that emerged on March 31st is shown in \autoref{fig:fipi}.

### 2016-04-02 bundestag.py

@@ -147,7 +147,7 @@ def get_word_correlations(legis=17):
         wordCors[pa] = dict([(wordidx2word[x],co[x]) for x in co.nonzero()[0] if wordidx2word[x] not in stops])
     json.dump(dict(wordCors),open(OUT_DIR+'/wordCors-%d.json'%legis,'wb'))
 
-def list_top_words(legis=17,topwhat=20):
+def list_top_words(legis=17,topwhat=10):
     import pylab
     fn = OUT_DIR+'/wordCors-%d.json'%legis
     cors = json.loads(open(fn).read())
@@ -156,14 +156,16 @@ def list_top_words(legis=17,topwhat=20):
     stops = get_stops(legislationPeriod=legis)
     
     for party in cors.keys():
-        pylab.figure(figsize=(3,12))
-        tmp = cors[party][:topwhat] + [('...',0)] +  cors[party][-topwhat:]
-        words = [x[0] for x in tmp if x[0] not in stops]
-        wordcors = [x[1] for x in tmp if x[0] not in stops]
+        pylab.figure(figsize=(3,10))
+        c = [x for x in cors[party] if x[0] not in stops]
+        tmp = c[:topwhat] + [('...',0)] +  c[-topwhat:]
+        words = [x[0] for x in tmp]
+        wordcors = [x[1] for x in tmp]
         pylab.barh(arange(len(words)),wordcors,color=colors[party])
         pylab.yticks(arange(len(words)),words)
         pylab.ylim(-.2,len(words))
-        pylab.xticks([-.2,0,.2])
+        xl = max(abs(array(wordcors))) + .01
+        pylab.xticks([-xl,0,xl])
         pylab.title(party)
         pylab.xlabel('Correlation')
         font = {'family' : 'normal', 'size'   : 16}

### 2016-04-02 bundestag.py
Added topic aggregation for manifesto evaluation data.

@@ -197,6 +197,19 @@ def get_raw_text(folder="data", legislationPeriod=18):
     files = filter(lambda x: x.split('/')[-1].split('_')[0] in partyIds,files)
     return zip(*chain(*filter(None,map(csv2DataTuple,files))))
 
+def get_raw_text_topics(folder="data", legislationPeriod=18):
+    '''
+    Loads raw text and labels from manifestoproject csv files
+    (Downloaded from https://visuals.manifesto-project.wzb.eu)
+    '''
+    parties = bundestagParties[legislationPeriod]
+    partyIds = [str(partyManifestoMap[p]) for p in parties]
+    year = '2013'
+    if legislationPeriod==17:year='2009'
+    files = glob.glob(folder+"/[0-9]*_%s.csv"%year)
+    files = filter(lambda x: x.split('/')[-1].split('_')[0] in partyIds,files)
+    return zip(*chain(*filter(None,map(csv2DataTupleTopics,files))))
+
 def csv2DataTuple(f):
     '''
     Extracts list of tuples of (text,label) for each manifestoproject file
@@ -206,6 +219,18 @@ def csv2DataTuple(f):
     partyId = f.split('/')[-1].split('_')[0]
     party = [k for (k,v) in partyManifestoMap.items() if str(v) == partyId]
     return zip(df['content'].tolist(), party * len(df))
+
+def csv2DataTupleTopics(f):
+    '''
+    Extracts list of tuples of (text,label) for each manifestoproject file
+    '''
+    df = pd.read_csv(f)
+    df['content'] = df['content'].astype('str')
+    df['topic'] = ((df['cmp_code'].dropna()) / 100).apply(lambda x: int(x))
+    partyId = f.split('/')[-1].split('_')[0]
+    party = [k for (k,v) in partyManifestoMap.items() if str(v) == partyId]
+    contentByTopic = df.groupby('topic')['content'].sum().tolist()
+    return zip(contentByTopic, party * len(contentByTopic))
 
 def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='default'):
     stops = get_stops()

### 2016-04-03 manuscript/political_bias_prediction.tex
Added topic aggregated manifesto party classification.

@@
 \subsection{Predicting political views}
+Parties change their policies and positions in the political spectrum.
+More reliable categories for political bias are party independent labels for political standpoints.
+These are provided by social scientists in the manifesto project \cite{manifesto}.
+A separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:models} on the manifesto code labels.
+As we do not have an out-of-domain evaluation set for this type of experiment we only report evaluation error.
+Note however that also in this experiment the evaluation data was never seen by any model during training time.
+Note that a precision of close to 0.5 is rather high considering the large number of labels.
 
@@
 \hline\hline
-test set    &  0.48   &   0.47  &    0.47  &    2651\\
 evaluation set    &  0.47    &  0.46 &     0.46 &     2946\\

### 2016-04-03 manuscript/political_bias_prediction.tex
Pre-final draft with all figures and tables.

@@
 For the analyses based on political view labels we considered all parties, also those that did not make it into the parliament.
+\\
+The length of each annotated statement in the party manifestos was rather short.
+The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters, measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.
+This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos.
+However for a classifier (as well as for humans) such short sentences can be rather difficult to classify.
+In order to obtain less 'noisy' data points from each party all statements were aggregated into political topics using the manifesto code labels: each political view label is a three digit code, the first digit represents the political domain.
+In total there were eight political domains (topics): {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups} and a topic {\em undefined}, for a complete list see also \cite{leftright}.
+These 8 topics were used to aggregate all statements in each manifesto into topics.
+Most party manifestos covered all eight of them, some party manifestos in the 17th Bundestag only covered seven.
 
@@
 \begin{enumerate}
-\item {\bf Classification of party affiliation} (5 class problem for the 17th legislation period, 4 class problem for the 18th legislation period)
-\item {\bf Classification of government membership} (binary problem)
-\item {\bf Classification of political views} as expressed by the political codes of the manifestoproject (56 class problem)
+\item {\bf Classification of party affiliation} 
+\item {\bf Classification of government membership}
+\item {\bf Classification of political views}
 \end{enumerate}
 
+Party affiliation is a five class problem for the 17th legislation period, and a four class problem for the 18th legislation period.
+Government membership is a binary problem.
+Political view classification is based on the labels of the manifestoproject, see also \autoref{sec:data}, and is a 56 class problem.
 For each of first two problems, party affiliation and government membership prediction, we trained classifiers on the parliament speeches.
@@
 Out-of-domain evaluation data were the texts from party manifestos.
-For the political view prediction setting we first split all labeled manifesto sentences in both legislative periods into a training/test and evaluation set of 90\% (train/test) and 10\% (evaluation), then we split the train/test part again with a 90\%/10\% ratio for training/evaluating of the hyperparameters.
+For the political view prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
 
@@
 When looking at predicting party affiliation from text data from the same domain that was used for training precision and recall values of above 0.6 are obtained.
+\\
+
 For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4.  
-\\
+A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
+In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, we aggregated the manifesto data based on the topic.
+The manifesto codes contain labels for different topics, which were used to concatenate the data points of each party to only eight data points per party, each data point corresponding to one of the eight topics defined and annotated by the manifesto project annotators.
+The results are shown in table \autoref{tab:results_17_topic} and \autoref{tab:results_18_topic} and clearly show that when the texts to be classified and the word count statistics are sufficiently dense the classification performance on out of domain data becomes close to precisions of 1.0.
 
@@
 
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lcccc}
+& \multicolumn{4}{c}{\bf 17th Bundestag} \\
+    &         precision    &recall &  f1-score  & N  \\
+    \hline
+cducsu     &  0.64  &    1.00  &    0.78    &     7\\
+       fdp    &   1.00    &  1.00    &  1.00    &     7\\
+    gruene  &     1.00  &    0.86  &    0.92    &     7\\
+     linke    &   1.00   &   1.00     & 1.00    &     7\\
+       spd   &    0.80   &   0.50    &  0.62     &    8\\
+    \hline
+    \hline
+    avg / total  &     0.88   &   0.86   &   0.86  &      36\\
+\end{tabular}
+\quad
+\begin{tabular}{lcccc}
+& \multicolumn{4}{c}{\bf 18th Bundestag} \\
+    &         precision    &recall &  f1-score  & N  \\
+    \hline
+  cducsu     &  0.50   &   1.00  &    0.67     &    8\\
+       gruene  &     0.00   &   0.00   &   0.00  &       8\\
+      linke     &  1.00   &   0.88   &   0.93  &       8\\
+        spd     &  0.56  &    0.62  &    0.59   &      8\\
+
+avg / total   &    0.51 &     0.62   &   0.55   &     32\\
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:results_topic}
+{\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
+Instead of sentence level predictions as in \autoref{fig:results_17}, here predictions per political topic are shown (see \autoref{sec:data}); in contrast to single sentence level predictions (see \autoref{tab:confusion}) the predictions made on topic level are in many cases reliable.
+One effect present in the single sentence level predicitions however is even more pronounced in the topic level predictions: the policy of the green party manifesto is completely absorbed in the parliament speeches of the right-conservative CDU/CSU party -- all manifesto topics of the green party are predicted to be from one of the parties of the governing coalition, CDU/CSU or SPD.}
+\end{table}
+
+%trained with cv on bundestag speeches 17
+%tested on topic aggregated (first digit of manifestocode) manifesto data
+%
+%cducsu       0.64      1.00      0.78         7
+%        fdp       1.00      1.00      1.00         7
+%     gruene       1.00      0.86      0.92         7
+%      linke       1.00      1.00      1.00         7
+%        spd       0.80      0.50      0.62         8
+%
+%avg / total       0.88      0.86      0.86        36
+%
+%Confusion Matrix (rows=true, cols=predicted)
+%cducsu, fdp, gruene, linke, spd
+%[7, 0, 0, 0, 0]
+%[0, 7, 0, 0, 0]
+%[0, 0, 6, 0, 1]
+%[0, 0, 0, 7, 0]
+%[4, 0, 0, 0, 4]
+%
+%0.86111111111111116
+%
+%trained with cv on bundestag speeches 18
+%tested on topic aggregated (first digit of manifestocode) manifesto data
+%
+%*** Training Set (CVd) ***
+%             precision    recall  f1-score   support
+%
+%     cducsu       0.50      1.00      0.67         8
+%     gruene       0.00      0.00      0.00         8
+%      linke       1.00      0.88      0.93         8
+%        spd       0.56      0.62      0.59         8
+%
+%avg / total       0.51      0.62      0.55        32
+%
+%Confusion Matrix (rows=true, cols=predicted)
+%cducsu, fdp, gruene, linke, spd
+%[8, 0, 0, 0]
+%[4, 0, 0, 4]
+%[1, 0, 7, 0]
+%[3, 0, 0, 5]
+%
+
 In order to investigate the errors the models made we extracted confusion matrices for the predictions on the out-of-domain evaluation data, the manifesto data.
@@
 
-% Confusion matrix
+% Confusion matrix sentence level
 \begin{table}[t]\label{tab:conf_mat_four_class}
@@
 \end{tabular}
-\quad
+\hfill
 \begin{tabular}{lcc|cccc}
@@
 \vspace{1em}
-\caption{\label{tab:confusion} Confusion matrices for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
 In particular the green party suffers from this effect: when trained on the speeches in the 18th Bundestag, 2092 statements out of the green party's manifesto were classified as belonging to the cducsu party.
 This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
+\end{table}
+
+% Confusion matrix topic level
+\begin{table}[t]\label{tab:conf_mat_four_class}
+\begin{tabular}{lcc|ccccc}
+ \multicolumn{8}{c}{\bf 17th Bundestag}\\
+ \\
+&&& \multicolumn{5}{c}{Predicted}\\
+&&& cducsu & fdp& gruene& linke& spd\\
+\hline
+\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}} &&cducsu &7& 0& 0& 0& 0\\
+&&fdp&0& 7& 0& 0& 0\\
+&&gruene&0& 0& 6& 0& 1\\
+&&linke&0& 0& 0& 7& 0\\
+&&spd&4& 0& 0& 0& 4\\
+\end{tabular}
+\hfill
+\begin{tabular}{lcc|cccc}
+ \multicolumn{7}{c}{\bf 18th Bundestag}\\
+ \vspace{1em}\\
+&&& \multicolumn{4}{c}{Predicted}\\
+&&& cducsu & gruene& linke& spd\\
+\hline
+\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&8& 0& 0& 0\\
+&&gruene&4& 0& 0& 4\\
+&&linke&1& 0& 7& 0\\
+&&spd&3& 0& 0& 5\\
+\end{tabular}
+\vspace{1em}
+\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); for interpretations see also \autoref{tab:results_topic}.}
 \end{table}
@@
 Note however that also in this experiment the evaluation data was never seen by any model during training time.
-Note that a precision of close to 0.5 is rather high considering the large number of labels.
+Despite the rather low values, a precision of close to 0.5 is rather high considering the large number of labels and the difficulty of the problem. \\
 
@@
 \begin{center}
-\begin{tabular}{lrrrr}
-    &         precision    &recall &  f1-score  & N\\
+\begin{tabular}{lcrrrr}
+code & meaning  &      precision    &recall &  f1-score  & N\\
 \hline\hline
-evaluation set    &  0.47    &  0.46 &     0.46 &     2946\\
+%
+       501  & environmentalism + &      0.62   &   0.61 &     0.61  &     165\\
+        202 &   democracy + &   0.58  &    0.55   &   0.57   &    122\\
+        701    & labour +&  0.57  &    0.54   &   0.56      & 129\\
+                201    &freedom/human rights +  & 0.58   &   0.54   &   0.56   &    159\\
+         106   & peace + & 0.52&      0.57   &   0.55    &    21\\
+\dots\\
+        302     &centralism + & 0.25     & 0.20  &    0.22  &      10\\
+        401    &  free enterprise + &0.20    &  0.19    &  0.20   &     52\\
+        505    &welfare - &  0.13   &   0.14   &   0.14   &     14\\
+        409    & keynesian demand +&  0.14  &    0.12  &    0.13   &      8\\
+            0    & undefined &  0.09  &    0.12   &   0.10    &    17\\
+avg / total    &  0.47    &  0.46 &     0.46 &     2946\\
 %

### 2016-04-03 manuscript/political_bias_prediction.tex
Polished text.

@@
 \author{
-%Felix Bie\ss{}mann%
+Felix Biessmann
 \thanks{}
@@
 Standard text features extracted from speeches and manifestos of political parties are used to predict political bias in terms of political party affiliation and in terms of political views.
-Results indicate that political bias can be predicted with well above chance accuracy.
+Results indicate that political bias can be predicted with above chance accuracy.
 Mistakes of the model can be interpreted with respect to changes of policies of political actors.
@@
 
-Automated predictions of political bias can be problematic for ethical reasons.
-One trusts political experts, as they will take responsibility for what they say about a text, and they can explain their decisions.
+When human experts determine political bias of texts they will take responsibility for what they say about a text, and they can explain their decisions.
 This is a key difference to many statistical learning approaches.
 Not only is the responsibility question not entirely clear, it can also be difficult to interpret some of the decisions.
-In order to validate the predictions of the models we devise two strategies that allow for better interpretations of the models.
-First we are looking at univariate measures of how discriminative the features are that the model uses.
-This somewhat allows to relate the political texts to the kind of information that political experts use, too.
-Second we use sentiment analysis to investigate whether this aspect of language also has discriminatory power.
+In order to validate the predictions of the models three strategies that allow for better interpretations of the models are proposed.
+First interpretations of the model misclassifications are related to changes in party policies.
+Second univariate measures of correlation between text features and party affiliation allow to relate the predictions to the kind of information that political experts use for interpreting texts.
+Third sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
 \\
 
-In the following \autoref{sec:data} gives an overview of data acquisition and preprocessing, \autoref{sec:model} presents the model, in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+In the following \autoref{sec:data} gives an overview of data acquisition and preprocessing, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
 
@@
 
-\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy of political bias.
-The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}, we used an open source API to that data to query the data in cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
-This API returns for each speech the speaker and its party affiliation.
+\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
+The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}, we used an open source API to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
 In total we obtained 22784 speeches for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.
@@
 The data released in this project comprises the complete manifestos for each party that ran for election enriched with annotations by political experts.
-Each sentence (in some cases also parts of sentences) are labels with one of 56 political labels.
-Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare, ...}.
+Each sentence (in some cases also parts of sentences) is annotated with one of 56 political labels.
+Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare, ...}, for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
 The set of labels was developed by political scientists at the WZB and released for public use.
 We obtained all manifestos of parties that were running for election in this and the last legislative period.
-In total this resulted in 29451 political statements that had two types of labels, for one we stored the party affiliation of each political statement.
-This party affiliation label was used to evaluate the party evaluation classifiers trained on the parliament speeches.
+In total this resulted in 29451 political statements that had two types of labels: First the party affiliation of each political statement which was used to evaluate the party evaluation classifiers trained on the parliament speeches.
 For this purpose we constrained the data to only those parties that were elected into the parliament.
-Next to the party affiliation we were mainly interested in the political view labels annotated by human experts.
+Next to the party affiliation the political view labels were extracted.
 For the analyses based on political view labels we considered all parties, also those that did not make it into the parliament.
-\\
+
 The length of each annotated statement in the party manifestos was rather short.
@@
 However for a classifier (as well as for humans) such short sentences can be rather difficult to classify.
-In order to obtain less 'noisy' data points from each party all statements were aggregated into political topics using the manifesto code labels: each political view label is a three digit code, the first digit represents the political domain.
+In order to obtain less 'noisy' data points from each party all statements were aggregated into political topics using the manifesto code labels.
+Each political view label is a three digit code, the first digit represents the political domain.
 In total there were eight political domains (topics): {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups} and a topic {\em undefined}, for a complete list see also \cite{leftright}.
@@
 Parliament speeches were often interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.
-Strings of each semantic unit were tokenized and transformed into bag-of-word vectors as implemented in scikit-learn \cite{scikit-learn}.
-The general idea of bag-of-words vectors is to simply count occurences of words (or word sequences, also called {\em n-grams}) for each data point.
+Strings of each semantic unit were tokenised and transformed into bag-of-word vectors as implemented in scikit-learn \cite{scikit-learn}.
+The general idea of bag-of-words vectors is to simply count occurrences of words (or word sequences, also called {\em n-grams}) for each data point.
 A data point is usually a documents, here it is the semantic units of parliament speeches and manifesto sentences, respectively.
-The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of our dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
-We tried several options for vectorizing the speeches, including term-frequency-inverse-document-frequency normalisation using n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
+The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of the dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
+We tried several options for vectorizing the speeches, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
 All of these hyperparameters were subjected to hyperparameter optimization as explained in \autoref{sec:crossvalidation}.
@@
 \section{Classification Model and Training Procedure}\label{sec:model}
-We used a multinomial logistic regression model in order to classify bag-of-words feature vectors.
-Let $y\in\{1,2,\dots,K\}$ be the true  label (party affiliation or political view, see \autoref{sec:data}), where $K$ is the total number of labels and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ is the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then
+Bag-of-words feature vectors were used to train a multinomial logistic regression model.
+Let $y\in\{1,2,\dots,K\}$ be the true  label, where $K$ is the total number of labels and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ is the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then
 \begin{eqnarray}\label{eq:logreg_multiclass}
@@
 
-We considered three different classification problems: 
+Three different classification problems were considered: 
 \begin{enumerate}
-\item {\bf Classification of party affiliation} 
-\item {\bf Classification of government membership}
-\item {\bf Classification of political views}
+\item {\bf Classification of party affiliation} (five class / four class problem)
+\item {\bf Classification of government membership} (binary problem)
+\item {\bf Classification of political views} (56 class problem)
 \end{enumerate}
@@
 Party affiliation is a five class problem for the 17th legislation period, and a four class problem for the 18th legislation period.
-Government membership is a binary problem.
-Political view classification is based on the labels of the manifestoproject, see also \autoref{sec:data}, and is a 56 class problem.
+Political view classification is based on the labels of the manifesto project, see \autoref{sec:data} and \cite{leftright}.
 For each of first two problems, party affiliation and government membership prediction, we trained classifiers on the parliament speeches.
-For the third problem we trained only on the manifesto data as only for those the political view labels were available.
+For the third problem we trained only on the manifesto data for which political view labels were available.
 
 \subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
-The model pipeline contained a number of {\em hyperparameters} that we did not want to tune by hand. 
-We first split the training data into a training data set that was used for optimization of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data; next to the in-domain test data, we also used out-of domain data wherever possible in order to evaluate how well the model would generalize to text domains that the model has not seen during training.
-The test data for hyperparameter optimization was 10\% of the entire training data.
-Hyperparameters were optimized using grid search and 3-fold cross-validation within the training set only: Within the training set a cross-validation split was made to obtain train/test data for the grid search and for each setting of hyperparameters the entire pipeline was trained and evaluated -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimization.
-For the best setting of all hyperparameters, we trained the pipeline again on all training data and evaluated it on both evaluation data sets.
-For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a legislative period.
+The model pipeline contained a number of {\em hyperparameters} that were optimised using cross-validation.  
+We first split the training data into a training data set that was used for optimisation of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data; wherever possible the generalisation performance of the models was also evaluated on out-of domain data.
+Hyperparameters were optimised using grid search and 3-fold cross-validation within the training set only: A cross-validation split was made to obtain train/test data for the grid search and for each setting of hyperparameters the entire pipeline was trained and evaluated -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimisation.
+For the best setting of all hyperparameters the pipeline was trained again on all training data and evaluated on the evaluation data sets.
+For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a given legislative period.
 Out-of-domain evaluation data were the texts from party manifestos.
@@
 \subsection{Sentiment analysis}\label{sec:sentiment_analysis_methods}
-We used a publicly available key word list to extract sentiments from the text subjected to classification \cite{remquahey2010}.
-The sentiment index used for attributing positive or negative sentiment to a text was computed  as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which was constructed by simply taking the values from the sentiment dictionary.
+A publicly available key word list was used to extract sentiments from the text subjected to classification \cite{remquahey2010}.
+The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which constructed from the values in the sentiment dictionary.
 
-\subsection{Analysing bag-of-words features}\label{sec:correlations_methods}
-In order to get  better picture of which features contribute to the classification we computed correlation coefficients between each word and each party.
-Often linear models are interpreted by just looking at the weights of the model.
-Since the text features are obtained from real data and thus correlated we cannot inspect the model by interpreting the coefficients of $\vec{W}$, see e.g.
-\cite{Haufe2013}.
-We thus only use univariate correlation to investigate how discriminative features are.
+\subsection{Analysis of bag-of-words features}\label{sec:correlations_methods}
+While interpretability of linear models is often propagated as one of their main advantages, doing so naively without modelling the noise covariances can lead to wrong conclusions, see e.g.
+\cite{Zien2009, Haufe2013}; interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
+Thus direct interpretation of the model coefficients $\vec{W}$ is problematic.
+In order to allow for better interpretation of the predictions and to assess which features contribute to the classification correlation coefficients between each word and the party affiliation label were computed.
 The words corresponding to the top positive and negative correlations are shown in section \autoref{sec:word_party_correlations}.
@@
 
-The following sections provide an overview of the results for both types of political bias prediction tasks, party affiliation and political view.
-Some interpretations of discriminative features are given one section will highlight a finding on the correlation of political power and positive sentiment of political texts.
+The following sections give an overview of the results for all political bias prediction tasks.
+Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section.
 
 \subsection{Predicting political party affiliation}
-The results for the political party affiliation prediction on held-out parliament data and on evaluation data are shown in \autoref{tab:results_17} for the 17th Bundestag and in \autoref{tab:results_18} for the 18th Bundestag, respectively.
+The results for the political party affiliation prediction on held-out parliament data and on evaluation data are listed in \autoref{tab:results_17} for the 17th Bundestag and in \autoref{tab:results_18} for the 18th Bundestag, respectively.
 Shown are the evaluation results for in-domain data (held-out parliament speech texts) as well as the out-of-domain data results obtained from testing the models on the texts extracted from party manifestos.
-When looking at predicting party affiliation from text data from the same domain that was used for training precision and recall values of above 0.6 are obtained.
+When predicting party affiliation on text data from the same domain that was used for training the model average precision and recall values of above 0.6 are obtained.
 \\
@@
 In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, we aggregated the manifesto data based on the topic.
-The manifesto codes contain labels for different topics, which were used to concatenate the data points of each party to only eight data points per party, each data point corresponding to one of the eight topics defined and annotated by the manifesto project annotators.
-The results are shown in table \autoref{tab:results_17_topic} and \autoref{tab:results_18_topic} and clearly show that when the texts to be classified and the word count statistics are sufficiently dense the classification performance on out of domain data becomes close to precisions of 1.0.
+The manifesto code political topics labels were used to concatenate texts of each party to one of eight topics, see \autoref{sec:data}.
+The results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve reliably precision and recall values close to 1.0.
 
@@
     \hline
+        \hline
 cducsu     &  0.64  &    1.00  &    0.78    &     7\\
@@
        spd   &    0.80   &   0.50    &  0.62     &    8\\
-    \hline
     \hline
@@
     \hline
+    \hline
   cducsu     &  0.50   &   1.00  &    0.67     &    8\\
@@
         spd     &  0.56  &    0.62  &    0.59   &      8\\
-
+\hline
 avg / total   &    0.51 &     0.62   &   0.55   &     32\\
@@
 {\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
-Instead of sentence level predictions as in \autoref{fig:results_17}, here predictions per political topic are shown (see \autoref{sec:data}); in contrast to single sentence level predictions (see \autoref{tab:confusion}) the predictions made on topic level are in many cases reliable.
-One effect present in the single sentence level predicitions however is even more pronounced in the topic level predictions: the policy of the green party manifesto is completely absorbed in the parliament speeches of the right-conservative CDU/CSU party -- all manifesto topics of the green party are predicted to be from one of the parties of the governing coalition, CDU/CSU or SPD.}
+In contrast to single sentence level predictions (see \autoref{tab:results_17}, \autoref{tab:results_18}, \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.
+Note that all manifesto topics of the green party in the 18th Bundestag are predicted to be from the parties of the governing coalition, CDU/CSU or SPD.}
 \end{table}
 
-%trained with cv on bundestag speeches 17
-%tested on topic aggregated (first digit of manifestocode) manifesto data
-%
-%cducsu       0.64      1.00      0.78         7
-%        fdp       1.00      1.00      1.00         7
-%     gruene       1.00      0.86      0.92         7
-%      linke       1.00      1.00      1.00         7
-%        spd       0.80      0.50      0.62         8
-%
-%avg / total       0.88      0.86      0.86        36
-%
-%Confusion Matrix (rows=true, cols=predicted)
-%cducsu, fdp, gruene, linke, spd
-%[7, 0, 0, 0, 0]
-%[0, 7, 0, 0, 0]
-%[0, 0, 6, 0, 1]
-%[0, 0, 0, 7, 0]
-%[4, 0, 0, 0, 4]
-%
-%0.86111111111111116
-%
-%trained with cv on bundestag speeches 18
-%tested on topic aggregated (first digit of manifestocode) manifesto data
-%
-%*** Training Set (CVd) ***
-%             precision    recall  f1-score   support
-%
-%     cducsu       0.50      1.00      0.67         8
-%     gruene       0.00      0.00      0.00         8
-%      linke       1.00      0.88      0.93         8
-%        spd       0.56      0.62      0.59         8
-%
-%avg / total       0.51      0.62      0.55        32
-%
-%Confusion Matrix (rows=true, cols=predicted)
-%cducsu, fdp, gruene, linke, spd
-%[8, 0, 0, 0]
-%[4, 0, 0, 4]
-%[1, 0, 7, 0]
-%[3, 0, 0, 5]
-%
-
-In order to investigate the errors the models made we extracted confusion matrices for the predictions on the out-of-domain evaluation data, the manifesto data.
-So the model, trained on party speeches in the parliament was tested on the texts parties published before the elections in order to get elected.
-One example in \autoref{tab:confusion} shows that these mistakes the model makes can give valuable insights into the changes in policy a party decides to make.
-The green party was promoting policies for renewable energy and agains nuclear energy in their manifestos prior to both legislative periods.
-When training a classifier on what representatives in the parliament say however, the statements of the green party are more often predicted to be from the conservative government than from the party that originally promoted these green ideas.
-This effect is also reflected in the precision values on out-of-domain data: The green party has the highest precision values by a large margin in both legislation periods, indicating that there is a relatively high true-positive rate.
-A similar effect was can be observed with the social democratic party (SPD): Evaluating the parliament classifier on manifestos it seems that a significant amount of speeches get classified as conservative government while being from the SPD.
+In order to investigate the errors the models made confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
+One example illustrates that the mistakes the model makes can be associated with changes in the party policy.
+The green party has been promoting policies for renewable energy and against nuclear energy in their manifestos prior to both legislative periods.
+Yet the statements of the green party are more often predicted to be from the government parties than from the party that originally promoted these green ideas, reflecting the trend that these legislative periods governing parties took over policies from the green party.
+This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag all manifesto topics of the green party are predicted to be from one of the parties of the governing coalition, CDU/CSU or SPD.
 \\
 
-Next to the party affiliation we also predicted whether or not a text is from a party that belonged to a given governing coalition of the Bundestag.
+\paragraph{Government membership prediction} Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag.
 In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
 While the in-domain evaluation precision and recall values are much better in this case, reaching values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
+While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved.
 
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
-\begin{tabular}{lcc|ccccc}
+\begin{tabular}{lccccccc}
  \multicolumn{8}{c}{\bf 17th Bundestag}\\
@@
 \hfill
-\begin{tabular}{lcc|cccc}
+\begin{tabular}{lcccccc}
  \multicolumn{7}{c}{\bf 18th Bundestag}\\
@@
 \caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
-In particular the green party suffers from this effect: when trained on the speeches in the 18th Bundestag, 2092 statements out of the green party's manifesto were classified as belonging to the cducsu party.
 This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
@@
 \vspace{1em}
-\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); for interpretations see also \autoref{tab:results_topic}.}
+\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right).}
 \end{table}
@@
 Parties change their policies and positions in the political spectrum.
-More reliable categories for political bias are party independent labels for political standpoints.
-These are provided by social scientists in the manifesto project \cite{manifesto}.
-A separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:models} on the manifesto code labels.
-As we do not have an out-of-domain evaluation set for this type of experiment we only report evaluation error.
+More reliable categories for political bias are party independent labels for political views, see \autoref{sec:data}.
+A separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:model}.
+As there was no out-of-domain evaluation set available in this setting only evaluation error on in-domain data is reported.
 Note however that also in this experiment the evaluation data was never seen by any model during training time.
-Despite the rather low values, a precision of close to 0.5 is rather high considering the large number of labels and the difficulty of the problem. \\
+In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes. 
+Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels. \\
 
@@
             0    & undefined &  0.09  &    0.12   &   0.10    &    17\\
-avg / total    &  0.47    &  0.46 &     0.46 &     2946\\
+avg / total  &  &  0.47    &  0.46 &     0.46 &     2946\\
 %
@@
 
-\subsection{Speech sentiment correlates with political power}\label{sec:sentiment_result}
-In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
-The average sentiment of each political party is shown in \autoref{fig:party_sentiments} for the two legislative periods under investigation.
-While the methodology used for estimating the sentiment (see \autoref{sec:sentiment_analysis_methods}) is simple and limited to a one dimensional spectrum of sentiment, it can yield some insight into the emotional content of a speech.
-High values indicate more pronounced usage of positive words, whereas negative values indicate more pronounced usage of words associated with negative emotional content.
-
-The results show an interesting relationship between political power and sentiment.
-Political power was evaluated in two ways: a) in terms of the number of seats a party has and b) in terms of membership of the government.
-Correlating either of these two proxies of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
-This pattern is evident from the barplot in \autoref{fig:party_sentiments} and in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
-
-Note that there is one party, the social democrats (SPD), which have a large number of seats and were in the opposition in the 17th Bundestag and member of the government in the 18th Bundestag: The average sentiment of this party switched sign from negative to positive when they joined the government, indicating that positive sentiment is a strong indicator of government membership of affiliation when analysing a text.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf} 
-%
-\end{center}
-\caption{
-\label{fig:party_sentiments}
-Speech sentiments computed for speeches of each party; parties are ordered according to political power measured in number of seats in the parliament from left to right.
-The more political power a party has, the more positive the speech content is.
-Note that while the SPD (red) was in the opposition in the 17th Bundestag it became part of the government in the 18th Bundestag: their seats in the parliament increased and the average sentiment of their speeches switched sign from negative to overall positive sentiment.
-}
-\end{figure}
-
-\begin{table}[t]
-\begin{center}
-\begin{tabular}{lcc}
-   Sentiment vs. &          Gov. Member    &  Seats\\
-\hline\hline
-17th Bundestag    &  0.84 & 0.70\\
-18th Bundestag   &  0.98 & 0.89\\
-%
-\end{tabular}
-\end{center}
-\caption{
-\label{tab:sentiments}
-Correlation coefficient between average sentiment of political speeches of a party in the german bundestag with two measures of their political power, a) their membership in the government and b) with the number of seats that party occupies in the parliament.
-Shown are correlations for the 17th legislative period and the (ongoing) 18th legislative period.
-}
-\end{table}
-
 \subsection{Correlations between words and parties}\label{sec:word_party_correlations}
 
-While interpretability of linear models is often propagated as one of their main advantages, doing so naively without modelling the noise covariances can lead to wrong conclusions, see e.g.
-\cite{Zien2009, Haufe2013} -- essentially interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; as this assumption is violated by our data simple univariate correlation measures are used to interpret the features used in the linear predictors, see also \autoref{sec:correlations_methods}.
-The 10 highest and lowest correlations between individual words and each party are shown in \autoref{fig:party_word_correlations}.
-While the correlations do indicate that there are -- despite extensive and careful cleaning procedures -- still some unspecific words in the top words, we conjecture that at least some of these words can be useful for a better interpretation of the model and the results, even if not every word can be easily associated with political views.
-
-In the following we give some examples of words that appear to be preferentially used or avoided by each respective party.
-Even though interpretations of these quantitative results are somewhat problematic, as they neglect the context in which these words were mentioned, we find some interesting patterns.
+The 10 highest and lowest correlations between individual words and the party affiliation label are shown for each party in \autoref{fig:party_word_correlations}.
+Correlations were computed on the data from the current, 18th, legislative period.
+Some unspecific stopwords are excluded.
+The following paragraphs highlight some examples of words that appear to be preferentially used or avoided by each respective party.
+Even though interpretations of these results are problematic in that they neglect the context in which these words were mentioned some interesting patterns can be found and related to the actual policies the parties are promoting.
 
@@
 The conservative christian party often uses words related to a pro-economy attitute, such as competitiveness ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
-The latter refers to the ongoing debates about whether or not the governments should be allowed to collect data and thus restrict fundamental civil rights in order to better secure the population from terrorist attacks.
-However the party rarely uses the word war ({\em krieg}).
+The latter could be related to the ongoing debates about whether or not the governments should be allowed to collect data and thus restrict fundamental civil rights in order to better secure the population.
+In contrast to the parties of the opposition, the conservatives rarely mention the word war ({\em krieg}) or related words.
 
@@
 
+\subsection{Speech sentiment correlates with political power}\label{sec:sentiment_result}
+In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
+The average sentiment of each political party is shown in \autoref{fig:party_sentiments}.
+High values indicate more pronounced usage of positive words, whereas negative values indicate more pronounced usage of words associated with negative emotional content.
+
+The results show an interesting relationship between political power and sentiment.
+Political power was evaluated in two ways: a) in terms of the number of seats a party has and b) in terms of membership of the government.
+Correlating either of these two indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
+This pattern is evident from the data in \autoref{fig:party_sentiments} and in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+
+Note that there is one party, the social democrats (SPD), which has many seats and switched from opposition to government with the 18th Bundestag: With its participation in the government the average sentiment of this party switched sign from negative to positive, suggesting that positive sentiment is a strong indicator of government membership.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf} 
+%
+\end{center}
+\caption{
+\label{fig:party_sentiments}
+Speech sentiments computed for speeches of each party; parties are ordered according to the number of seats in the parliament.
+There is a trend for more positive speech content with more political power.
+Note that the SPD (red) switched from opposition to government in the 18th Bundestag: their seats in the parliament increased and the average sentiment of their speeches switched sign from negative to overall positive sentiment.
+}
+\end{figure}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lcc}
+   Sentiment vs. &          Gov. Member    &  Seats\\
+\hline\hline
+17th Bundestag    &  0.84 & 0.70\\
+18th Bundestag   &  0.98 & 0.89\\
+%
+\end{tabular}
+\end{center}
+\caption{
+\label{tab:sentiments}
+Correlation coefficient between average sentiment of political speeches of a party in the german Bundestag with two indicators of political power, a) membership in the government and b) the number of seats a party occupies in the parliament.
+}
+\end{table}
+
 \subsection{An example web application}
-To show an example use case of the above approach a web application was implemented that downloads regularly all articles from major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}}, applies some simple topic modelling to them and assigns articles to a topic and predicts the political view of an article using the manifesto code classifier trained on all party manifestos.
-Within each topic it is then possible to get an order (from left to right) overview of the opinions on that topic.
-The left right index is computed according to the outputs of the manifestocodes and deterministic left-right mapping defined by human experts \cite{leftright}.
+To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
+For each news article topic, headlines of articles are plotted along with the predictions of the political view of an article and two labels derived deterministically from the 56 class output, a left right index and the political domain of a text, see \cite{leftright}.
+Within each topic it is then possible to get an ordered (from left to right) overview of the articles on that topic.
 An example of one topic that emerged on March 31st is shown in \autoref{fig:fipi}.
-The demo is live and running\footnote{\url{http://to-be-included-in-camera-ready-version.com}} and the code is open sourced an available on git-hub\footnote{\url{http://to-be-included-in-camera-ready-version.com}}.
+A preliminary demo is live at \cite{fipidemo} and the code is available on github\cite{fipi}.
 \begin{figure}
@@
 \section{Conclusions, Limitations and Outlook}\label{sec:conclusion}
-The goal of this study was to present a simple approach for automating political bias prediction and to quantify how well this is possible.
+This study presents a simple approach for automated political bias prediction.
 The results of these experiments show that automated political bias prediction is possible with above chance accuracy.
-Moreover there are some interesting aspects of these analyses that could be interesting for social science researchers, too.
+When texts are sufficiently long even texts from different domains can be classified reliably.
+\\
 
-For one the misclassifications of a model can be related to the changes in policy of a party.
-This can be helpful to quantitatively back up such a change in policy with more than just anecdotal evidence from single texts.
-Second analysing the word-party correlations shows some discriminative words that are also used by the classifier to determine political bias.
-This insight can be related to the political views of a party and allows for validation of the models by human experts.
-Third when correlating the sentiment of a speech with proxies of political power there is a strong and significant positive correlation between political power and positive sentiment.
-This supports the (probably trivial) intuition that if we read a text that has a positive opinion on what the party in power does, the author of the text is probably affiliated with those in power.
-While such an insight of something rather obvious in itself might not be very useful, the technology behind it is: Sentiment analysis, and in particular such a simple approach as used here, can be easily automated and scaled up to massive amounts of text data.
-Having established such a way of relating this aspect of media content to the political bias allows to automate political bias direction using the proxy of sentiments, which are a rather domain-independent measure.
+Some aspects of these analyses could be interesting for social science researchers and three of these are highlighted in this study.
+First the misclassifications of a model can be related to the changes in policy of a party.
+A prominent example in the current legislative period of the german Bundestag is the green party; the misclassifications of the model suggest that many of the policies of the green party appear to have been absorbed by the governing parties (CDU/CSU and SPD).
+Such analyses could be helpful to quantitatively investigate a change in policy.
+Second analysing the word-party correlations shows that some discriminative words can be related to the political views of a party; this allows for validation of the models by human experts.
+Third when correlating the sentiment of a speech with proxies of political power there is a strong positive correlation between political power and positive sentiment.
+This supports the intuition that if we read a text that has a positive opinion on what the party in power does the author of the text is probably affiliated with those in power.
+While such an insight in itself might seem not very surprising this quantitatifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data.
+Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains.
+\\
 
@@
 Every training data set that is publicly available has an inherent bias as it is taken from a different domain.
-This study tried to quantify the impact of this, but not in all settings we had out-of-domain data to quantify generalization performance across text domains.
-For the cases where we had evaluation data from two domains there was a pronounced shift in performance between in-domain evaluation set and out-of-domain evaluation set.
-This was to be expected, given that the texts are written for a different audience and in completely different styles.
-This results in quite some non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
-Another important issue was that the semantic units in the case of the out-of-domain data were of much shorter length (often just parts of a sentence) and hence much noisier than the in-domain evaluation data.
-\\
+This study tried to quantify the impact of this, but not in all settings out-of-domain data was available to quantify generalisation performance across text domains.
+For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in-domain evaluation set and out-of-domain evaluation set.
+This was to be expected, given that the texts are written for a different audience and in different styles, resulting in non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
+However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance becomes better and in some cases reaches reliable predictions.\\
 
-Taken together the results show that when training on one domain and evaluating on an entirely different domain one can still obtain above chance performance, even though there is a significant drop in performance.
-The generalization performance across domains will be an important topic of future research.
-The aim of this study was mainly to assess the impact of this effect, but if the goal is to minimize the generalization performance the simplest solution, next to more complicated approaches from semisupervised learning, non-stationary methods or covariate shift approaches, is to simply include data from different -- possibly very heterogeneous -- domains in the training data set.
-A popular approach is to integrate data from social networks such as Facebook, as advocated for instance in \cite{Arzheimer2016}.
+The generalisation performance across domains will be an important topic of future research.
+Next to more complicated approaches tailored to non-stationary data exhibiting covariate shifts, a simple direction of research could be to include data from different domains in the training data set and thus minimising the difference in data distributions between training and test.
+A popular approach is to integrate data from social networks such as Facebook posts, as advocated for instance in \cite{Arzheimer2016}.
 \\
@@
 This study would not have been possible without the open data initiatives behind the code and projects that led to the publicly available data sets and APIs described in \autoref{sec:data}.
+Some backend configurations for the web application were taken from an earlier collaboration with Daniel Kirsch.
+Pola Lehmann and Michael Gaebler provided helpful feedback on an earlier version of the manuscript.
 %

### 2016-04-04 manuscript/political_bias_prediction.tex
Final draft.

@@
 Felix Biessmann
-\thanks{}
+\thanks{felix.biessmann@gmail.com}
 %
@@
 Every day media generate large amounts of text.
-An unbiased view on what media report on requires an understanding of the political bias of media content.
+An unbiased view on media reports requires an understanding of the political bias of media content.
 Assistive technology for estimating the political bias of texts can be helpful in this context.
@@
 But ordering media content with respect to the political spectrum at scale requires automated prediction of political bias.
-And determining a political spectrum that goes beyond the two dimensional left-right scheme is difficult in itself.
-The aim of this study is to provide some empirical evidence indicating that leveraging some valuable open data sources political bias prediction is possible with above chance accuracy.
+The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
 \\
@@
 This is a key difference to many statistical learning approaches.
-Not only is the responsibility question not entirely clear, it can also be difficult to interpret some of the decisions.
+Not only is the responsibility question problematic, it can also be difficult to interpret some of the decisions.
 In order to validate the predictions of the models three strategies that allow for better interpretations of the models are proposed.
@@
 
-In the following \autoref{sec:data} gives an overview of data acquisition and preprocessing, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
 
@@
 %
-We used publicly available data sets of german political texts and standard libraries for processing the text. The following sections describe the details of data acquisition and feature extraction.
+All experiments were run on publicly available data sets of german political texts and standard libraries for processing the text.
+The following sections describe the details of data acquisition and feature extraction.
 
 \subsection{Data}
-We obtained annotated political text data from sources a) the discussions and speeches held in the german parliament ({\em Bundestag}) and b) all manifesto texts from parties running for election in the german parliament in the current 18th and the last, 17th, legislation period.
+Annotated political text data was obtained from two sources: a) the discussions and speeches held in the german parliament ({\em Bundestag}) and b) all manifesto texts of parties running for election in the german parliament in the current 18th and the last, 17th, legislation period.
 
 \paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
-The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}, we used an open source API to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
-In total we obtained 22784 speeches for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.
+The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}; an open source API was used to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
+In total 22784 speeches were extracted for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.
 
 \paragraph{Party manifesto data}
-For party manifestos we also use an openly accessible API provided by the Wissenschaftszentrum Berlin (WZB).
+For party manifestos another openly accessible API was used, provided by the Wissenschaftszentrum Berlin (WZB).
 The API is released as part of the {\em Manifestoproject} \cite{manifesto}.
@@
 Each sentence (in some cases also parts of sentences) is annotated with one of 56 political labels.
-Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare, ...}, for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
+Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
 The set of labels was developed by political scientists at the WZB and released for public use.
-We obtained all manifestos of parties that were running for election in this and the last legislative period.
-In total this resulted in 29451 political statements that had two types of labels: First the party affiliation of each political statement which was used to evaluate the party evaluation classifiers trained on the parliament speeches.
-For this purpose we constrained the data to only those parties that were elected into the parliament.
+All manifestos of parties that were running for election in this and the last legislative period were obtained.
+In total this resulted in 29451 political statements that had two types of labels: First the party affiliation of each political statement; this label was used to evaluate the party evaluation classifiers trained on the parliament speeches.
+For this purpose the data acquisition was constrained to only those parties that were elected into the parliament.
 Next to the party affiliation the political view labels were extracted.
-For the analyses based on political view labels we considered all parties, also those that did not make it into the parliament.
+For the analyses based on political view labels all parties were considered, also those that did not make it into the parliament.
 
 The length of each annotated statement in the party manifestos was rather short.
-The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters, measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.
+The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters.
+Measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.
 This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos.
 However for a classifier (as well as for humans) such short sentences can be rather difficult to classify.
-In order to obtain less 'noisy' data points from each party all statements were aggregated into political topics using the manifesto code labels.
+In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into political topics using the manifesto code labels.
 Each political view label is a three digit code, the first digit represents the political domain.
@@
 The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of the dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
-We tried several options for vectorizing the speeches, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
+Several options for vectorizing the speeches were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
 All of these hyperparameters were subjected to hyperparameter optimization as explained in \autoref{sec:crossvalidation}.
@@
 Where $\|~\|_F$ denotes the Frobenius Norm and $\gamma$ is a regularization parameter controlling the complexity of the model. 
- We optimized the regularization parameter on a log-scaled grid from $10^{-4,\dots,4}$ and the regularization constant was adopted to reflect asymmetric class frequency distributions.
+The regularization parameter was optimized on a log-scaled grid from $10^{-4,\dots,4}$.
 The performance of the model was optimized using the classification accuracy, but we also report all other standard measures, precision ($TP / (FP + TP$), recall ($TP / (TP + FN)$) and f1-score ($2\times (Prec.
 \times Rec) / (Prec + Rec.)$).
+\\
 
@@
 Political view classification is based on the labels of the manifesto project, see \autoref{sec:data} and \cite{leftright}.
-For each of first two problems, party affiliation and government membership prediction, we trained classifiers on the parliament speeches.
-For the third problem we trained only on the manifesto data for which political view labels were available.
+For each of first two problems, party affiliation and government membership prediction, classifiers were trained on the parliament speeches.
+For the third problem classifiers were trained only on the manifesto data for which political view labels were available.
 
 \subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
-The model pipeline contained a number of {\em hyperparameters} that were optimised using cross-validation.  
+The model pipeline contained a number of  hyperparameters that were optimised using cross-validation.  
 We first split the training data into a training data set that was used for optimisation of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data; wherever possible the generalisation performance of the models was also evaluated on out-of domain data.
@@
 \subsection{Sentiment analysis}\label{sec:sentiment_analysis_methods}
-A publicly available key word list was used to extract sentiments from the text subjected to classification \cite{remquahey2010}.
-The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and sentiment vector $\vec{s}\in\R^d$ which constructed from the values in the sentiment dictionary.
+A publicly available key word list was used to extract sentiments \cite{remquahey2010}.
+A sentiment vector $\vec{s}\in\R^d$ was constructed from the values in the sentiment dictionary.
+The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and $\vec{s}$
+
+\begin{align}
+\frac{\vec{s}^\top \vec{x}}{\|\vec{s}\|\|\vec{x}\|}
+\end{align}
 
@@
 Thus direct interpretation of the model coefficients $\vec{W}$ is problematic.
-In order to allow for better interpretation of the predictions and to assess which features contribute to the classification correlation coefficients between each word and the party affiliation label were computed.
+In order to allow for better interpretation of the predictions and to assess which features are discriminative correlation coefficients between each word and the party affiliation label were computed.
 The words corresponding to the top positive and negative correlations are shown in section \autoref{sec:word_party_correlations}.
@@
 The results for the political party affiliation prediction on held-out parliament data and on evaluation data are listed in \autoref{tab:results_17} for the 17th Bundestag and in \autoref{tab:results_18} for the 18th Bundestag, respectively.
-Shown are the evaluation results for in-domain data (held-out parliament speech texts) as well as the out-of-domain data results obtained from testing the models on the texts extracted from party manifestos.
-When predicting party affiliation on text data from the same domain that was used for training the model average precision and recall values of above 0.6 are obtained.
-\\
+Shown are the evaluation results for in-domain data (held-out parliament speech texts) as well as the out-of-domain data; the party manifesto out-of-domain predictions were made on the sentence level.
 
+When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained. 
 For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4.  
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, we aggregated the manifesto data based on the topic.
+In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated based on the topic.
 The manifesto code political topics labels were used to concatenate texts of each party to one of eight topics, see \autoref{sec:data}.
-The results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve reliably precision and recall values close to 1.0.
+The topic level results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve reliably precision and recall values close to 1.0.
 
@@
 Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively.
+Predictions on the manifesto data was done on {\bf sentence level}.
 }
@@
 Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively.
+Predictions on the manifesto data was done on {\bf sentence level}.
 }
@@
 Yet the statements of the green party are more often predicted to be from the government parties than from the party that originally promoted these green ideas, reflecting the trend that these legislative periods governing parties took over policies from the green party.
-This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag all manifesto topics of the green party are predicted to be from one of the parties of the governing coalition, CDU/CSU or SPD.
+This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag predicts all manifesto topics of the green party to be from one of the parties of the governing coalition, CDU/CSU or SPD.
 \\
@@
 In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
-While the in-domain evaluation precision and recall values are much better in this case, reaching values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
-While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved.
+While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
+While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
 
@@
 \vspace{1em}
-\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
 This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
@@
             0    & undefined &  0.09  &    0.12   &   0.10    &    17\\
+            \hline
 avg / total  &  &  0.47    &  0.46 &     0.46 &     2946\\
@@
 \label{tab:results_avg_political_view}
-Classification performance on the 56 class prediction problem of political views, see \autoref{sec:data}.
+Classification performance of 56 political views, see \autoref{sec:data}.
 }
@@
 
-\paragraph{\bf Green party (Gruene)}
+\paragraph{\bf Green party (gruene)}
 The green party heavily criticised the secret negotiations about the TiSA agreement\footnote{\url{https://en.wikipedia.org/wiki/Trade_in_Services_Agreement}} and insists in formal inquiries that the representatives of the green party put forward in this matter ({\em fragen, anfragen}).
-They also often put forward critical questions related to army projects ({\em R\"ustungsprojekte, Wehrbericht}) or the military development in east europe ({\em  Jalta}\footnote{Referring to the \url{https://en.wikipedia.org/wiki/Yalta_Conference}}).
+They also often ask questions related to army projects ({\em R\"ustungsprojekte, Wehrbericht}) or the military development in east europe ({\em  Jalta}\footnote{Referring to the \url{https://en.wikipedia.org/wiki/Yalta_Conference}}).
 
@@
 \paragraph{\bf Conservative party (CDU/CSU)}
-The conservative christian party often uses words related to a pro-economy attitute, such as competitiveness ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
+The conservative christian party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
 The latter could be related to the ongoing debates about whether or not the governments should be allowed to collect data and thus restrict fundamental civil rights in order to better secure the population.
@@
 One of the main limitations of an automated political bias prediction system is the availability of training data.
-Every training data set that is publicly available has an inherent bias as it is taken from a different domain.
+Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain.
 This study tried to quantify the impact of this, but not in all settings out-of-domain data was available to quantify generalisation performance across text domains.
@@
 This was to be expected, given that the texts are written for a different audience and in different styles, resulting in non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
-However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance becomes better and in some cases reaches reliable predictions.\\
+However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better and in some cases reliable predictions can be obtained even beyond the training text domain.\\
 
@@
 Next to more complicated approaches tailored to non-stationary data exhibiting covariate shifts, a simple direction of research could be to include data from different domains in the training data set and thus minimising the difference in data distributions between training and test.
-A popular approach is to integrate data from social networks such as Facebook posts, as advocated for instance in \cite{Arzheimer2016}.
+A popular approach is to integrate data from social networks such as Facebook posts see for instance \cite{Arzheimer2016}.
 \\
@@
 All data sets used in this study were publicly available, all code for experiments and the live web application can be found online \cite{fipi, fipidemo}.
-%-- in fact even the very definition of a spectrum could be considered a difficult problem. The aim of this study is to investigate
 
@@
 This study would not have been possible without the open data initiatives behind the code and projects that led to the publicly available data sets and APIs described in \autoref{sec:data}.
+I would like to thank Friedrich Lindenberg for factoring out the \url{https://github.com/bundestag/plpr-scraper} from his bundestag project.
 Some backend configurations for the web application were taken from an earlier collaboration with Daniel Kirsch.
 Pola Lehmann and Michael Gaebler provided helpful feedback on an earlier version of the manuscript.
+Pola Lehman also helped with getting access to and documentation on the Manifestoproject data.
 %

### 2016-05-22 manuscript/rebuttal.txt
revision with reviewers comments

I would like to thank all three reviewers for their careful and helpful feedback.
Especially I would like to thank reviewer 2 for providing valuable references of related work. I found these both highly relevant and interesting and I do wish I would have found them before. I read and added those and others I found meanwhile to the bibliography. Also I added a discussion of the related work to the revised manuscript and pointed out common results as well as new aspects in my results that I couldn’t find in those references. I do hope that the reviewers would consider taking this into account.

I would like to address first the three main concerns of the reviewers and then answer some specific concerns in more detail.

1) Missing related work:
I did extensive literature research, compared and discussed as many studies I could cover and added them to the revised manuscript.

2) The analysis only covers german.
That is partly due to the availability of data, but also because I didn't feel comfortable with other languages - but I will definitely look into more languages asap.

3) Low and badly explained accuracy of predictions:
Thanks to the references provided by reviewer 2, I was able to compare the results with previous studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
I found my results to be at least on par with or better than those reported in previous studies (on comparable data sources). Importantly also the explanations and empirical tests thereof provided in my submitted manuscript were in line and in parts extended the findings in those studies. The relevant parts in the manuscript now reflect these points.

a) the results of binary party classification (in the Canadian parliament, see Yu et al, 2008) was comparable (accuracies just above 0.8) to the accuracy in the present study (0.85 for the 17th Bundestag)
b) the results of 0.63 accuracy on the five class classification on the 17th Bundestag is actually slightly better than the only comparable study I found: in the European parliament, Hirst et al, 2014 obtained a classification accuracy of 0.61 on a five class problem of political affiliation prediction.
c) The main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts, as demonstrated in the submission. Both effects were described before in Yu et al, 2008 and Hirst et al 2014, respectively. Also the results on the correlation between political power and speech sentiment complement the findings of Hirst et al on ideology-unspecific language in Canadian parliament debates.

DETAILLED COMMENTS

REVIEWER 2
Q: [...] other classifiers
A: I did try others (random forests, kernel based methods), but they did not improve the results.

Q: What about stemming of words?
A: I did include this in the hyperparameter optimization, but it did not improve results, so i excluded it for brevity.

Q: [sentence level too fine, topic too coarse ]
A: I tried paragraphs for manifestoproject data earlier, results were between sentence and topic performance. Hirst et al, 2014 show a great analysis of text length and accuracy.

REVIEWER 3

Q: each sentence [...] associated to many labels?
A: No.

Q: [...] 3-fold cross-validation sufficent [...]
A: [...] One split into train/hyperparameter-optimization/evaluation is common practice (see e.g. Netflix prize); i did 5 fold nested CV, results were the same.

Q: d is the size of the dictionary [...] union of the words [...] in the corpus and the words [...] in the sentiment dictionary
A: Yes

Q: the sentiment vector s includes all sentiment-related words, both positive and negative.
A: It contains the sentiment polarity values associated with the key words. The manuscript now reads "was constructed from the sentiment polarity values in the sentiment dictionary." (before: "was constructed from the sentiment values in the sentiment dictionary.")

Q: What is N in Table 1?
A: Number of data points, the revised manuscript includes this now.

Q: [remark on poor out-of-domain generalization results Table 3]
A: These effects are in line with previous findings (now referenced and discussed) and I tried to explain some of the reasons with experiments the results of which are also inline with previous findings. It is just a very challenging problem and there is not a lot of data at our disposal: party policies compared to those in their manifestos _have_ to change as part of the coalition negotiations and democratic processes in the parliaments.  But the manuscript shows how some of these prediction failures can be related to the policy changes.

Q: [...] difference in performance between the 17th and 18th bundestag
A: one is a 5 class the other a 4 class problem, also the 18th Bundestag is still running, so the amount of data is different. The revised manuscript reflects these differences better.

Q: [remark on topic level results in Table 3 and Table 5]
A: These experiments were mostly meant to show the effect of text length, but the results nonetheless show that in some (not all) cases highly accurate predictions are possible.

Q: [...] performance [...] of political views is not [...] high. [...] based on what this may be considered relevant and/or useful
A: The performance is significantly above chance for some (not all) classes (chance level being 1/56=0.018 assuming balanced classes). Even if the results are not perfect the model can be still highly relevant for assistive technology e.g. for speeding up human annotations in active learning scenarios.

Q: [...] experiment on correlations between words and parties raises
A: The experiments on word correlations and sentiment correlations are separate and independent experiments. They are also discussed independently.

Q: [remark on table 9]
A: The notion of sentiment used in the manuscript has polarity. The revised manuscript reflects this better.

### 2016-05-22 manuscript/political_bias_prediction.tex
revision with reviewers comments

@@
 Assistive technology can help in this context to try and obtain a more unbiased sample of information.
-\\
 
@@
 The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
-\\
 
@@
 Third sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
-\\
 
 In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+
+\section{Related Work}\label{sec:related}
+Throughout the last years automated content analyses for political texts content have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools.
+While a complete overview is beyond the scope of this work, the following paragraphs list similarities and differences between this study and previous work.
+For a more complete overview we refer the reader to \cite{Grimmer2013,Kaal2014}.
+
+A similar approach to the one presented here was taken in \cite{Yu2008}.
+The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches.
+They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one data source and tested on another.
+They found that generalization of the classifier as quantified by cross-validation was in some cases strongly affected by the domain.
+Generalization was also affected by the time difference between the political speeches used for training and those used for testing.
+%
+Other work has focused on developing dedicated methods for predicting political bias, what is often referred to as {\em scaling} in these studies; two have become particularly popular, WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}.
+Both methods have been applied in many studies and their strengths and weaknesses compared, which led to several improvements, see e.g.
+\cite{Lowe09scalingpolicy}.
+Validation is done by comparing parameter estimates of the different models amongst each other or with human expert annotations.
+Models are usually not evaluated by testing them on held-out data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult.
+In general these approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting.
+%
+Also texts from social network data were investigated in the context of automated political text analysis in the 2008 U.S.
+presidential election \cite{Gryc2014}.
+The study highlighted the importance of features such as sentiment or social network structure but the results also show that standard BOW features combined with linear classifiers yield competitive results for sentiment classification on political tweets.
+%
+Finally Hirst et al conducted a large number of experiments that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
+The authors show that a classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary.
+They also argue that emotional content plays an important role in automatic analysis of political texts.
+Furthermore their results show a clear depencency between length of a political text and the accuracy with which it can be classified correctly.
+
+Taken together, there is a large body of literature on an ever expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts.
+However except for few exceptions most previous work has focused to a large extent on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or to assignment of a one dimensional policy position (mostly left vs right).
+Often this is a too coarse approach to take into account subtle differences in political policies.
+This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies, and second the political view labels considered are more fine grained than in previous studies.
+Another important contribution of this work is that many existing studies are primarily concerned with a posteriori analysis of historical data.
+This work aims at prediction of political bias on out-of-domain data with a focus on the practical application on new data, for which a prototypical web application is provided.
+The experiments follow the established principles of cross validation to estimate out-of-domain generalization and complement the work of \cite{Yu2008} and \cite{Hirst2014} by evaluating data from the german parliament.
+We confirm and extend some of their findings on why and when generalization performance is impacted negatively.
 
@@
 The general idea of bag-of-words vectors is to simply count occurrences of words (or word sequences, also called {\em n-grams}) for each data point.
-A data point is usually a documents, here it is the semantic units of parliament speeches and manifesto sentences, respectively.
+A data point is usually a document, here it is the semantic units of parliament speeches and manifesto sentences, respectively.
 The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of the dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
@@
 A publicly available key word list was used to extract sentiments \cite{remquahey2010}.
-A sentiment vector $\vec{s}\in\R^d$ was constructed from the values in the sentiment dictionary.
-The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between bag-of-word vectors $\vec{x}\in\R^d$ and $\vec{s}$
+A sentiment vector $\vec{s}\in\R^d$ was constructed from the sentiment polarity values in the sentiment dictionary.
+The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors $\vec{x}\in\R^d$ and $\vec{s}$
 
@@
 In order to allow for better interpretation of the predictions and to assess which features are discriminative correlation coefficients between each word and the party affiliation label were computed.
-The words corresponding to the top positive and negative correlations are shown in section \autoref{sec:word_party_correlations}.
+The words corresponding to the top positive and negative correlations are shown in \autoref{sec:word_party_correlations}.
 
@@
 When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
+These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem of prediction party affiliation in the European parliament; the accuracy for the 17th Bundestag is 0.63, results of the 18th Bundestag are difficult to compare as the number of parties is four and the legislation period is not finished yet.
 For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4.
+This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
@@
 The manifesto code political topics labels were used to concatenate texts of each party to one of eight topics, see \autoref{sec:data}.
-The topic level results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve reliably precision and recall values close to 1.0.
+The topic level results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve in the case of some parties reliably precision and recall values close to 1.0.
+This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
 
 \begin{table}[t]
+\caption{
+\label{tab:results_17}
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively.
+Predictions on the manifesto data was done on {\bf sentence level}; $N$ denotes number of data points in the evaluation set.
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_17}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively.
-Predictions on the manifesto data was done on {\bf sentence level}.
-}
 \end{table}
@@
 \begin{table}[t]
+\caption{
+\label{tab:results_18}
+Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively.
+Predictions on the manifesto data was done on {\bf sentence level}.
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_18}
-Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively.
-Predictions on the manifesto data was done on {\bf sentence level}.
-}
+
 \end{table}
@@
 \begin{table}[t]
+\caption{
+\label{tab:results_topic}
+{\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
+In contrast to single sentence level predictions (see \autoref{tab:results_17}, \autoref{tab:results_18}, \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.
+Note that all manifesto topics of the green party in the 18th Bundestag are predicted to be from the parties of the governing coalition, CDU/CSU or SPD.}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_topic}
-{\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
-In contrast to single sentence level predictions (see \autoref{tab:results_17}, \autoref{tab:results_18}, \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.
-Note that all manifesto topics of the green party in the 18th Bundestag are predicted to be from the parties of the governing coalition, CDU/CSU or SPD.}
+
 \end{table}
@@
 While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
-While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
+This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
+The authors report classification accuracy of up to 0.87, the accuracy in the 17th Bundestag was While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
 
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
+\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
 \begin{tabular}{lccccccc}
@@
 \vspace{1em}
-\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
-This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
 \end{table}
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
+\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right).}
 \begin{tabular}{lccccccc}
@@
 \end{tabular}
-\vspace{1em}
-\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right).}
+\vspace{1em}\\
 \end{table}
@@
 \begin{table}[t]
+\caption{
+\label{tab:results_binary_17}
+Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches into government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_binary_17}
-Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches into government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
-}
 \end{table}
@@
 \begin{table}[t]
+\caption{
+\label{tab:results_binary_18}
+Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_binary_18}
-Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
-}
 \end{table}
@@
 \begin{table}[t]
+\caption{
+\label{tab:results_avg_political_view}
+Classification performance of 56 political views, see \autoref{sec:data}.
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:results_avg_political_view}
-Classification performance of 56 political views, see \autoref{sec:data}.
-}
 \end{table}
@@
 Even though interpretations of these results are problematic in that they neglect the context in which these words were mentioned some interesting patterns can be found and related to the actual policies the parties are promoting.
-
 \paragraph{\bf Left party (linke)}
@@
 Rarely the party addresses concerns related to security ({\em sicherheit}).
-
 \paragraph{\bf Green party (gruene)}
@@
 They also often ask questions related to army projects ({\em R\"ustungsprojekte, Wehrbericht}) or the military development in east europe ({\em  Jalta}\footnote{Referring to the \url{https://en.wikipedia.org/wiki/Yalta_Conference}}).
-
 \paragraph{\bf Social democratic party (SPD)}
@@
 \begin{table}[t]
+\caption{
+\label{tab:sentiments}
+Correlation coefficient between average sentiment of political speeches of a party in the german Bundestag with two indicators of political power, a) membership in the government and b) the number of seats a party occupies in the parliament.
+}
 \begin{center}
@@
 \end{center}
-\caption{
-\label{tab:sentiments}
-Correlation coefficient between average sentiment of political speeches of a party in the german Bundestag with two indicators of political power, a) membership in the government and b) the number of seats a party occupies in the parliament.
-}
 \end{table}
@@
 This study presents a simple approach for automated political bias prediction.
-The results of these experiments show that automated political bias prediction is possible with above chance accuracy.
-When texts are sufficiently long even texts from different domains can be classified reliably.
-\\
+The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases.
 
@@
 First the misclassifications of a model can be related to the changes in policy of a party.
-A prominent example in the current legislative period of the german Bundestag is the green party; the misclassifications of the model suggest that many of the policies of the green party appear to have been absorbed by the governing parties (CDU/CSU and SPD).
 Such analyses could be helpful to quantitatively investigate a change in policy.
@@
 Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains.
-\\
 
@@
 For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in-domain evaluation set and out-of-domain evaluation set.
-This was to be expected, given that the texts are written for a different audience and in different styles, resulting in non-stationarities or covariate shifts between the in-domain and out-of-domain evaluation data.
-However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better and in some cases reliable predictions can be obtained even beyond the training text domain.\\
-
-The generalisation performance across domains will be an important topic of future research.
-Next to more complicated approaches tailored to non-stationary data exhibiting covariate shifts, a simple direction of research could be to include data from different domains in the training data set and thus minimising the difference in data distributions between training and test.
-A popular approach is to integrate data from social networks such as Facebook posts see for instance \cite{Arzheimer2016}.
-\\
+This effect was reported previously, see e.g.
+\cite{Yu2008}.
+Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}.
+However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better and in some cases reliable predictions can be obtained even beyond the training text domain.
 
@@
 \subsection*{Acknowledgements}
-This study would not have been possible without the open data initiatives behind the code and projects that led to the publicly available data sets and APIs described in \autoref{sec:data}.
 I would like to thank Friedrich Lindenberg for factoring out the \url{https://github.com/bundestag/plpr-scraper} from his bundestag project.

### 2016-05-22 bundestag.py

@@ -102,7 +102,7 @@ def get_raw_text_bundestag(folder="data/out", legislationPeriod=17):
 def run_experiments():
     for legis in [17,18]:
         #classify_speeches_binary_manifesto(legis);
-        #classify_speeches_binary_parliament(legis);
+        classify_speeches_binary_parliament(legis);
         #classify_speeches_party_manifesto(legis);
         classify_speeches_party_parliament(legis);
         #get_sentiments(legis)
@@ -292,6 +292,7 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     labelsStr = [str(x) for x in final_clf.steps[-1][1].classes_]
     report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(evalLabels, predictedEval).tolist(): report += str(line)+"\n" 
+    import pdb;pdb.set_trace()
     # dump report
     open(OUT_DIR+'/report-'+saveId+'.txt','wb').write(report)
     return report
@@ -377,8 +378,6 @@ def classify_speeches_party(legislationPeriod = 18):
             speech['speaker_party'] is not None and \
             speech['speaker_party'] in bundestagParties[legislationPeriod] and \
             len(speech['text']) > MINLEN:
-                import pdb
-                pdb.set_trace()
                 prediction = pclf.predict(speech['text'])
                 predictedParty.append(sp.argmax(prediction.values()))
                 trueParty.append(prediction.keys().index(speech['speaker_party']))
@@ -388,6 +387,8 @@ def classify_speeches_party(legislationPeriod = 18):
     for line in metrics.confusion_matrix(trueParty, predictedParty).tolist():
         report += str(line)+"\n" 
     open(OUT_DIR+'/report','wb').write(report)
+    import pdb;pdb.set_trace()
+    report += "Accuracy: %d" + accuracy_score(trueParty,predictedParty)
     return report
 
 def plot_leftright():

### 2016-05-22 manuscript/rebuttal.txt
now comments less than 5000 words

@@
 A: It contains the sentiment polarity values associated with the key words.
-The manuscript now reads "was constructed from the sentiment polarity values in the sentiment dictionary." (before: "was constructed from the sentiment values in the sentiment dictionary.")
 
@@
 
-Q: [remark on poor out-of-domain generalization results Table 3]
-A: These effects are in line with previous findings (now referenced and discussed) and I tried to explain some of the reasons with experiments the results of which are also inline with previous findings.
-It is just a very challenging problem and there is not a lot of data at our disposal: party policies compared to those in their manifestos _have_ to change as part of the coalition negotiations and democratic processes in the parliaments.
-But the manuscript shows how some of these prediction failures can be related to the policy changes.
-
 Q: [...] difference in performance between the 17th and 18th bundestag
-A: one is a 5 class the other a 4 class problem, also the 18th Bundestag is still running, so the amount of data is different. The revised manuscript reflects these differences better.
-
-Q: [remark on topic level results in Table 3 and Table 5]
-A: These experiments were mostly meant to show the effect of text length, but the results nonetheless show that in some (not all) cases highly accurate predictions are possible.
+A: one is a 5 class the other a 4 class problem, also the 18th Bundestag is still running, so the amount of data is different.
 
 Q: [...] performance [...] of political views is not [...] high. [...] based on what this may be considered relevant and/or useful
-A: The performance is significantly above chance for some (not all) classes (chance level being 1/56=0.018 assuming balanced classes).
-Even if the results are not perfect the model can be still highly relevant for assistive technology e.g.
-for speeding up human annotations in active learning scenarios.
+A: Even if the results are not perfect the model can be still highly relevant for assistive technology e.g. for speeding up human annotations in active learning scenarios.
 
 Q: [...] experiment on correlations between words and parties raises
-A: The experiments on word correlations and sentiment correlations are separate and independent experiments. They are also discussed independently.
+A: The experiments on word correlations and sentiment correlations are separate and independent experiments and discussed independently.
 
 Q: [remark on table 9]
-A: The notion of sentiment used in the manuscript has polarity. The revised manuscript reflects this better.
+A: The notion of sentiment used in the manuscript has polarity.

### 2016-05-23 manuscript/rebuttal.txt
with pdf
shortened reply, polished ms text

@@
-I would like to thank all three reviewers for their careful and helpful feedback.
+I would like to thank all three reviewers for their careful review and helpful feedback.
 Especially I would like to thank reviewer 2 for providing valuable references of related work.
@@
 I read and added those and others I found meanwhile to the bibliography.
-Also I added a discussion of the related work to the revised manuscript and pointed out common results as well as new aspects in my results that I couldn’t find in those references.
-I do hope that the reviewers would consider taking this into account.
+Also I added a discussion of the related work to a revised manuscript and pointed out common results as well as new aspects in my results that I couldn’t find in those references.
+
+I do hope that the reviewers would consider taking this into account and I hope that it would be possible to add these added references and their discussion to the camera ready version of the manuscript?
 
@@
 2) The analysis only covers german.
-That is partly due to the availability of data, but also because I didn't feel comfortable with other languages - but I will definitely look into more languages asap.
+That is mostly due to the availability of data - but I will definitely look into more languages asap.
 
 3) Low and badly explained accuracy of predictions:
-Thanks to the references provided by reviewer 2, I was able to compare the results with previous studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
-I found my results to be at least on par with or better than those reported in previous studies (on comparable data sources).
-Importantly also the explanations and empirical tests thereof provided in my submitted manuscript were in line and in parts extended the findings in those studies.
-The relevant parts in the manuscript now reflect these points.
+I compared the results with previous studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
+My results are on par with or better than those reported in previous studies (on comparable data sources).
+Importantly also the explanations and empirical tests thereof provided in my submitted manuscript were in line with and complemented the findings in those studies:
 
-a) the results of binary party classification (in the Canadian parliament, see Yu et al, 2008) was comparable (accuracies just above 0.8) to the accuracy in the present study (0.85 for the 17th Bundestag)
-b) the results of 0.63 accuracy on the five class classification on the 17th Bundestag is actually slightly better than the only comparable study I found: in the European parliament, Hirst et al, 2014 obtained a classification accuracy of 0.61 on a five class problem of political affiliation prediction.
-c) The main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts, as demonstrated in the submission.
-Both effects were described before in Yu et al, 2008 and Hirst et al 2014, respectively.
-Also the results on the correlation between political power and speech sentiment complement the findings of Hirst et al on ideology-unspecific language in Canadian parliament debates.
+a) the results of binary party classification (see Yu et al, 2008) was comparable (accuracies between 0.8 and 0.87) to the accuracy in the present study (0.85 for the 17th Bundestag)
+b) the accuracy in the five class classification of the 17th Bundestag (0.63) is slightly better than the only comparable study I found: Hirst et al, 2014 obtained a classification accuracy of 0.61 on a five class problem of political affiliation prediction in the European parliament.
+c) As demonstrated in the submission the main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts.
+Both effects were described before in Yu et al, 2008 and Hirst et al 2014, respectively, in similar (but not the same) settings.
 
-DETAILLED COMMENTS
+Also it would be worth noting that even if the accuracies are (as those of comparable studies) are low, they nonetheless are on average above chance.
+While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.
+
+DETAILED COMMENTS
 
@@
 Q: [...] other classifiers
-A: I did try others (random forests, kernel based methods), but they did not improve the results.
+A: I did try others (random forests, kernel based methods); results on in-domain data were slightly better but results on out of domain data were significantly worse (for all regularizations i tried).
 
 Q: What about stemming of words?
-A: I did include this in the hyperparameter optimization, but it did not improve results, so i excluded it for brevity.
+A: I did include this in earlier hyperparameter optimizations, but it did not improve results, so i excluded it for brevity.
 
 Q: [sentence level too fine, topic too coarse ]
-A: I tried paragraphs for manifestoproject data earlier, results were between sentence and topic performance. Hirst et al, 2014 show a great analysis of text length and accuracy.
+A: I tried paragraphs for manifestoproject data earlier, results were between sentence and topic performance. Hirst et al, 2014 show a comprehensive analysis of text length and accuracy, though.
 
@@
 Q: [...] 3-fold cross-validation sufficent [...]
-A: [...] One split into train/hyperparameter-optimization/evaluation is common practice (see e.g. Netflix prize); i did 5 fold nested CV, results were the same.
+A: [...] A single random split into train/hyperparameter-optimization/evaluation is common practice (see e.g. Netflix prize); I did 5 fold nested CV, results were the same.
 
@@
 
-Q: [...] performance [...] of political views is not [...] high. [...] based on what this may be considered relevant and/or useful
-A: Even if the results are not perfect the model can be still highly relevant for assistive technology e.g. for speeding up human annotations in active learning scenarios.
-
 Q: [...] experiment on correlations between words and parties raises
-A: The experiments on word correlations and sentiment correlations are separate and independent experiments and discussed independently.
+A: The experiments on word correlations and sentiment correlations are separate and independent experiments and are discussed independently.
 

### 2016-05-23 manuscript/political_bias_prediction.tex
shortened reply, polished ms text

@@
 
-In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+In the following \autoref{sec:related} briefly surveys some related work, thereafter \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
 
 \section{Related Work}\label{sec:related}
-Throughout the last years automated content analyses for political texts content have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools.
+Throughout the last years automated content analyses for political texts have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools.
 While a complete overview is beyond the scope of this work, the following paragraphs list similarities and differences between this study and previous work.
@@
 The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches.
-They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one data source and tested on another.
-They found that generalization of the classifier as quantified by cross-validation was in some cases strongly affected by the domain.
+They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one domain and tested on another.
+They found that accuracies of the model when trained on one domain and test on another was significantly decreased.
 Generalization was also affected by the time difference between the political speeches used for training and those used for testing.
-%
-Other work has focused on developing dedicated methods for predicting political bias, what is often referred to as {\em scaling} in these studies; two have become particularly popular, WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}.
-Both methods have been applied in many studies and their strengths and weaknesses compared, which led to several improvements, see e.g.
+
+Other work has focused on developing dedicated methods for predicting political bias such as for instance two methods known as WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}, respectively, and improved versions thereof, see e.g.
 \cite{Lowe09scalingpolicy}.
-Validation is done by comparing parameter estimates of the different models amongst each other or with human expert annotations.
-Models are usually not evaluated by testing them on held-out data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult.
-In general these approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting.
+Validation in these studies is often done by comparing parameter estimates of the different models amongst each other or with human expert annotations.
+These models are often not evaluated on held-out data or with respect to the same type of labels used to train or partition the data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult.
+These approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting.
 %
@@
 The study highlighted the importance of features such as sentiment or social network structure but the results also show that standard BOW features combined with linear classifiers yield competitive results for sentiment classification on political tweets.
-%
-Finally Hirst et al conducted a large number of experiments that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
+
+Finally Hirst et al conducted a large number of experiments in the Canadian parliament and the European parliament that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
 The authors show that a classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary.
@@
 
-Taken together, there is a large body of literature on an ever expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts.
+Taken together, there is a large body of literature in this expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts.
 However except for few exceptions most previous work has focused to a large extent on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or to assignment of a one dimensional policy position (mostly left vs right).
 Often this is a too coarse approach to take into account subtle differences in political policies.
-This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies, and second the political view labels considered are more fine grained than in previous studies.
+This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies; second the political view labels considered are more fine grained than in previous studies.
 Another important contribution of this work is that many existing studies are primarily concerned with a posteriori analysis of historical data.
-This work aims at prediction of political bias on out-of-domain data with a focus on the practical application on new data, for which a prototypical web application is provided.
-The experiments follow the established principles of cross validation to estimate out-of-domain generalization and complement the work of \cite{Yu2008} and \cite{Hirst2014} by evaluating data from the german parliament.
-We confirm and extend some of their findings on why and when generalization performance is impacted negatively.
+This work aims at prediction of political bias on out-of-domain data with a focus on the practical application of the model on new data, for which a prototypical web application is provided.
+The experiments estimate out-of-domain generalization and thus complement the work of \cite{Yu2008, Hirst2014} with results from data of the german parliament.
+In particular the results of this study confirm and extend some of their findings on why and when generalization performance is impacted negatively.
 
@@
 This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
-The authors report classification accuracy of up to 0.87, the accuracy in the 17th Bundestag was While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
+The authors report classification accuracies between 0.8 and 0.87, the accuracy in the 17th Bundestag was 0.85.
+While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
 
@@
 The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases.
+It is worth noting that even if the accuracies are (as those of comparable studies \cite{Yu2008, Hirst2014}) are not perfect, they nonetheless are on average above chance.
+While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.
 
-Some aspects of these analyses could be interesting for social science researchers and three of these are highlighted in this study.
+One of the main limiting factors of an automated political bias prediction system is the availability of training data.
+Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain.
+This study tried to quantify the impact of this, but not in all settings out of domain data was available to quantify generalisation performance across text domains.
+For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in domain evaluation set and out of domain evaluation set.
+This effect was reported previously for similar data, see e.g.
+\cite{Yu2008}.
+Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}.
+However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better; in some cases reliable predictions can be obtained even beyond the training text domain.
+
+Some aspects of these analyses could be interesting for social science researchers; three of these are highlighted here.
 First the misclassifications of a model can be related to the changes in policy of a party.
@@
 Third when correlating the sentiment of a speech with proxies of political power there is a strong positive correlation between political power and positive sentiment.
-This supports the intuition that if we read a text that has a positive opinion on what the party in power does the author of the text is probably affiliated with those in power.
-While such an insight in itself might seem not very surprising this quantitatifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data.
+While such an insight in itself might seem not very surprising this quantifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data.
 Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains.
-
-One of the main limitations of an automated political bias prediction system is the availability of training data.
-Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain.
-This study tried to quantify the impact of this, but not in all settings out-of-domain data was available to quantify generalisation performance across text domains.
-For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in-domain evaluation set and out-of-domain evaluation set.
-This effect was reported previously, see e.g.
-\cite{Yu2008}.
-Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}.
-However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better and in some cases reliable predictions can be obtained even beyond the training text domain.
+\\
 

### 2016-05-24 manuscript/rebuttal.txt
revision for rebuttal phase

@@
 
+An updated version can be found here
+
 I would like to address first the three main concerns of the reviewers and then answer some specific concerns in more detail.
@@
 I did extensive literature research, compared and discussed as many studies I could cover and added them to the revised manuscript.
+My results are well in line with previous findings but extend existing work with new analyses.
 
 2) The analysis only covers german.
-That is mostly due to the availability of data - but I will definitely look into more languages asap.
+That is mostly due to the availability of data - but I will look into more languages asap.
 
 3) Low and badly explained accuracy of predictions:
-I compared the results with previous studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
-My results are on par with or better than those reported in previous studies (on comparable data sources).
-Importantly also the explanations and empirical tests thereof provided in my submitted manuscript were in line with and complemented the findings in those studies:
+I compared the results with previous similar studies (Yu et al, "Classifying party affiliation from political speech", 2008 and Hirst et al, "Text to Ideology or Text to Party Status?", 2014).
+My results are on par with or better than the results reported in those studies on comparable data.
+Importantly also the explanations and empirical tests of these accuracies provided in my manuscript were in line with the findings in those studies:
 
 a) the results of binary party classification (see Yu et al, 2008) was comparable (accuracies between 0.8 and 0.87) to the accuracy in the present study (0.85 for the 17th Bundestag)
-b) the accuracy in the five class classification of the 17th Bundestag (0.63) is slightly better than the only comparable study I found: Hirst et al, 2014 obtained a classification accuracy of 0.61 on a five class problem of political affiliation prediction in the European parliament.
-c) As demonstrated in the submission the main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts.
+b) the accuracy in the five class classification of the 17th Bundestag (0.63) is slightly better than the classification accuracy of 0.61 which Hirst et al obtained on a five class problem of political affiliation prediction in the European parliament.
+c) As demonstrated in the my submission the main sources of poor accuracy are the out-of-domain generalization challenge and the short length of the texts.
 Both effects were described before in Yu et al, 2008 and Hirst et al 2014, respectively, in similar (but not the same) settings.
 
-Also it would be worth noting that even if the accuracies are (as those of comparable studies) are low, they nonetheless are on average above chance.
-While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.
+Also it would be worth noting that even if the accuracies are low (as those of comparable studies), they nonetheless are on average above chance.
+While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting for priorization of to-be-labeled data.
 
-DETAILED COMMENTS
+MORE DETAILED COMMENTS
 
@@
 Q: [...] other classifiers
-A: I did try others (random forests, kernel based methods); results on in-domain data were slightly better but results on out of domain data were significantly worse (for all regularizations i tried).
+A: I did try others (random forests, kernel based methods); results on in-domain data were slightly better but results on out of domain data were significantly worse (despite regularization).
 
@@
 Q: [sentence level too fine, topic too coarse ]
-A: I tried paragraphs for manifestoproject data earlier, results were between sentence and topic performance. Hirst et al, 2014 show a comprehensive analysis of text length and accuracy, though.
+A: I tried paragraphs for manifestoproject data earlier; results were between sentence and topic performance. Hirst et al, 2014 show a comprehensive analysis of text length and accuracy, though.
 
@@
 Q: What is N in Table 1?
-A: Number of data points, the revised manuscript includes this now.
+A: Number of data points, the revised manuscript includes this.
 

### 2016-05-24 manuscript/political_bias_prediction.tex
revision for rebuttal phase

@@
 But ordering media content with respect to the political spectrum at scale requires automated prediction of political bias.
-The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
+The aim of this study is to provide empirical evidence  indicating that leveraging open data sources of german texts, automated political bias prediction is possible with above chance accuracy.
+These experimental results  confirm and extend previous findings \cite{Yu2008, Hirst2014}; a novel contribution of this work is a proof of concept which applies this technology to sort news article recommendations according to their political bias.
 
@@
 Not only is the responsibility question problematic, it can also be difficult to interpret some of the decisions.
-In order to validate the predictions of the models three strategies that allow for better interpretations of the models are proposed.
-First interpretations of the model misclassifications are related to changes in party policies.
+In order to validate and explain the predictions of the models three strategies that allow for better interpretations of the models are proposed.
+First the model misclassifications are related to changes in party policies.
 Second univariate measures of correlation between text features and party affiliation allow to relate the predictions to the kind of information that political experts use for interpreting texts.
@@
 The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches.
-They used data from the two chambers of the US congress, House and Senat, in order to investigate generalization performance of a classifier trained on one domain and tested on another.
-They found that accuracies of the model when trained on one domain and test on another was significantly decreased.
+They used data from the two chambers of the US congress, House and Senat, in order to assess generalization performance of a classifier trained on data from one chamber and tested on data from another.
+They found that accuracies of the model when trained on one domain and tested on another were significantly decreased.
 Generalization was also affected by the time difference between the political speeches used for training and those used for testing.
 
-Other work has focused on developing dedicated methods for predicting political bias such as for instance two methods known as WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}, respectively, and improved versions thereof, see e.g.
+Other work has focused on developing dedicated methods for predicting political bias.
+Two popular methods are WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}, or improved versions thereof, see e.g.
 \cite{Lowe09scalingpolicy}.
-Validation in these studies is often done by comparing parameter estimates of the different models amongst each other or with human expert annotations.
-These models are often not evaluated on held-out data or with respect to the same type of labels used to train or partition the data, which renders comparisons with this work or other work such as \cite{Yu2008} difficult.
-These approaches have been very valuable for a posteriori analysis of historical data but do not seem to be used as much for analyses of new data in a predictive analytics setting.
-%
-Also texts from social network data were investigated in the context of automated political text analysis in the 2008 U.S.
-presidential election \cite{Gryc2014}.
-The study highlighted the importance of features such as sentiment or social network structure but the results also show that standard BOW features combined with linear classifiers yield competitive results for sentiment classification on political tweets.
+These approaches have been very valuable for {\em a posteriori} analysis of historical data but they do not seem to be used as much for analyses of new data in a predictive analytics setting.
+Moreover direct comparisons of the results obtained with these so called {\em scaling} methods with the results of the present study or those of studies as \cite{Yu2008} are difficult, due to the different modeling and evaluation approaches: Validations of WordFish/WordScore based analyses often compare parameter estimates of the different models rather than predictions of these models on held-out data with respect to the same type of labels used to train the models.
 
-Finally Hirst et al conducted a large number of experiments in the Canadian parliament and the European parliament that offer an interesting comparison to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
-The authors show that a classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary.
-They also argue that emotional content plays an important role in automatic analysis of political texts.
-Furthermore their results show a clear depencency between length of a political text and the accuracy with which it can be classified correctly.
+Finally Hirst et al conducted a large number of experiments on data from the Canadian parliament and the European parliament; these experiments can be directly compared to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
+The authors show that a linear classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary.
+The authors also argue that emotional content plays an important role in automatic analysis of political texts.
+Furthermore their results show a clear dependency between length of a political text and the accuracy with which it can be classified correctly.
 
 Taken together, there is a large body of literature in this expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts.
-However except for few exceptions most previous work has focused to a large extent on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or to assignment of a one dimensional policy position (mostly left vs right).
-Often this is a too coarse approach to take into account subtle differences in political policies.
-This work focuses on more fine grained political view prediction: for one the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies; second the political view labels considered are more fine grained than in previous studies.
-Another important contribution of this work is that many existing studies are primarily concerned with a posteriori analysis of historical data.
+Except for few exceptions most previous work has focused on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or on assignment of a one dimensional policy position (mostly left vs right).
+Yet many applications require to take into account more subtle differences in political policies.
+This work focuses on more fine grained political view prediction: for one, the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies; second the political view labels considered are more fine grained than in previous studies.
+While previous studies used such labels only for partitioning training data \cite{Slapin08ascaling} (which is not possible at test time in real-world applications where these labels are not known) the experiments presented in this study directly predict these labels.
+Another important contribution of this work is that many existing studies are primarily concerned with {\em a posteriori} analysis of historical data.
 This work aims at prediction of political bias on out-of-domain data with a focus on the practical application of the model on new data, for which a prototypical web application is provided.
-The experiments estimate out-of-domain generalization and thus complement the work of \cite{Yu2008, Hirst2014} with results from data of the german parliament.
-In particular the results of this study confirm and extend some of their findings on why and when generalization performance is impacted negatively.
+The experiments on out-of-domain generalization complement the work of \cite{Yu2008, Hirst2014} with results from data of the german parliament and novel sentiment analyses.
 
@@
 The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases.
-It is worth noting that even if the accuracies are (as those of comparable studies \cite{Yu2008, Hirst2014}) are not perfect, they nonetheless are on average above chance.
+It is worth noting that even if the accuracies are not perfect, they are above chance and comparable with results of comparable studies \cite{Yu2008, Hirst2014}.
 While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.
@@
 Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain.
-This study tried to quantify the impact of this, but not in all settings out of domain data was available to quantify generalisation performance across text domains.
-For the cases where evaluation data from two domains was available there was a pronounced drop in performance between in domain evaluation set and out of domain evaluation set.
+This study tried to quantify the impact of this effect.
+For the cases in which evaluation data from two domains was available there was a pronounced drop in prediction accuracy between the in domain evaluation set and the out of domain evaluation set.
 This effect was reported previously for similar data, see e.g.
@@
 Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}.
-However when considering texts of sufficient length (for instance by aggregating all texts of a given political topic, see \autoref{tab:results_topic}) the classification performance became better; in some cases reliable predictions can be obtained even beyond the training text domain.
+When considering texts of sufficient length (for instance by aggregating all texts of a given political topic) classification performance improved and in some cases reliable predictions could be obtained even beyond the training text domain.
 
@@
 Second analysing the word-party correlations shows that some discriminative words can be related to the political views of a party; this allows for validation of the models by human experts.
-Third when correlating the sentiment of a speech with proxies of political power there is a strong positive correlation between political power and positive sentiment.
+Third when correlating the sentiment of a speech with measures of political power there is a strong positive correlation between political power and positive sentiment.
 While such an insight in itself might seem not very surprising this quantifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data.

### 2016-06-02 manuscript/rebuttal.txt
Added news paper validation
Comparing party predictions of news predictions with human judgements of party affiliation of news papers, taken from http://www.mediasystemsineurope.org/download.htm

@@
 
-I do hope that the reviewers would consider taking this into account and I hope that it would be possible to add these added references and their discussion to the camera ready version of the manuscript?
-
-An updated version can be found here
+An updated version can be found here https://github.com/felixbiessmann/fipi/blob/poltext/manuscript/political_bias_prediction.pdf; 
+related work is discussed in the new section 2.
+I do hope that the reviewers would consider taking this into account.
 

### 2016-06-02 api.py

@@ -17,9 +17,9 @@ DEBUG = os.environ.get('DEBUG') != None
 VERSION = 0.1
 
 # Schedules news reader to be run at 00:00
-scheduler = BackgroundScheduler()
-scheduler.add_job(get_news, 'interval', minutes=360)
-scheduler.start()
+#scheduler = BackgroundScheduler()
+#scheduler.add_job(get_news, 'interval', minutes=360)
+#scheduler.start()
 
 @retry(stop_max_attempt_number=5)
 def fetch_url(url):

### 2016-06-02 bundestag.py

@@ -197,6 +197,17 @@ def list_top_words(legis=17,topwhat=10):
         pylab.rc('font', **font)
         pylab.savefig(OUT_DIR+'/party_word_correlations-%s-%d.pdf'%(party,legis),bbox_inches='tight')
 
+def newsFileTuple(nf):
+    return [(x['text'],x['source']) for x in json.load(open(nf))]
+
+def get_raw_text_news(folder="data", prefix="news-2016"):
+    '''
+    Loads raw text and labels from news article files
+    (Downloaded using newsreader.get_daily_news())
+    '''
+    files = glob.glob(folder+"/"+prefix+"*")
+    return zip(*chain(*filter(None,map(newsFileTuple,files))))
+
 def get_raw_text_manifesto(folder="data", legislationPeriod=17):
     '''
     Loads raw text and labels from manifestoproject csv files
@@ -257,6 +268,51 @@ def csv2DataTupleTopics(f):
     contentByTopic = df.groupby('topic')['content'].sum()
     return zip(contentByTopic.tolist(), [party[0] + ":%d"%topic for topic in contentByTopic.index.tolist()])
 
+def news_experiment(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='default'):
+    stops = get_stops()
+    text_clf = Pipeline([('vect', CountVectorizer()),
+                            ('tfidf', TfidfTransformer()),
+                            ('clf',LogisticRegression(class_weight='balanced'))])
+    parameters = {'vect__ngram_range': [(1,2)],\
+           #'vect__stop_words':(None,stops),\
+           #'tfidf__use_idf': (True,False),\
+           'clf__C': (10.**sp.arange(1,5,2.)).tolist(),
+           #'vect__max_df':[.3, .5, .75],
+        }
+    saveId = idstr+"-"+randid()
+    X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
+    # optimize hyperparams on training set
+    gs_clf = GridSearchCV(text_clf, parameters, cv=StratifiedKFold(y_train, folds), n_jobs=-1,verbose=1)
+    # train on training set
+    best_clf = gs_clf.fit(X_train, y_train).best_estimator_
+    # test on test set
+    test_clf = text_clf.set_params(**best_clf.get_params()).fit(X_train,y_train)
+    predictedTest = test_clf.predict(X_test)
[…]

### 2016-06-02 newsreader.py

@@ -16,7 +16,7 @@ from scipy import double,zeros
 
 codes = [x['label'] for x in json.load(open("data/nullPrediction.json"))['manifestocode']]
 
-def get_news(sources=['spiegel','faz','welt','zeit']):
+def get_news(sources=['spiegel','sz','faz','welt','zeit']):
     '''
     Collects all news articles from political ressort of major German newspapers
     Articles are transformed to BoW vectors and assigned to a political party
@@ -61,7 +61,7 @@ def get_news(sources=['spiegel','faz','welt','zeit']):
             titles = site.findAll("a", { "class" : "as_teaser-kicker" })
             urls = [a['href'] for a in titles]
          
-        if source is 'sz-without-readability':
+        if source is 'sz':
             # fetching articles from sueddeutsche.de/politik
             url = 'http://www.sueddeutsche.de/politik'
             site = BeautifulSoup(urllib2.urlopen(url).read())
@@ -99,6 +99,81 @@ def get_news(sources=['spiegel','faz','welt','zeit']):
     # store current news and topics
     json.dump(articles,open('news.json','wb'))
     json.dump(topics,open('topics.json','wb'))
+
+def get_daily_news(sources=['faz','welt','sz']):
+    '''
+    Collects all news articles from political ressort of major German newspapers
+    Articles are transformed to BoW vectors and assigned to a political party
+    For better visualization, articles' BoW vectors are also clustered into topics
+
+    INPUT
+    folder      the model folder containing classifier and BoW transformer
+    sources     a list of strings for each newspaper for which a crawl is implemented
+                default ['zeit','sz']
+
+    '''
+    from bs4 import BeautifulSoup
+    from api import fetch_url
+    import urllib2
+    
+    articles = []
+    
+    for source in sources:
+
+        if source is 'spiegel':
+            # fetching articles from sueddeutsche.de/politik
+            url = 'http://www.spiegel.de/politik'
+            site = BeautifulSoup(urllib2.urlopen(url).read())
+            titles = site.findAll("div", { "class" : "teaser" })
+            urls = ['http://www.spiegel.de'+a.findNext('a')['href'] for a in titles]
+         
+        if source is 'faz':
+            # fetching articles from sueddeutsche.de/politik
+            url = 'http://www.faz.net/aktuell/politik'
+            site = BeautifulSoup(urllib2.urlopen(url).read())
[…]

### 2016-06-08 bundestag.py
more hyperparameters for news experiment.

@@ -274,10 +274,10 @@ def news_experiment(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='d
                             ('tfidf', TfidfTransformer()),
                             ('clf',LogisticRegression(class_weight='balanced'))])
     parameters = {'vect__ngram_range': [(1,2)],\
-           #'vect__stop_words':(None,stops),\
-           #'tfidf__use_idf': (True,False),\
-           'clf__C': (10.**sp.arange(1,5,2.)).tolist(),
-           #'vect__max_df':[.3, .5, .75],
+           'vect__stop_words':(None,stops),\
+           'tfidf__use_idf': (True,False),\
+           'clf__C': (10.**sp.arange(1,5,1.)).tolist(),
+           'vect__max_df':[.3, .5, .75],
         }
     saveId = idstr+"-"+randid()
     X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
@@ -309,9 +309,9 @@ def news_experiment(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='d
     npp = [[corrs[idx] for idx,evl in enumerate(evalLabels) if evl == l] for l in labels.keys()]
     perc = [25,50,75]
     corrsPercentiles = dict([(np,[sp.percentile(npp[labels.keys().index(np)],p) for p in perc]) for np in labels.keys()])
-    import pdb;pdb.set_trace()
     
     print("\n".join([x[0]+":\tperc "+"\t".join(["%d:%0.2f"%y for y in zip(perc,x[1])]) for x in corrsPercentiles.items()]))
+    import pdb;pdb.set_trace()
 
 def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='default'):
     stops = get_stops()

### 2016-06-08 manuscript/poltext.tex
now with text in correct file

--- 2016-05-24 manuscript/political_bias_prediction.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
-%
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
-%
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%% ICML 2015 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\documentclass[runningheads,a4paper]{llncs}
+% Use the following line _only_ if you're still using LaTeX 2.09.
+%\documentstyle[icml2015,epsf,natbib]{article}
+% If you rely on Latex2e packages, like most moden people use this:
+\documentclass{article}
 
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-% For citations
+% use Times
+\usepackage{times}
+% For figures
+\usepackage{graphicx} % more modern
+%\usepackage{epsfig} % less modern
+\usepackage{subfigure}
+
 \usepackage{natbib}
@@
 \usepackage{pbox}
-\usepackage{url}
-\urldef{\mailsa}\path|{felix.biessmann,|
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+
+% As of 2011, we use the hyperref package to produce hyperlinks in the
+% resulting PDF.  If this breaks your system, please commend out the
+% following usepackage line and replace \usepackage{icml2015} with
+% \usepackage[nohyperref]{icml2015} above.
+\usepackage{hyperref}
+
+% Packages hyperref and algorithmic misbehave sometimes.  We can fix
+% this with the following command.
+\newcommand{\theHalgorithm}{\arabic{algorithm}}
+
+% Employ the following version of the ``usepackage'' statement for
+% submitting the draft version of the paper for review.  This will set
+% the note in the first column to ``Under review.  Do not distribute.''
+\usepackage{icml2015}
+
+% Employ this version of the ``usepackage'' statement after the paper has
+% been accepted, when creating the final version.  This will set the
+% note in the first column to ``Proceedings of the...''
+%\usepackage[accepted]{icml2015}
+
+% The \icmltitle you define below is probably too long as a header.
+% Therefore, a short form for the running title is supplied here:
+\icmltitlerunning{Automating Political Bias Prediction}
 
@@
 
-\mainmatter  % start of an individual contribution
+\twocolumn[
+\icmltitle{Automating Political Bias Prediction}
 
-% first the title is needed
-\title{Automating Political Bias Prediction}
+% It is OKAY to include author information, even for blind
+% submissions: the style file will automatically remove it for you
+% unless you've provided the [accepted] option to the icml2015
+% package.
+\icmlauthor{Felix Biessmann}{felix.biessmann@gmail.com}
+\icmladdress{}
+\icmlauthor{Pola Lehmann}{pola.lehmann@wzb.eu}
+\icmladdress{Wissenschaftszentrum Berlin}
 
-% a short form should be given in case it is too long for the running head
-\titlerunning{Political Bias Prediction}
+% You may provide any keywords that you 
+% find helpful for describing your paper; these are used to populate 
+% the "keywords" metadata in the PDF but will not be shown in the document
+\icmlkeywords{}
 
-% the name(s) of the author(s) follow(s) next
-%
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
-%
-\author{
-Felix Biessmann
-\thanks{felix.biessmann@gmail.com}
-%
-\authorrunning{Political Bias Prediction}}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-%\institute{
-%\mailsa\\
-%\mailsb\\
-%\mailsc\\
-%\url{http://www.springer.com/lncs}}
-
-%
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
-%
-
-\toctitle{Political Bias Prediction}
-\tocauthor{}
-\maketitle
+\vskip 0.3in
+]
 
@@
 In the following \autoref{sec:related} briefly surveys some related work, thereafter \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
-
-\section{Related Work}\label{sec:related}
-Throughout the last years automated content analyses for political texts have been conducted on a variety of text data sources (parliament data \, blogs, tweets, news articles, party manifestos) with a variety of methods, including sentiment analysis, stylistic analyses, standard bag-of-word (BOW) text feature classifiers and more advanced natural language processing tools.
-While a complete overview is beyond the scope of this work, the following paragraphs list similarities and differences between this study and previous work.
-For a more complete overview we refer the reader to \cite{Grimmer2013,Kaal2014}.
-
-A similar approach to the one presented here was taken in \cite{Yu2008}.
-The authors extracted BOW feature vectors and applied linear classifiers to predict political party affiliation of US congress speeches.
-They used data from the two chambers of the US congress, House and Senat, in order to assess generalization performance of a classifier trained on data from one chamber and tested on data from another.
-They found that accuracies of the model when trained on one domain and tested on another were significantly decreased.
-Generalization was also affected by the time difference between the political speeches used for training and those used for testing.
-
-Other work has focused on developing dedicated methods for predicting political bias.
-Two popular methods are WordFish \cite{Slapin08ascaling} and WordScores \cite{Laver2003}, or improved versions thereof, see e.g.
-\cite{Lowe09scalingpolicy}.
-These approaches have been very valuable for {\em a posteriori} analysis of historical data but they do not seem to be used as much for analyses of new data in a predictive analytics setting.
-Moreover direct comparisons of the results obtained with these so called {\em scaling} methods with the results of the present study or those of studies as \cite{Yu2008} are difficult, due to the different modeling and evaluation approaches: Validations of WordFish/WordScore based analyses often compare parameter estimates of the different models rather than predictions of these models on held-out data with respect to the same type of labels used to train the models.
-
-Finally Hirst et al conducted a large number of experiments on data from the Canadian parliament and the European parliament; these experiments can be directly compared to the present study both in terms of methodology but also with respect to their results \cite{Hirst2014}.
-The authors show that a linear classifier trained on parliament speeches uses language elements of defense and attack to classify speeches, rather than ideological vocabulary.
-The authors also argue that emotional content plays an important role in automatic analysis of political texts.
-Furthermore their results show a clear dependency between length of a political text and the accuracy with which it can be classified correctly.
-
-Taken together, there is a large body of literature in this expanding field in which scientists from quantitative empirical disciplines as well as political science experts collaborate on the challenging topic of automated analysis of political texts.
-Except for few exceptions most previous work has focused on binary classification\footnote{Many parliaments only have two parties and many studies chose binary classification schemes within manually defined topics and generic schemes such as conservative vs liberal or left vs right.} or on assignment of a one dimensional policy position (mostly left vs right).
-Yet many applications require to take into account more subtle differences in political policies.
-This work focuses on more fine grained political view prediction: for one, the case of the german parliament is more diverse than two parliament systems, allowing for a distinction between more policies; second the political view labels considered are more fine grained than in previous studies.
-While previous studies used such labels only for partitioning training data \cite{Slapin08ascaling} (which is not possible at test time in real-world applications where these labels are not known) the experiments presented in this study directly predict these labels.
-Another important contribution of this work is that many existing studies are primarily concerned with {\em a posteriori} analysis of historical data.
-This work aims at prediction of political bias on out-of-domain data with a focus on the practical application of the model on new data, for which a prototypical web application is provided.
-The experiments on out-of-domain generalization complement the work of \cite{Yu2008, Hirst2014} with results from data of the german parliament and novel sentiment analyses.
 
@@
 
-\begin{table}[t]
+\begin{table*}[t]
 \caption{
@@
 \end{center}
-\end{table}
+\end{table*}
 
-\begin{table}[t]
+\begin{table*}[t]
 \caption{
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 \end{document}
+
+% This document was modified from the file originally made available by
+% Pat Langley and Andrea Danyluk for ICML-2K. This version was
+% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
+% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
+% slightly modified from the 2009 version by Kiri Wagstaff and 
+% Sam Roweis's 2008 version, which is slightly modified from 
+% Prasad Tadepalli's 2007 version which is a lightly 
+% changed version of the previous year's version by Andrew Moore, 
+% which was in turn edited from those of Kristian Kersting and 
+% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2016-06-10 manuscript/poltext.tex
made some changes, basically with regard to the manifestos and tried to rearrange the results section a little

@@
 %
-Modern media generate a large amount of content at an ever increasing rate.
-Keeping an unbiased view on what media report on requires to understand the political bias of texts.
+Analysis and classifications of political text is and has been a very important tool to generate political science data (Benoit et al.
+2015).
+Traditionally such classifications are done by experts, who read and label the text of interest (see for example the Manifesto Project (Budge et al.
+2001; Klingemann et al.
+2006), the Comparative Agendas Project or Poltext (Pétry/Duval 2015)).
+This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
+The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists.
+Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve (Benoit et al.
+2009)
+
+One area which produces large amounts of text which are of high interest for political scientists is modern media.
+Standard newspapers, but also new forms as twitter, facebook or blogs produce an ever growing amount of text influencing political debate.
+These texts can be of interest to political scientist in various analyses.
+The question we will focus on in this analysis is the potential political bias that can be detected in these sources.
 In many cases it is obvious which political bias an author has.
 In other cases some expertise is required to judge the political bias of a text.
-%
-When dealing with large amounts of text however there are simply not enough experts to examine all possible sources and publications.
+Keeping an unbiased view on what media report on requires to understand the political bias of texts.
+Different newspaper outlets are, for example, often ascribed as favoring different positions on the political left-right axis.
+Data from the European Media Systems Survey show that experts clearly associate different media outlets with specific parties (http://www.mediasystemsineurope.org).
+However such expert surveys on the political bias of newspapers can normally only give us a very general tendency of a newspaper, they do not distinguish between different journalists or different sections.
 Assistive technology can help in this context to try and obtain a more unbiased sample of information.
+It gives us the possibility to get a more nuanced view of the spectrum of political positions favored in important news outlets in a country.
+In this paper we will test whether it is possible to automatically detect political bias in text and if such a model can be used on different kind of texts.
+The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
+We will test our model on both in-domain and out-of-domain data.
+Prediction on out-of-domain data is less precise as on in-domain data, but still generating satisfying results.
 
-Ideally one would choose for each topic a sample of reports from the entire political spectrum in order to form an unbiased opinion.
-But ordering media content with respect to the political spectrum at scale requires automated prediction of political bias.
-The aim of this study is to provide empirical evidence  indicating that leveraging open data sources of german texts, automated political bias prediction is possible with above chance accuracy.
-These experimental results  confirm and extend previous findings \cite{Yu2008, Hirst2014}; a novel contribution of this work is a proof of concept which applies this technology to sort news article recommendations according to their political bias.
-
-When human experts determine political bias of texts they will take responsibility for what they say about a text, and they can explain their decisions.
-This is a key difference to many statistical learning approaches.
-Not only is the responsibility question problematic, it can also be difficult to interpret some of the decisions.
 In order to validate and explain the predictions of the models three strategies that allow for better interpretations of the models are proposed.
@@
 
-In the following \autoref{sec:related} briefly surveys some related work, thereafter \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
 
@@
 %
-All experiments were run on publicly available data sets of german political texts and standard libraries for processing the text.
+All experiments were run on publicly available data sets of German political texts and standard libraries for processing the text.
 The following sections describe the details of data acquisition and feature extraction.
@@
 \subsection{Data}
-Annotated political text data was obtained from two sources: a) the discussions and speeches held in the german parliament ({\em Bundestag}) and b) all manifesto texts of parties running for election in the german parliament in the current 18th and the last, 17th, legislation period.
+Annotated political text data was obtained from two sources: a) the discussions and speeches held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the current 18th and the last, 17th, legislative period.
+For the 18th period we also added the manifesto data of three additional parties that did not win seats but nonetheless played a very important role in the political discourse in the run-up to the election.
 
 \paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
-The texts of parliament protocols are available through the website of the german bundestag\footnote{\url{https://www.bundestag.de/protokolle}}; an open source API was used to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
+The protocols of plenary debates are available through the website of the German Bundestag\footnote{\url{https://www.bundestag.de/protokolle}}; an open source API was used to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
 In total 22784 speeches were extracted for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.
@@
 \paragraph{Party manifesto data}
-For party manifestos another openly accessible API was used, provided by the Wissenschaftszentrum Berlin (WZB).
-The API is released as part of the {\em Manifestoproject} \cite{manifesto}.
-The data released in this project comprises the complete manifestos for each party that ran for election enriched with annotations by political experts.
-Each sentence (in some cases also parts of sentences) is annotated with one of 56 political labels.
-Examples of these labels are {\em pro/contra protectionism, decentralism, centralism, pro/contra welfare}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
-The set of labels was developed by political scientists at the WZB and released for public use.
-All manifestos of parties that were running for election in this and the last legislative period were obtained.
-In total this resulted in 29451 political statements that had two types of labels: First the party affiliation of each political statement; this label was used to evaluate the party evaluation classifiers trained on the parliament speeches.
+The party manifesto text was taken from the Manifesto Corpus (add citation).
+The Corpus is accessible through an publicly available API published by the {\em Manifesto Project} \cite{manifesto}.
+The data released in this project mainly comprises the complete manifestos of all parties that have won seats at a national election.
+Each quasi-sentence \footnote{A quasi-sentence has the length of an argument.
+It is never longer than one sentence.} is annotated with one of 56 policy goals.
+Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
+In total the 12 manifestos from the 17th and 18th legislative period have been divided into 29451 quasi-sentences.
+Each observation has been added with two labels: the party affiliation and the manually assigned policy goal aimed at in each specific statement.
+The party affiliation was used to evaluate the party evaluation classifiers trained on the parliament speeches.
 For this purpose the data acquisition was constrained to only those parties that were elected into the parliament.
-Next to the party affiliation the political view labels were extracted.
-For the analyses based on political view labels all parties were considered, also those that did not make it into the parliament.
+For the analyses based on policy issues all parties were considered, also those three that did not make it into parliament.
 
 The length of each annotated statement in the party manifestos was rather short.
-The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters.
-Measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.
-This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos.
+The median length was 95 characters or 12 words.
+\footnote{The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters.
+Measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.} This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos and it a good test for media data, as some data like tweets, facebook texts etc.
+are often very short, too.
 However for a classifier (as well as for humans) such short sentences can be rather difficult to classify.
-In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into political topics using the manifesto code labels.
-Each political view label is a three digit code, the first digit represents the political domain.
-In total there were eight political domains (topics): {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups} and a topic {\em undefined}, for a complete list see also \cite{leftright}.
-These 8 topics were used to aggregate all statements in each manifesto into topics.
+Benoit et al.
+(2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
+In the Manifesto Project the human coders do not code these statements separately but within context.
+In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains.
+Each of the 56 policy categories from the Manifesto Project belongs to one of seven policy domains.
+These were used to aggregate the data into topics.
+An eight topic is generated for all statements not belonging to any one the 56 policy issues: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups} and a topic {\em undefined}, for a complete list see also \cite{leftright}.
 Most party manifestos covered all eight of them, some party manifestos in the 17th Bundestag only covered seven.
@@
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-First each data set was segmented into semantic units; in the case of parliament discussions this were the speeches, in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 political view labels.
+First each data set was segmented into semantic units; in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 policy issues, in the case of parliament discussions this were the speeches.
 Parliament speeches were often interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.
@@
 \item {\bf Classification of government membership} (binary problem)
-\item {\bf Classification of political views} (56 class problem)
+\item {\bf Classification of policy issues} (56 class problem)
 \end{enumerate}
@@
 Party affiliation is a five class problem for the 17th legislation period, and a four class problem for the 18th legislation period.
-Political view classification is based on the labels of the manifesto project, see \autoref{sec:data} and \cite{leftright}.
-For each of first two problems, party affiliation and government membership prediction, classifiers were trained on the parliament speeches.
-For the third problem classifiers were trained only on the manifesto data for which political view labels were available.
+The policy issue classification is based on the categories of the Manifesto Project, see \autoref{sec:data} and \cite{leftright}.
+For both the party affiliation and the government membership prediction, classifiers were trained on the parliament speeches.
+For the third problem classifiers were trained only on the manifesto data for which policy issue labels were available.
 
@@
 Out-of-domain evaluation data were the texts from party manifestos.
-For the political view prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
+For the policy issue prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
 
@@
 
-The following sections give an overview of the results for all political bias prediction tasks.
+The following section gives an overview of the results for all political bias prediction tasks.
 Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section.
@@
 \subsection{Predicting political party affiliation}
-The results for the political party affiliation prediction on held-out parliament data and on evaluation data are listed in \autoref{tab:results_17} for the 17th Bundestag and in \autoref{tab:results_18} for the 18th Bundestag, respectively.
-Shown are the evaluation results for in-domain data (held-out parliament speech texts) as well as the out-of-domain data; the party manifesto out-of-domain predictions were made on the sentence level.
+The results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th and 18th Bundestag are listed in \autoref{tab:results_in-domain}.
+Shown are the evaluation results for in-domain data.
 
@@
 These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem of prediction party affiliation in the European parliament; the accuracy for the 17th Bundestag is 0.63, results of the 18th Bundestag are difficult to compare as the number of parties is four and the legislation period is not finished yet.
-For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4.
+But as the main purpose of this paper is to generate a model that can be used to predict political bias on new data, like media outtakes, we are interested in the question if the model also produces satisfying results on out-of-domain data.
+We are therefore also predicting party affiliation on party manifesto text.
+For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
 This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated based on the topic.
-The manifesto code political topics labels were used to concatenate texts of each party to one of eight topics, see \autoref{sec:data}.
-The topic level results are shown in \autoref{tab:results_topic} and \autoref{tab:confusion_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve in the case of some parties reliably precision and recall values close to 1.0.
+In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated based on the political domain into which each quasi-sentence was coded, see \autoref{sec:data}.
+The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve in the case of some parties reliable precision and recall values close to 1.0.
 This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
@@
 \caption{
-\label{tab:results_17}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on test set and evaluation set, respectively.
-Predictions on the manifesto data was done on {\bf sentence level}; $N$ denotes number of data points in the evaluation set.
+\label{tab:results_in-domain}
+Classification performance on the party affiliation prediction problem for data from the 17th and 18th legislative period on in-domain data. $N$ denotes number of data points in the evaluation set.
 }
@@
 \begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Held-out parliament speeches}\\
+& \multicolumn{4}{c}{\bf 17th legislative period} & \multicolumn{4}{c}{\bf 18th legislative period}\\
     &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
 \hline \hline
-       cducsu   &    0.62  &    0.81  &    0.70  &     706&0.26    &  0.58   &   0.36    &  2030\\
-        fdp    &   0.70   &   0.37  &    0.49    &   331&0.38   &   0.28    &  0.33   &   2319\\
-     gruene &      0.59  &    0.40   &   0.48   &    298&0.47  &    0.20   &   0.28 &     3747\\
-      linke    &   0.71   &   0.61  &    0.65    &   338&0.30   &   0.47   &   0.37    &  1701\\
-        spd   &    0.60   &   0.69  &    0.65   &    606&0.26   &   0.16   &   0.20  &    2278\\
+       cducsu   &    0.62  &    0.81  &    0.70  &     706&0.66    &  0.82   &   0.73    &  456\\
+        fdp    &   0.70   &   0.37  &    0.49    &   331&   &      &     &   0\\
+     gruene &      0.59  &    0.40   &   0.48   &    298&0.68  &    0.54   &   0.60 &     173\\
+      linke    &   0.71   &   0.61  &    0.65    &   338&0.77   &   0.58   &   0.66    &  173\\
+        spd   &    0.60   &   0.69  &    0.65   &    606&0.60   &   0.54   &   0.57  &    330\\
 \hline
-avg / total &      0.64   &   0.63   &   0.62    &  2279 &0.35    &  0.31 &     0.30   &  12075
+avg / total &      0.64   &   0.63   &   0.62    &  2279 &0.66    &  0.66 &     0.30   &  1132
 %
@@
 \caption{
-\label{tab:results_18}
+\label{tab:results_out-of-domain}
 Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively.
@@
 \begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+& \multicolumn{4}{c}{\bf Party Manifestos} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+& \multicolumn{4}{c}{\bf 17th legislative period} & \multicolumn{4}{c}{\bf 18th legislative period}\\
     &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
 \hline \hline
-    cducsu    &   0.66   &   0.82   &   0.73    &   456 & 0.32  &    0.64  &    0.43    &  2983\\
-     gruene   &    0.68    &  0.54   &   0.60    &   173   &0.59   &   0.15   &   0.23   &   5674\\
-      linke     &  0.77  &    0.58    &  0.66    &   173 & 0.36   &   0.48   &   0.41   &   2555\\
-        spd     &  0.60  &    0.54   &   0.57    &   330 & 0.26 &     0.31   &   0.28     & 2989\\
+    cducsu    &   0.26   &   0.58   &   0.36    &   2030 & 0.32  &    0.64  &    0.43    &  2983\\
+    fdp    &   0.38   &   0.28   &   0.33    &   2319 &   &      &        &  0\\
+     gruene   &    0.47    &  0.20   &   0.28    &  3747   &0.59   &   0.15   &   0.23   &   5674\\
+      linke     &  0.30  &    0.47    &  0.37    &   1701 & 0.36   &   0.48   &   0.41   &   2555\\
+        spd     &  0.26  &    0.16   &   0.20    &   2278 & 0.26 &     0.31   &   0.28     & 2989\\
 \hline
-avg / total    &   0.66  &    0.66  &    0.65   &   1132&  0.42 &     0.34  &    0.32&     14201\\
+avg / total    &   0.35  &    0.31  &    0.30   &   12075 &  0.42 &     0.34  &    0.32&     14201\\
 %
@@
 \begin{tabular}{lcccc}
+& \multicolumn{4}{c}{\bf Party Manifestos} \\
 & \multicolumn{4}{c}{\bf 17th Bundestag} \\
@@
 \begin{tabular}{lcccc}
+& \multicolumn{4}{c}{\bf Party Manifestos} \\
 & \multicolumn{4}{c}{\bf 18th Bundestag} \\
@@
 
-In order to investigate the errors the models made confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
+In order to investigate the errors conducted by the models confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
 One example illustrates that the mistakes the model makes can be associated with changes in the party policy.
 The green party has been promoting policies for renewable energy and against nuclear energy in their manifestos prior to both legislative periods.
-Yet the statements of the green party are more often predicted to be from the government parties than from the party that originally promoted these green ideas, reflecting the trend that these legislative periods governing parties took over policies from the green party.
-This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag predicts all manifesto topics of the green party to be from one of the parties of the governing coalition, CDU/CSU or SPD.
+Yet the statements of the green party are more often predicted to be from the government parties than from the party that originally promoted these green ideas, reflecting the trend that during the legislative period governing parties took over policies from the green party.
+This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag predicts all manifesto content of the green party to be from one of the parties of the governing coalition, CDU/CSU or SPD.
 \\
-
-\paragraph{Government membership prediction} Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag.
-In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
-While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
-This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
-The authors report classification accuracies between 0.8 and 0.87, the accuracy in the 17th Bundestag was 0.85.
-While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
 
@@
 \begin{tabular}{lccccccc}
+ \multicolumn{8}{c}{\bf Party Manifestos}\\
  \multicolumn{8}{c}{\bf 17th Bundestag}\\
@@
 \begin{tabular}{lcccccc}
+ \multicolumn{7}{c}{\bf Party Manifestos}\\
  \multicolumn{7}{c}{\bf 18th Bundestag}\\
@@
 \subsection{Predicting political views}
+%Dieser Absatz hat ja nicht wirklich was mit political bias zu tun. Wollen wir den also aus K\"uerzungsgr\"unden rausnehmen?
 Parties change their policies and positions in the political spectrum.
-More reliable categories for political bias are party independent labels for political views, see \autoref{sec:data}.
+More reliable categories for political bias are party independent labels for policy issues, see \autoref{sec:data}.
 A separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:model}.
-As there was no out-of-domain evaluation set available in this setting only evaluation error on in-domain data is reported.
-Note however that also in this experiment the evaluation data was never seen by any model during training time.
+As there was no out-of-domain evaluation set available in this setting only the evaluation error on in-domain data is reported.
+Note however that in this experiment too the evaluation data was never seen by any model during training time.
 In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes.
@@
 \subsection{Correlations between words and parties}\label{sec:word_party_correlations}
-
-The 10 highest and lowest correlations between individual words and the party affiliation label are shown for each party in \autoref{fig:party_word_correlations}.
+We now know that party affiliation can be predicted using our model.
+But which words are most important for the classification performed here?
+To get a more precise idea of what differentiates the parties from each other the 10 highest and lowest correlations between individual words and the party affiliation label are shown for each party in \autoref{fig:party_word_correlations}.
 Correlations were computed on the data from the current, 18th, legislative period.
@@
 \paragraph{\bf Left party (linke)}
-The left party mostly criticises measures that affect social welfare negatively, such as the {\em Hartz IV} program.
-Main actors that are blamed for decisions of the conservative governments by the left party are big companies ({\em konzerne}).
-Rarely the party addresses concerns related to security ({\em sicherheit}).
+The Left party talks mainly about issues related to work ({\em Besch\"aftigten}) and unemployment ({\em Hartz IV, Erwerbslosen}). 
 \paragraph{\bf Green party (gruene)}
-The green party heavily criticised the secret negotiations about the TiSA agreement\footnote{\url{https://en.wikipedia.org/wiki/Trade_in_Services_Agreement}} and insists in formal inquiries that the representatives of the green party put forward in this matter ({\em fragen, anfragen}).
-They also often ask questions related to army projects ({\em R\"ustungsprojekte, Wehrbericht}) or the military development in east europe ({\em  Jalta}\footnote{Referring to the \url{https://en.wikipedia.org/wiki/Yalta_Conference}}).
-\paragraph{\bf Social democratic party (SPD)}
-The social democrats often use words related to rights of the working class, as reflected by the heavy use of the {\em International Labour Organisation} (ILO) or rights of employes ({\em Arbeitnehmerrechte}).
-They rarely talk about competition ({\em Wettbewerb}) or climate change ({\em klimapolitik}).
+The Green party talks about the formal procedures available to opposition parties to question government policy ({\em fragen, anfragen}) and, interestingly, about the reform of the Germany army ({\em R\"ustungsprojekte, Wehrbericht}).
+\paragraph{\bf Social Democratic Party (SPD)}
+The SPD often uses words related to rights of the working class ({\em International Labour Organisation, Arbeitnehmerrechte}) and the governing coalition ({\em bundestagsfraktion, koalitionspartner}).
 
-\paragraph{\bf Conservative party (CDU/CSU)}
-The conservative christian party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
-The latter could be related to the ongoing debates about whether or not the governments should be allowed to collect data and thus restrict fundamental civil rights in order to better secure the population.
-In contrast to the parties of the opposition, the conservatives rarely mention the word war ({\em krieg}) or related words.
+\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+The CDU/CSU party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
 
@@
 \begin{center}
-\includegraphics[width=2.8cm]{images/party_word_correlations-linke-18.pdf} 
-\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-18.pdf} 
-\includegraphics[width=3cm]{images/party_word_correlations-spd-18.pdf} 
-\includegraphics[width=3cm]{images/party_word_correlations-cducsu-18.pdf}
+\includegraphics[width=2.8cm]{images/party_word_correlations-linke-17.pdf}
+\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-17.pdf}
+\includegraphics[width=3cm]{images/party_word_correlations-spd-17.pdf}
+\includegraphics[width=3cm]{images/party_word_correlations-cducsu-17.pdf}
 %
@@
 
-\subsection{Speech sentiment correlates with political power}\label{sec:sentiment_result}
-In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
+\subsection{Predicting government status}\label{sec:sentiment_result}
+Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag.
+In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
+While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
+This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
+The authors report classification accuracies between 0.8 and 0.87, the accuracy in the 17th Bundestag was 0.85.
+While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
+\paragraph{Speech sentiment correlates with political power} In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
 The average sentiment of each political party is shown in \autoref{fig:party_sentiments}.
@@
 The results show an interesting relationship between political power and sentiment.
-Political power was evaluated in two ways: a) in terms of the number of seats a party has and b) in terms of membership of the government.
+Political power was evaluated in terms of membership of the government.
 Correlating either of these two indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
@@
 \subsection{An example web application}
-To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major german news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
-For each news article topic, headlines of articles are plotted along with the predictions of the political view of an article and two labels derived deterministically from the 56 class output, a left right index and the political domain of a text, see \cite{leftright}.
+To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major German news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
+For each news article topic, headlines of articles are plotted along with the predictions of the policy issue of an article and two labels derived deterministically from the 56 class output, a left right index and the political domain of a text, see \cite{leftright}.
 Within each topic it is then possible to get an ordered (from left to right) overview of the articles on that topic.
@@
 \begin{center}
-\includegraphics[width=10cm]{images/fipi-screenshot}
+%\includegraphics[width=10cm]{images/fipi-screenshot}
 %
@@
 \subsection*{Acknowledgements}
-I would like to thank Friedrich Lindenberg for factoring out the \url{https://github.com/bundestag/plpr-scraper} from his bundestag project.
+We would like to thank Friedrich Lindenberg for factoring out the \url{https://github.com/bundestag/plpr-scraper} from his Bundestag project.
 Some backend configurations for the web application were taken from an earlier collaboration with Daniel Kirsch.
-Pola Lehmann and Michael Gaebler provided helpful feedback on an earlier version of the manuscript.
-Pola Lehman also helped with getting access to and documentation on the Manifestoproject data.
+Michael Gaebler provided helpful feedback on an earlier version of the manuscript.
 %

### 2016-06-12 manuscript/poltext.tex
Some deletions.

@@
 2015).
-Traditionally such classifications are done by experts, who read and label the text of interest (see for example the Manifesto Project (Budge et al.
+Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project (Budge et al.
 2001; Klingemann et al.
-2006), the Comparative Agendas Project or Poltext (Pétry/Duval 2015)).
-This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
+2006), the Comparative Agendas Project or Poltext (Pétry/Duval 2015).} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
 The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists.
@@
 \subsection{Data}
-Annotated political text data was obtained from two sources: a) the discussions and speeches held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the current 18th and the last, 17th, legislative period.
-For the 18th period we also added the manifesto data of three additional parties that did not win seats but nonetheless played a very important role in the political discourse in the run-up to the election.
+Annotated political text data was obtained from two sources: a) the plenary debates held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the current 18th and the last, 17th, legislative period.
+
+%For the 18th period we also added the manifesto data of three additional parties that did not win seats but nonetheless played a very important role in the political discourse in the run-up to the election.
+Wollen wir die einfach rauslassen?
+Käme ja eh nur in einem kleinen Teil der Analyse vor und macht die Beschreibung an mehreren Stellen kürzer.
 
@@
 The data released in this project mainly comprises the complete manifestos of all parties that have won seats at a national election.
-Each quasi-sentence \footnote{A quasi-sentence has the length of an argument.
-It is never longer than one sentence.} is annotated with one of 56 policy goals.
+Each quasi-sentence\footnote{A quasi-sentence has the length of an argument.
+It is never longer than one sentence.} is annotated with one of 56 policy issue categories.
 Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
-In total the 12 manifestos from the 17th and 18th legislative period have been divided into 29451 quasi-sentences.
-Each observation has been added with two labels: the party affiliation and the manually assigned policy goal aimed at in each specific statement.
-The party affiliation was used to evaluate the party evaluation classifiers trained on the parliament speeches.
+In total the 9 manifestos from the 17th and 18th legislative period have been divided into 29451 quasi-sentences.
+Each observation has been added with two labels: the party affiliation and the manually assigned policy issue aimed at in each specific statement.
+
+%The party affiliation was used to evaluate the party evaluation classifiers trained on the parliament speeches.
 For this purpose the data acquisition was constrained to only those parties that were elected into the parliament.
@@
 
-The length of each annotated statement in the party manifestos was rather short.
-The median length was 95 characters or 12 words.
-\footnote{The longest statement was 522 characters long, the 25\%/50\%/75\% percentiles were 63/95/135 characters.
-Measured in words the longest data point was 65 words and the 25\%/50\%/75\% percentiles were 8/12/17 words, respectively.} This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos and it a good test for media data, as some data like tweets, facebook texts etc.
+The length of each annotated statement in the party manifestos is rather short.
+The median length is 95 characters or 12 words.\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.} This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos and is a good test for media data, as some data like tweets, facebook texts etc.
 are often very short, too.
@@
 (2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
-In the Manifesto Project the human coders do not code these statements separately but within context.
-In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains.
-Each of the 56 policy categories from the Manifesto Project belongs to one of seven policy domains.
-These were used to aggregate the data into topics.
-An eight topic is generated for all statements not belonging to any one the 56 policy issues: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups} and a topic {\em undefined}, for a complete list see also \cite{leftright}.
-Most party manifestos covered all eight of them, some party manifestos in the 17th Bundestag only covered seven.
+In the Manifesto Project the human coders therefore do not code these statements separately but within context.
+In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains using the 56policy categories from the Manifesto Project.
+Within the Manifesto Project category scheme each of these belongs to one of seven policy domains.
+These were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
+An eight topic is generated for all statements not belonging to any one the 56 policy issues, for a complete list see also \cite{leftright}.
+Most party manifestos covered all eight topics, some party manifestos in the 17th Bundestag only covered seven.
 
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-First each data set was segmented into semantic units; in the case of the party manifesto data semantic units were the sentences or sentence parts associated with one of the 56 policy issues, in the case of parliament discussions this were the speeches.
+First each data set was segmented into semantic units; in the case of the party manifesto data the semantic units were the quasi-sentences associated with one of the 56 policy issues, in the case of parliament discussions this were the speeches.
 Parliament speeches were often interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.
@@
 
-\section{Classification Model and Training Procedure}\label{sec:model}
+\section{Classification Model and Training}\label{sec:model}
 Bag-of-words feature vectors were used to train a multinomial logistic regression model.
@@
 \end{eqnarray}
-%
-We estimated $\vec{W}$ using quasi-newton gradient descent.
-The optimization function was obtained by adding a penalization term to the negative log-likelihood of the multinomial logistic regression objective and the optimization hence found the $\vec{W}$ that minimized
-\begin{equation}\label{eq:objective}
-L(\vec{W}, \vec{x}, \gamma) = - \log{\frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}}}+ \gamma \| \vec{W} \|_{F}
-\end{equation}
-Where $\|~\|_F$ denotes the Frobenius Norm and $\gamma$ is a regularization parameter controlling the complexity of the model.
-The regularization parameter was optimized on a log-scaled grid from $10^{-4,\dots,4}$.
 The performance of the model was optimized using the classification accuracy, but we also report all other standard measures, precision ($TP / (FP + TP$), recall ($TP / (TP + FN)$) and f1-score ($2\times (Prec.
@@
 For the policy issue prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
+% Falls wir die Medien nochreinnehmen, muss das bei out-of-domain auch noch erwähnt werden.
 
@@
 Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section.
+Because of space restrictions, we will only report results fromt he 17th legislative period here.
+Results are very similar for the 18th legislative period and can be accessed via github.
 
 \subsection{Predicting political party affiliation}
-The results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th and 18th Bundestag are listed in \autoref{tab:results_in-domain}.
-Shown are the evaluation results for in-domain data.
+The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
 
 When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
-These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem of prediction party affiliation in the European parliament; the accuracy for the 17th Bundestag is 0.63, results of the 18th Bundestag are difficult to compare as the number of parties is four and the legislation period is not finished yet.
-But as the main purpose of this paper is to generate a model that can be used to predict political bias on new data, like media outtakes, we are interested in the question if the model also produces satisfying results on out-of-domain data.
+These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament; the accuracy for the 17th Bundestag is 0.63, results of the 18th Bundestag are difficult to compare as the number of parties is four and the legislation period is not finished yet.
+
+But as the main purpose of this paper is to generate a model that can be used to predict political bias on new data, like media texts, we are interested in the question if the model also produces satisfying results on out-of-domain data.
 We are therefore also predicting party affiliation on party manifesto text.
@@
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated based on the political domain into which each quasi-sentence was coded, see \autoref{sec:data}.
-The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out of domain data can achieve in the case of some parties reliable precision and recall values close to 1.0.
+In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into the previously described party domains, see \autoref{sec:data}.
+The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
 This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
@@
 \label{tab:results_in-domain}
-Classification performance on the party affiliation prediction problem for data from the 17th and 18th legislative period on in-domain data. $N$ denotes number of data points in the evaluation set.
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on in-domain data. $N$ denotes number of data points in the evaluation set.
 }
 \begin{center}
-\begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Held-out parliament speeches}\\
-& \multicolumn{4}{c}{\bf 17th legislative period} & \multicolumn{4}{c}{\bf 18th legislative period}\\
-    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\begin{tabular}{lcccc}
+    &         precision    &recall &  f1-score  & N  \\
 \hline \hline
-       cducsu   &    0.62  &    0.81  &    0.70  &     706&0.66    &  0.82   &   0.73    &  456\\
-        fdp    &   0.70   &   0.37  &    0.49    &   331&   &      &     &   0\\
-     gruene &      0.59  &    0.40   &   0.48   &    298&0.68  &    0.54   &   0.60 &     173\\
-      linke    &   0.71   &   0.61  &    0.65    &   338&0.77   &   0.58   &   0.66    &  173\\
-        spd   &    0.60   &   0.69  &    0.65   &    606&0.60   &   0.54   &   0.57  &    330\\
+       cducsu   &    0.62  &    0.81  &    0.70  &     706\\
+        fdp    &   0.70   &   0.37  &    0.49    &   331\\
+     gruene &      0.59  &    0.40   &   0.48   &    298\\
+      linke    &   0.71   &   0.61  &    0.65    &   338\\
+        spd   &    0.60   &   0.69  &    0.65   &    606\\
 \hline
-avg / total &      0.64   &   0.63   &   0.62    &  2279 &0.66    &  0.66 &     0.30   &  1132
+avg / total &      0.64   &   0.63   &   0.62    &  2279 
 %
@@
 \label{tab:results_out-of-domain}
-Classification performance on the party affiliation prediction problem for data from the 18th legislative period on test set and evaluation set, respectively.
-Predictions on the manifesto data was done on {\bf sentence level}.
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on out-of-domain data.  Predictions are done on the {\bf sentence level}.
 }
 \begin{center}
-\begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Party Manifestos} & \multicolumn{4}{c}{\bf Party Manifestos}\\
-& \multicolumn{4}{c}{\bf 17th legislative period} & \multicolumn{4}{c}{\bf 18th legislative period}\\
-    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+\begin{tabular}{lcccc}
+    &         precision    &recall &  f1-score  & N  \\
 \hline \hline
-    cducsu    &   0.26   &   0.58   &   0.36    &   2030 & 0.32  &    0.64  &    0.43    &  2983\\
-    fdp    &   0.38   &   0.28   &   0.33    &   2319 &   &      &        &  0\\
-     gruene   &    0.47    &  0.20   &   0.28    &  3747   &0.59   &   0.15   &   0.23   &   5674\\
-      linke     &  0.30  &    0.47    &  0.37    &   1701 & 0.36   &   0.48   &   0.41   &   2555\\
-        spd     &  0.26  &    0.16   &   0.20    &   2278 & 0.26 &     0.31   &   0.28     & 2989\\
+    cducsu    &   0.26   &   0.58   &   0.36    &   2030 \\
+    fdp    &   0.38   &   0.28   &   0.33    &   2319 \\
+     gruene   &    0.47    &  0.20   &   0.28    &  3747\\
+      linke     &  0.30  &    0.47    &  0.37    &   1701\\
+        spd     &  0.26  &    0.16   &   0.20    &   2278\\
 \hline
-avg / total    &   0.35  &    0.31  &    0.30   &   12075 &  0.42 &     0.34  &    0.32&     14201\\
+avg / total    &   0.35  &    0.31  &    0.30   &   12075\\
 %
@@
 \end{tabular}
-\quad
-\begin{tabular}{lcccc}
-& \multicolumn{4}{c}{\bf Party Manifestos} \\
-& \multicolumn{4}{c}{\bf 18th Bundestag} \\
-    &         precision    &recall &  f1-score  & N  \\
-    \hline
-    \hline
-  cducsu     &  0.50   &   1.00  &    0.67     &    8\\
-       gruene  &     0.00   &   0.00   &   0.00  &       8\\
-      linke     &  1.00   &   0.88   &   0.93  &       8\\
-        spd     &  0.56  &    0.62  &    0.59   &      8\\
-\hline
-avg / total   &    0.51 &     0.62   &   0.55   &     32\\
-\end{tabular}
 \end{center}
@@
 
-In order to investigate the errors conducted by the models confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
-One example illustrates that the mistakes the model makes can be associated with changes in the party policy.
-The green party has been promoting policies for renewable energy and against nuclear energy in their manifestos prior to both legislative periods.
-Yet the statements of the green party are more often predicted to be from the government parties than from the party that originally promoted these green ideas, reflecting the trend that during the legislative period governing parties took over policies from the green party.
-This effect is even more pronounced in the topic level predictions: a model trained on data from the 18th Bundestag predicts all manifesto content of the green party to be from one of the parties of the governing coalition, CDU/CSU or SPD.
+\subsection{Validation}
+\paragraph{Differentiability of policy issues}
+One reason why some statements are difficult to classify could be that some issues are more difficult to differentiate than others.
+To test this a separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:model}.
+As there was no out-of-domain evaluation set available in this setting only the evaluation error on in-domain data is reported.
+Note however that in this experiment too the evaluation data was never seen by any model during training time.
+In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes.
+Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels. Results are not surprising, but are equal to those by human coders. \\
+
+\begin{table}
+\caption{
+\label{tab:results_avg_political_view}
+Classification of 56 political views, see \autoref{sec:data}.
+}
+\begin{center}
+\footnotesize
+\begin{tabular}{lcrrrr}
+code & meaning  &      prec    &recall & N\\
+\hline\hline
+%
+       501  & environmentalism + &      0.62   &   0.61 &      165\\
+        202 &   democracy + &   0.58  &    0.55   &      122\\
+        701    & labour +&  0.57  &    0.54   &   129\\
+                201    &freedom/human rights +  & 0.58   &   0.54   &      159\\
+         106   & peace + & 0.52&      0.57   &       21\\
+\dots\\
+        302     &centralism + & 0.25     & 0.20  &          10\\
+        401    &  free enterprise + &0.20    &  0.19    &       52\\
+        505    &welfare - &  0.13   &   0.14   &      14\\
+        409    & keynesian demand +&  0.14  &      0.13   &      8\\
+            0    & undefined &  0.09  &    0.12   &    17\\
+            \hline
+avg / total  &  &  0.47    &  0.46 &    2946\\
+%
+\end{tabular}
+\end{center}
+\end{table}
+
+\paragraph{Policy change}
+Another explanation for missclassification could lie in the fact that parties change their policy positions.
+In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
+On the topic level results for the 17th legislative period are pretty good except for the SPD.
+In the 18th legislative period predictions are very bad for the Green party.
+One explanation for this missclassification could lie in policy changes.
+For example, the Green party has devoted large parts of its manifesto to renewable and against nuclear energy.
+After Fukushima the other parties adopted these positions.
 \\
 
+% Die könnten wir vielleicht rausnehmen.
 % Confusion matrix sentence level
 \begin{table}[t]\label{tab:conf_mat_four_class}
-\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right); the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
+\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period; the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
 This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
@@
 \end{tabular}
-\hfill
-\begin{tabular}{lcccccc}
- \multicolumn{7}{c}{\bf Party Manifestos}\\
- \multicolumn{7}{c}{\bf 18th Bundestag}\\
- \vspace{1em}\\
-&&& \multicolumn{4}{c}{Predicted}\\
-&&& cducsu & gruene& linke& spd\\
-\hline
-\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&1912& 156& 331& 584\\
-&&gruene&2092& 827& 1311& 1444\\
-&&linke&596& 186& 1216& 557\\
-&&spd&1284& 226& 563& 916\\
-\end{tabular}
-\vspace{1em}
 \end{table}
@@
 \end{tabular}
-\hfill
-\begin{tabular}{lcccccc}
- \multicolumn{7}{c}{\bf Party Manifestos}\\
- \multicolumn{7}{c}{\bf 18th Bundestag}\\
- \vspace{1em}\\
-&&& \multicolumn{4}{c}{Predicted}\\
-&&& cducsu & gruene& linke& spd\\
-\hline
-\multirow{4}{*}{\rotatebox{90}{\pbox{4.7cm}{\centering True}}}&&cducsu&8& 0& 0& 0\\
-&&gruene&4& 0& 0& 4\\
-&&linke&1& 0& 7& 0\\
-&&spd&3& 0& 0& 5\\
-\end{tabular}
-\vspace{1em}\\
 \end{table}
@@
 
-\subsection{Predicting political views}
-%Dieser Absatz hat ja nicht wirklich was mit political bias zu tun. Wollen wir den also aus K\"uerzungsgr\"unden rausnehmen?
-Parties change their policies and positions in the political spectrum.
-More reliable categories for political bias are party independent labels for policy issues, see \autoref{sec:data}.
-A separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:model}.
-As there was no out-of-domain evaluation set available in this setting only the evaluation error on in-domain data is reported.
-Note however that in this experiment too the evaluation data was never seen by any model during training time.
-In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes.
-Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels. \\
-
-\begin{table}[t]
-\caption{
-\label{tab:results_avg_political_view}
-Classification performance of 56 political views, see \autoref{sec:data}.
-}
-\begin{center}
-\begin{tabular}{lcrrrr}
-code & meaning  &      precision    &recall &  f1-score  & N\\
-\hline\hline
 %
-       501  & environmentalism + &      0.62   &   0.61 &     0.61  &     165\\
-        202 &   democracy + &   0.58  &    0.55   &   0.57   &    122\\
-        701    & labour +&  0.57  &    0.54   &   0.56      & 129\\
-                201    &freedom/human rights +  & 0.58   &   0.54   &   0.56   &    159\\
-         106   & peace + & 0.52&      0.57   &   0.55    &    21\\
-\dots\\
-        302     &centralism + & 0.25     & 0.20  &    0.22  &      10\\
-        401    &  free enterprise + &0.20    &  0.19    &  0.20   &     52\\
-        505    &welfare - &  0.13   &   0.14   &   0.14   &     14\\
-        409    & keynesian demand +&  0.14  &    0.12  &    0.13   &      8\\
-            0    & undefined &  0.09  &    0.12   &   0.10    &    17\\
-            \hline
-avg / total  &  &  0.47    &  0.46 &     0.46 &     2946\\
-%
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
-We now know that party affiliation can be predicted using our model.
+%\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
+%We now know that party affiliation can be predicted using our model.
 But which words are most important for the classification performed here?
@@
 Some unspecific stopwords are excluded.
-The following paragraphs highlight some examples of words that appear to be preferentially used or avoided by each respective party.
+%The following paragraphs highlight some examples of words that appear to be preferentially used or avoided by each respective party.
 Even though interpretations of these results are problematic in that they neglect the context in which these words were mentioned some interesting patterns can be found and related to the actual policies the parties are promoting.
-\paragraph{\bf Left party (linke)}
-The Left party talks mainly about issues related to work ({\em Besch\"aftigten}) and unemployment ({\em Hartz IV, Erwerbslosen}). 
-\paragraph{\bf Green party (gruene)}
-The Green party talks about the formal procedures available to opposition parties to question government policy ({\em fragen, anfragen}) and, interestingly, about the reform of the Germany army ({\em R\"ustungsprojekte, Wehrbericht}).
-\paragraph{\bf Social Democratic Party (SPD)}
-The SPD often uses words related to rights of the working class ({\em International Labour Organisation, Arbeitnehmerrechte}) and the governing coalition ({\em bundestagsfraktion, koalitionspartner}).
-
-\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
-The CDU/CSU party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=2.8cm]{images/party_word_correlations-linke-17.pdf}
-\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-17.pdf}
-\includegraphics[width=3cm]{images/party_word_correlations-spd-17.pdf}
-\includegraphics[width=3cm]{images/party_word_correlations-cducsu-17.pdf}
+%% Vielleicht sollten wir hier dann auch Beispiel aus der 17 nehmen.
+%\paragraph{\bf Left party (linke)}
+%The Left party talks mainly about issues related to work ({\em Besch\"aftigten}) and unemployment ({\em Hartz IV, Erwerbslosen}). 
+%\paragraph{\bf Green party (gruene)}
+%The Green party talks about the formal procedures available to opposition parties to question government policy ({\em fragen, anfragen}) and, interestingly, about the reform of the Germany army ({\em R\"ustungsprojekte, Wehrbericht}).
+%\paragraph{\bf Social Democratic Party (SPD)}
+%The SPD often uses words related to rights of the working class ({\em International Labour Organisation, Arbeitnehmerrechte}) and the governing coalition ({\em bundestagsfraktion, koalitionspartner}).
 %
-\end{center}
-\caption{
-\label{fig:party_word_correlations}
-Correlations between words and party affiliation label for parliament speeches can help interpreting the features used by a predictive model.
+%\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+%The CDU/CSU party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=2.8cm]{images/party_word_correlations-linke-18.pdf}
+%\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-18.pdf}
+%\includegraphics[width=3cm]{images/party_word_correlations-spd-18.pdf}
+%\includegraphics[width=3cm]{images/party_word_correlations-cducsu-18.pdf}
+%%
+%\end{center}
+%\caption{
+%\label{fig:party_word_correlations}
+%Correlations between words and party affiliation label for parliament speeches can help interpreting the features used by a predictive model.
 Shown are the top 10 positively and negatively correlated text features for the current Bundestag.
 For interpretations see \autoref{sec:word_party_correlations}.}
-\end{figure}
+%\end{figure}
 
@@
 Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag.
-In \autoref{tab:results_binary_17} and \autoref{tab:results_binary_18} the results are shown for the 17th and the 18th Bundestag, respectively.
+In \autoref{tab:results_binary_17} the results are shown for the 17th legislative period.
 While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
@@
 While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
-\paragraph{Speech sentiment correlates with political power} In order to investigate the features that give rise to the classifiers' performance the bag-of-words features were analysed with respect to their sentiment.
+\\
+
+\paragraph{Speech sentiment correlates with political power} 
+But what are the underlying features that give rise to the classifiers performance?
+In order to investigate this the bag-of-words features were analysed with respect to their sentiment.
 The average sentiment of each political party is shown in \autoref{fig:party_sentiments}.
@@
 Political power was evaluated in terms of membership of the government.
-Correlating either of these two indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
+Correlating these indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
 This pattern is evident from the data in \autoref{fig:party_sentiments} and in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
@@
 \begin{center}
-%\includegraphics[width=10cm]{images/fipi-screenshot}
+\includegraphics[width=10cm]{images/fipi-screenshot}
 %

### 2016-06-13 manuscript/poltext.tex
Changed style from ICML to the ACL style file that the organisers sent around. Shortened text.

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%% ICML 2015 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% File poltext2016.tex
+%
+% Contact: jan.snajder@fer.hr or dsirinic@gmail.com
+%%
+%% Based on the style files for ACL-2015, which were, in turn,
+%% Based on the style files for ACL-2014, which were, in turn,
+%% Based on the style files for ACL-2013, which were, in turn,
+%% Based on the style files for ACL-2012, which were, in turn,
+%% based on the style files for ACL-2011, which were, in turn, 
+%% based on the style files for ACL-2010, which were, in turn, 
+%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
+%% based on the style files for EACL-2009 and IJCNLP-2008...
 
-% Use the following line _only_ if you're still using LaTeX 2.09.
-%\documentstyle[icml2015,epsf,natbib]{article}
-% If you rely on Latex2e packages, like most moden people use this:
-\documentclass{article}
+%% Based on the style files for EACL 2006 by 
+%%e.agirre@ehu.es or Sergi.Balari@uab.es
+%% and that of ACL 08 by Joakim Nivre and Noah Smith
 
-% use Times
+\documentclass[11pt]{article}
+\usepackage{poltext2016}
 \usepackage{times}
-% For figures
+\usepackage{url}
+\usepackage{latexsym}
+\usepackage{booktabs}
 \usepackage{graphicx} % more modern
@@
 \usepackage{pbox}
+\usepackage{color}
 
@@
 
-% Employ the following version of the ``usepackage'' statement for
-% submitting the draft version of the paper for review.  This will set
-% the note in the first column to ``Under review.  Do not distribute.''
-\usepackage{icml2015}
+\newcommand{\pola}[1]{\textcolor{blue}{#1}}
+\newcommand{\felix}[1]{\textcolor{green}{#1}}
 
-% Employ this version of the ``usepackage'' statement after the paper has
-% been accepted, when creating the final version.  This will set the
-% note in the first column to ``Proceedings of the...''
-%\usepackage[accepted]{icml2015}
+%\setlength\titlebox{5cm}
 
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Automating Political Bias Prediction}
+% You can expand the titlebox if you need extra space
+% to show all the authors. Please do not make the titlebox
+% smaller than 5cm (the original size); we will check this
+% in the camera-ready version and ask you to change it back.
+
+\title{Automating Political Bias Prediction}
+
+\author{Felix Biessmann\\
+  {\tt felix.biessmann@gmail.com} \\\And
+ Pola Lehmann \\
+  Wissenschaftszentrum Berlin \\
+  {\tt pola.lehmann@wzb.eu} \\}
+
+\date{}
 
 \begin{document}
-
-\twocolumn[
-\icmltitle{Automating Political Bias Prediction}
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2015
-% package.
-\icmlauthor{Felix Biessmann}{felix.biessmann@gmail.com}
-\icmladdress{}
-\icmlauthor{Pola Lehmann}{pola.lehmann@wzb.eu}
-\icmladdress{Wissenschaftszentrum Berlin}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{}
-
-\vskip 0.3in
-]
+\maketitle
 
@@
 The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
-We will test our model on both in-domain and out-of-domain data.
-Prediction on out-of-domain data is less precise as on in-domain data, but still generating satisfying results.
+We will test our model on both in-domain and out-of-domain data; the different train and test data sets are summarized in \autoref{tab:classification_settings}.
+The classification tasks we considered as proxies for political bias prediction are
+
+\begin{enumerate}
+\item {\bf Classification of party affiliation} (five class problem)
+\item {\bf Classification of government membership} (binary problem)
+\item {\bf Classification of policy issues} (56 class problem)
+\end{enumerate}
 
@@
 Third sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
+
+% Confusion matrix topic level
+\begin{table}[t]\label{tab:classification_settings}
+\footnotesize
+\begin{tabular}{ccc}
+Training Set & \multicolumn{2}{c}{Test set} \\
+& In-domain&Out-of-domain\\
+\hline
+\hline
+Parliament ($90\%$) & Parliament ($10\%$)& Manifestos\\
+Manifestos ($90\%$)& Manifestos ($10\%$) & -\\
+\end{tabular}
+\end{table}
 
@@
 \subsection{Data}
-Annotated political text data was obtained from two sources: a) the plenary debates held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the current 18th and the last, 17th, legislative period.
+Annotated political text data was obtained from two sources: a) the plenary debates held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the the last, 17th, legislative period.
 
@@
 The protocols of plenary debates are available through the website of the German Bundestag\footnote{\url{https://www.bundestag.de/protokolle}}; an open source API was used to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
-In total 22784 speeches were extracted for the 17th legislative period and 11317 speeches for the 18th period, queried until March 2016.
+In total 22784 speeches were extracted for the 17th legislative period.
 
 \paragraph{Party manifesto data}
-The party manifesto text was taken from the Manifesto Corpus (add citation).
-The Corpus is accessible through an publicly available API published by the {\em Manifesto Project} \cite{manifesto}.
+The party manifesto text was taken from the Manifesto Corpus  \cite{manifesto}.
 The data released in this project mainly comprises the complete manifestos of all parties that have won seats at a national election.
@@
 Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
-In total the 9 manifestos from the 17th and 18th legislative period have been divided into 29451 quasi-sentences.
-Each observation has been added with two labels: the party affiliation and the manually assigned policy issue aimed at in each specific statement.
+Each quasi-sentence has  two types of labels: the party affiliation and the manually assigned policy issue aimed at in each specific statement.
 
@@
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-First each data set was segmented into semantic units; in the case of the party manifesto data the semantic units were the quasi-sentences associated with one of the 56 policy issues, in the case of parliament discussions this were the speeches.
-Parliament speeches were often interrupted; in this case each uninterrupted part of a speech was considered a semantic unit.
-Strings of each semantic unit were tokenised and transformed into bag-of-word vectors as implemented in scikit-learn \cite{scikit-learn}.
-The general idea of bag-of-words vectors is to simply count occurrences of words (or word sequences, also called {\em n-grams}) for each data point.
-A data point is usually a document, here it is the semantic units of parliament speeches and manifesto sentences, respectively.
-The text of each semantic unit is transformed into a vector $\vec{x}\in\mathds{R}^d$ where $d$ is the size of the dictionary; the $w$th entry of $\vec{x}$ contains the (normalized) count of the $w$th word (or sequence of words) in our dictionary.
-Several options for vectorizing the speeches were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
+All text data was segmented into semantic units, quasi-sentences for manifesto data and (uninterrupted) speech parts for parliament speech data.
+Strings of each semantic unit were tokenised and transformed into bag-of-word (BOW) vectors as implemented in scikit-learn \cite{scikit-learn}.
+Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
 All of these hyperparameters were subjected to hyperparameter optimization as explained in \autoref{sec:crossvalidation}.
@@
 
-Three different classification problems were considered:
-\begin{enumerate}
-\item {\bf Classification of party affiliation} (five class / four class problem)
-\item {\bf Classification of government membership} (binary problem)
-\item {\bf Classification of policy issues} (56 class problem)
-\end{enumerate}
-
-Party affiliation is a five class problem for the 17th legislation period, and a four class problem for the 18th legislation period.
-The policy issue classification is based on the categories of the Manifesto Project, see \autoref{sec:data} and \cite{leftright}.
-For both the party affiliation and the government membership prediction, classifiers were trained on the parliament speeches.
-For the third problem classifiers were trained only on the manifesto data for which policy issue labels were available.
-
 \subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
 The model pipeline contained a number of  hyperparameters that were optimised using cross-validation.
-We first split the training data into a training data set that was used for optimisation of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data; wherever possible the generalisation performance of the models was also evaluated on out-of domain data.
-Hyperparameters were optimised using grid search and 3-fold cross-validation within the training set only: A cross-validation split was made to obtain train/test data for the grid search and for each setting of hyperparameters the entire pipeline was trained and evaluated -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimisation.
+We first split the training data into a training data set that was used for optimisation of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data.
+Hyperparameters were optimised using grid search and 3-fold cross-validation within the training set -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimisation.
 For the best setting of all hyperparameters the pipeline was trained again on all training data and evaluated on the evaluation data sets.
 For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a given legislative period.
-Out-of-domain evaluation data were the texts from party manifestos.
+Wherever possible the generalisation performance of the models was also evaluated on out-of domain texts, for which we used the texts from party manifestos.
 For the policy issue prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
@@
 
-\subsection{Analysis of bag-of-words features}\label{sec:correlations_methods}
-While interpretability of linear models is often propagated as one of their main advantages, doing so naively without modelling the noise covariances can lead to wrong conclusions, see e.g.
-\cite{Zien2009, Haufe2013}; interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
-Thus direct interpretation of the model coefficients $\vec{W}$ is problematic.
+\subsection{Interpreting bag-of-words models}\label{sec:correlations_methods}
+Interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
+Thus direct interpretation of the model coefficients $\vec{W}$ is problematic, see also \cite{Zien2009, Haufe2013}.
 In order to allow for better interpretation of the predictions and to assess which features are discriminative correlation coefficients between each word and the party affiliation label were computed.
@@
 When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
-These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament; the accuracy for the 17th Bundestag is 0.63, results of the 18th Bundestag are difficult to compare as the number of parties is four and the legislation period is not finished yet.
-
-But as the main purpose of this paper is to generate a model that can be used to predict political bias on new data, like media texts, we are interested in the question if the model also produces satisfying results on out-of-domain data.
-We are therefore also predicting party affiliation on party manifesto text.
+These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.
 For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
@@
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into the previously described party domains, see \autoref{sec:data}.
+In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into the previously described party domains.
 The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
@@
 \label{tab:results_out-of-domain}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on out-of-domain data.  Predictions are done on the {\bf sentence level}.
+Classification performance on the party affiliation prediction problem for data from the 17th legislative period on out-of-domain data.
+Predictions are done on the {\bf sentence level}.\label{tab:confusion}
 }
@@
 {\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
-In contrast to single sentence level predictions (see \autoref{tab:results_17}, \autoref{tab:results_18}, \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.
-Note that all manifesto topics of the green party in the 18th Bundestag are predicted to be from the parties of the governing coalition, CDU/CSU or SPD.}
+In contrast to single sentence level predictions (see  also \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.}
 \begin{center}
 \begin{tabular}{lcccc}
-& \multicolumn{4}{c}{\bf Party Manifestos} \\
-& \multicolumn{4}{c}{\bf 17th Bundestag} \\
     &         precision    &recall &  f1-score  & N  \\
@@
 In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes.
-Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels. Results are not surprising, but are equal to those by human coders. \\
+Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels.
+\felix{Results are not surprising, but are equal to those by human coders.} \\
 
@@
         701    & labour +&  0.57  &    0.54   &   129\\
-                201    &freedom/human rights +  & 0.58   &   0.54   &      159\\
+                201    &human rights +  & 0.58   &   0.54   &      159\\
          106   & peace + & 0.52&      0.57   &       21\\
@@
         505    &welfare - &  0.13   &   0.14   &      14\\
-        409    & keynesian demand +&  0.14  &      0.13   &      8\\
+        409    & keyn. demand +&  0.14  &      0.13   &      8\\
             0    & undefined &  0.09  &    0.12   &    17\\
             \hline
-avg / total  &  &  0.47    &  0.46 &    2946\\
+ total  &  &  0.47    &  0.46 &    2946\\
 %
@@
 In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
-On the topic level results for the 17th legislative period are pretty good except for the SPD.
-In the 18th legislative period predictions are very bad for the Green party.
-One explanation for this missclassification could lie in policy changes.
-For example, the Green party has devoted large parts of its manifesto to renewable and against nuclear energy.
-After Fukushima the other parties adopted these positions.
+On the topic level results for the 17th legislative period are pretty good except for the SPD, which the classifier often predicts as CDU/CSU.
+One explanation for this missclassification could lie in policy changes, due to e.g.
+the coalition contract that forces the SPD to eventually accept compromises with the other governing parties, CDU/CSU.
 \\
-
-% Die könnten wir vielleicht rausnehmen.
-% Confusion matrix sentence level
-\begin{table}[t]\label{tab:conf_mat_four_class}
-\caption{\label{tab:confusion} {\bf Confusion matrices (sentence level)} for predictions on evaluation data (party manifestos); classifiers were trained on parliament speeches for the 17th legislative period; the most prominent effect is the high likelihood for a party to be taken as the strongest, governing party, cdu/csu.
-This can be interpreted as a change in policies of the conservative party cdu/csu towards the policies of the green party.}
-\begin{tabular}{lccccccc}
- \multicolumn{8}{c}{\bf Party Manifestos}\\
- \multicolumn{8}{c}{\bf 17th Bundestag}\\
- \\
-&&& \multicolumn{5}{c}{Predicted}\\
-&&& cducsu & fdp& gruene& linke& spd\\
-\hline
-\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}}& &cducsu &1186 &289& 178& 198& 179\\
-&&fdp &882& 658& 236& 329& 214\\
-&&gruene &1174& 404& 764& 941& 464\\
-&&linke &388& 92& 214& 806& 201\\
-&&spd &999& 268& 240& 398& 373\\
-\end{tabular}
-\end{table}
 
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
-\caption{\label{tab:confusion_topic} {\bf Confusion matrices (topic level)} for predictions on evaluation data (party manifestos) for classifiers trained on parliament speeches for the 17th legislative period (left) and 18th legislative period (right).}
-\begin{tabular}{lccccccc}
- \multicolumn{8}{c}{\bf Party Manifestos}\\
- \multicolumn{8}{c}{\bf 17th Bundestag}\\
- \\
-&&& \multicolumn{5}{c}{Predicted}\\
-&&& cducsu & fdp& gruene& linke& spd\\
+\caption{\label{tab:confusion_topic} {\bf Topic level confusion matrices} on evaluation data (manifesto texts), classifiers were trained on speeches of 17th Bundestag.}
+\vspace{0.5em}
+\footnotesize
+\begin{tabular}{lcccccc}
+&& \multicolumn{5}{c}{\bf Predicted}\\
+&& cducsu & fdp& gruene& linke& spd\\
 \hline
-\multirow{5}{*}{\rotatebox{90}{\pbox{3cm}{\centering True}}} &&cducsu &7& 0& 0& 0& 0\\
-&&fdp&0& 7& 0& 0& 0\\
-&&gruene&0& 0& 6& 0& 1\\
-&&linke&0& 0& 0& 7& 0\\
-&&spd&4& 0& 0& 0& 4\\
+\hline
+\multirow{5}{*}{\rotatebox{90}{\pbox{2cm}{\centering {\bf True}}}} &cducsu &7& 0& 0& 0& 0\\
+&fdp&0& 7& 0& 0& 0\\
+&gruene&0& 0& 6& 0& 1\\
+&linke&0& 0& 0& 7& 0\\
+&spd&4& 0& 0& 0& 4\\
 \end{tabular}
@@
 \end{table*}
-
-\begin{table*}[t]
-\caption{
-\label{tab:results_binary_18}
-Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
-}
-\begin{center}
-\begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
-    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
-\hline \hline
-government   &    0.88  &    0.95    &  0.92   &    786   &0.52   &   0.66 &     0.58  &    5972\\
- opposition    &   0.86    &  0.71   &   0.78   &    346 & 0.69  &    0.56    &  0.62   &   8229\\
-\hline
-avg / total    &   0.88    &  0.88     & 0.87  &    1132 &  0.62   &   0.60    &  0.60 &    14201\\
-%
-\end{tabular}
-\end{center}
-\end{table*}
+%%
+%%\begin{table*}[t]
+%%\caption{
+%%\label{tab:results_binary_18}
+%%Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
+%%}
+%%\begin{center}
+%%\begin{tabular}{lcccccccc}
+%%& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
+%%    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
+%%\hline \hline
+%%government   &    0.88  &    0.95    &  0.92   &    786   &0.52   &   0.66 &     0.58  &    5972\\
+%% opposition    &   0.86    &  0.71   &   0.78   &    346 & 0.69  &    0.56    &  0.62   &   8229\\
+%%\hline
+%%avg / total    &   0.88    &  0.88     & 0.87  &    1132 &  0.62   &   0.60    &  0.60 &    14201\\
+%%%
+%%\end{tabular}
+%%\end{center}
+%%\end{table*}
 
@@
 This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
-The authors report classification accuracies between 0.8 and 0.87, the accuracy in the 17th Bundestag was 0.85.
+The authors report classification accuracies between 0.80 and 0.87, the accuracy in the 17th Bundestag was 0.85.
 While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
@@
 Correlating these indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
-This pattern is evident from the data in \autoref{fig:party_sentiments} and in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+This pattern is evident from the data in
+%\autoref{fig:party_sentiments} and 
+in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
 
-Note that there is one party, the social democrats (SPD), which has many seats and switched from opposition to government with the 18th Bundestag: With its participation in the government the average sentiment of this party switched sign from negative to positive, suggesting that positive sentiment is a strong indicator of government membership.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf}
+%Note that there is one party, the social democrats (SPD), which has many seats and switched from opposition to government with the 18th Bundestag: With its participation in the government the average sentiment of this party switched sign from negative to positive, suggesting that positive sentiment is a strong indicator of government membership.
 %
-\end{center}
-\caption{
-\label{fig:party_sentiments}
-Speech sentiments computed for speeches of each party; parties are ordered according to the number of seats in the parliament.
+%
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf}
+%%
+%\end{center}
+%\caption{
+%\label{fig:party_sentiments}
+%Speech sentiments computed for speeches of each party; parties are ordered according to the number of seats in the parliament.
 There is a trend for more positive speech content with more political power.
 Note that the SPD (red) switched from opposition to government in the 18th Bundestag: their seats in the parliament increased and the average sentiment of their speeches switched sign from negative to overall positive sentiment.
-}
-\end{figure}
+%}
+%\end{figure}
 
@@
 Within each topic it is then possible to get an ordered (from left to right) overview of the articles on that topic.
-An example of one topic that emerged on March 31st is shown in \autoref{fig:fipi}.
 A preliminary demo is live at \cite{fipidemo} and the code is available on github\cite{fipi}.
-\begin{figure}
-\begin{center}
-\includegraphics[width=10cm]{images/fipi-screenshot}
-%
-\end{center}
-\caption{
-\label{fig:fipi}
-A screen shot of an example web application using the political view prediction combined with topic modelling to provide a heterogeneous overview of a topic. }
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=10cm]{images/fipi-screenshot}
+%%
+%\end{center}
+%\caption{
+%\label{fig:fipi}
+%A screen shot of an example web application using the political view prediction combined with topic modelling to provide a heterogeneous overview of a topic. }
+%\end{figure}
 
@@
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was
-% created by Lise Getoor and Tobias Scheffer, it was slightly modified
-% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz,
-% slightly modified from the 2009 version by Kiri Wagstaff and
-% Sam Roweis's 2008 version, which is slightly modified from
-% Prasad Tadepalli's 2007 version which is a lightly
-% changed version of the previous year's version by Andrew Moore,
-% which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2016-06-13 manuscript/poltext.tex
Changed abstract to match what we actually submitted. removed more text.

@@
 
-\title{Automating Political Bias Prediction}
+\title{Predicting political party affiliation from text}
 
@@
 \begin{abstract}
-Every day media generate large amounts of text.
-An unbiased view on media reports requires an understanding of the political bias of media content.
-Assistive technology for estimating the political bias of texts can be helpful in this context.
-This study proposes a simple statistical learning approach to predict political bias from text.
-Standard text features extracted from speeches and manifestos of political parties are used to predict political bias in terms of political party affiliation and in terms of political views.
-Results indicate that political bias can be predicted with above chance accuracy.
-Mistakes of the model can be interpreted with respect to changes of policies of political actors.
-Two approaches are presented to make the results more interpretable: a) discriminative text features are related to the political orientation of a party and b) sentiment features of texts are correlated with a measure of political power.
-Political power appears to be strongly correlated with positive sentiment of a text.
-To highlight some potential use cases a web application shows how the model can be used for texts for which the political bias is not clear such as news articles.
+Every day large amounts of text are produced in public discourse.
+Some of this text is produced by actors whose political colour is very obvious, for example in the case of party elites.
+But there are also many actors raising their voices in the discourse, who cannot be clearly attributed to a political party, while their statements might actually by biased towards a specific party.
+Identifying such biases is crucial for political research as well as media consumers, especially when analysing the influence of the media on political discourse and vice versa.
+In this study we investigate to what extent political party affiliation can be predicted from textual content.
+Results indicate that automatic classification of political affiliation is possible with above chance accuracy also across text domains.
+We propose methods to better interpret these results and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
+We present an example of a web application that combines topic modelling with automatic political affiliation analyses of news articles based on the proposed model.
 \end{abstract}
@@
 \begin{enumerate}
-\item {\bf Classification of party affiliation} (five class problem)
+\item {\bf Classification of party affiliation}\\ (five class problem for 17th Bundestag)
 \item {\bf Classification of government membership} (binary problem)
-\item {\bf Classification of policy issues} (56 class problem)
+\item {\bf Classification of policy issues}\\ (56 class problem)
 \end{enumerate}
@@
 Third sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
-
-% Confusion matrix topic level
-\begin{table}[t]\label{tab:classification_settings}
-\footnotesize
-\begin{tabular}{ccc}
-Training Set & \multicolumn{2}{c}{Test set} \\
-& In-domain&Out-of-domain\\
-\hline
-\hline
-Parliament ($90\%$) & Parliament ($10\%$)& Manifestos\\
-Manifestos ($90\%$)& Manifestos ($10\%$) & -\\
-\end{tabular}
-\end{table}
 
@@
 \begin{eqnarray}\label{eq:logreg_multiclass}
-p(y=k|\vec{x},\vec{W}) = &\frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}} \qquad \textrm{with }  z_k=&\vec{w}_k^{\top}\vec{x} \\\nonumber
+p(y=k|\vec{x},\vec{W}) = \frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}}  \textrm{ with }  z_k=\vec{w}_k^{\top}\vec{x}.
 \end{eqnarray}
-The performance of the model was optimized using the classification accuracy, but we also report all other standard measures, precision ($TP / (FP + TP$), recall ($TP / (TP + FN)$) and f1-score ($2\times (Prec.
-\times Rec) / (Prec + Rec.)$).
-\\
 
@@
 For the policy issue prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
+The train and test splits are summarized in \autoref{tab:classification_settings}.
 % Falls wir die Medien nochreinnehmen, muss das bei out-of-domain auch noch erwähnt werden.
+% 
+\begin{table}[t]\label{tab:classification_settings}
+\footnotesize
+\begin{tabular}{ccc}
+Training Set & \multicolumn{2}{c}{Test set} \\
+& In-domain&Out-of-domain\\
+\hline
+\hline
+Parliament ($90\%$) & Parliament ($10\%$)& Manifestos\\
+Manifestos ($90\%$)& Manifestos ($10\%$) & -\\
+\end{tabular}
+\end{table}
 
@@
 A sentiment vector $\vec{s}\in\R^d$ was constructed from the sentiment polarity values in the sentiment dictionary.
-The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors $\vec{x}\in\R^d$ and $\vec{s}$
-
-\begin{align}
-\frac{\vec{s}^\top \vec{x}}{\|\vec{s}\|\|\vec{x}\|}
-\end{align}
+The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors and sentiment vectors.
 
@@
 \label{tab:results_in-domain}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on in-domain data. $N$ denotes number of data points in the evaluation set.
+{\bf In-domain classification performance} for data from the 17th legislative period on in-domain data. $N$ denotes number of data points in the evaluation set.
 }
@@
 \label{tab:results_out-of-domain}
-Classification performance on the party affiliation prediction problem for data from the 17th legislative period on out-of-domain data.
-Predictions are done on the {\bf sentence level}.\label{tab:confusion}
+{\bf Out-of-domain classification performance} (quasi-sentence level) of a classifier trained on speeches of the 17th legislative period of the Bundestag.
 }
@@
 \label{tab:results_topic}
-{\bf Topic level classification performance} on the party affiliation prediction problem for data from the evaluation set (manifesto texts) of the 17th legislative period.
-In contrast to single sentence level predictions (see  also \autoref{tab:confusion} for results and \autoref{sec:data} for topic definitions) the predictions made on topic level are reliable in many cases.}
+{\bf Out-of-domain classification performance} (topic level) on manifesto data of a classifier trained on the 17th legislative period of the Bundestag.
+Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}) the predictions made on topic level are more reliable.}
 \begin{center}

### 2016-06-14 bundestag.py
Added fb experiment.

@@ -20,6 +20,7 @@ import pandas as pd
 from scipy import random,unique,vstack,hstack,arange,array,mean,corrcoef
 import cPickle
 from sentiments import SentimentClassifier
+from fb import fbTraverserTop
 
 DATA_PATH = os.environ.get('DATA_PATH', 'data')
 TXT_DIR = os.path.join(DATA_PATH, 'txt')
@@ -200,6 +201,18 @@ def list_top_words(legis=17,topwhat=10):
 def newsFileTuple(nf):
     return [(x['text'],x['source']) for x in json.load(open(nf))]
 
+def get_raw_text_fb(folder="data/parteien-auf-fb", suffix=".json.gz"):
+    '''
+    Loads raw text and labels from news article files
+    (Downloaded using newsreader.get_daily_news())
+    '''
+    files = glob.glob(folder+"/*"+suffix)
+    return zip(*chain(*filter(None,map(fbFileTuple,files))))
+
+def fbFileTuple(fbf):
+    f = [ff[1] for ff in fbTraverserTop(fbf)]
+    return zip(f, [fbf.split("/")[-1].split(".")[0]] * len(f))
+
 def get_raw_text_news(folder="data", prefix="news-2016"):
     '''
     Loads raw text and labels from news article files
@@ -321,8 +334,8 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     parameters = {'vect__ngram_range': [(1,2)],\
            #'vect__stop_words':(None,stops),\
            #'tfidf__use_idf': (True,False),\
-           'clf__C': (10.**sp.arange(1,5,1.)).tolist(),
-           'vect__max_df':[.3, .5, .75],
+           'clf__C': (10.**sp.arange(-4,4,2.)).tolist(),
+           'vect__max_df':[.5, .75],
         }
     saveId = idstr+"-"+randid()
     X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
@@ -381,6 +394,11 @@ def classify_speeches_party_parliament(legislationPeriod = 18):
     trainData, trainLabels = get_raw_text_bundestag(legislationPeriod=legislationPeriod)
     evalData, evalLabels = get_raw_text(legislationPeriod=legislationPeriod)
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="parliament-train-%d"%legislationPeriod)
+
[…]

### 2016-06-14 fb.py

@@ -1,0 +1,125 @@
+import json,gzip
+from itertools import chain
+import cPickle
+from sklearn.feature_extraction.text import CountVectorizer
+from sklearn.pipeline import Pipeline
+from sklearn.linear_model import LogisticRegression
+from sklearn.grid_search import GridSearchCV
+from sklearn import cross_validation,metrics
+import scipy as sp
+import re
+
+DDIR = '/Users/felix/Data/fb-politics/'
+
+dat = [DDIR + x + ".json.gz" for x in ['afd','npd','pegida']]
+
+urlPat = r'(http://.*\.html)'
+
+def getUrls(post): 
+    urls = []
+    if post['message']: 
+        foundUrl = re.search(urlPat,post['message'])
+        if foundUrl:
+            name = ''
+            import pdb;pdb.set_trace()
+            if post.has_key('name'):
+                name += post['name']
+            urls.append((post['created'],post['id']+": "+name,foundUrl.groups(0)[0]))
+    if len(post['comments'])>0: 
+        urls += list(chain(*map(getUrls,post['comments'])))
+    return filter(lambda x: len(x)>0,urls)
+
+def getText(post): 
+    text = []
+    if post['message'] and len(post['message'])>0: 
+        text.append((post['created'][:10],post['message']))
+    if len(post['comments'])>0: 
+        text = text + list(chain(*map(getText,post['comments'])))
+    return text
+
+def getTextTop(post): 
+    text = []
+    if post['message'] and len(post['message'])>0: 
+        text.append((post['created'][:10],post['message']))
+    return text
+
+def train_bow(d = dat):
+    vect = CountVectorizer().fit(map(lambda x: x[1], chain(*map(fbTraverser, d))))
+    cPickle.dump(vect,open(DDIR+"vectorizer.pickle","wb"))
+
+def get_daily_bow_single(data, vect):
+    daily_bow = {}
+    for day,text in fbTraverser(data):
+        if daily_bow.has_key(day):
+            daily_bow[day] += vect.transform([text]) 
+        else: daily_bow[day] = vect.transform([text])
+    return daily_bow
+
+def get_daily_bow(data = dat):
+    vect = cPickle.load(open(DDIR+"vectorizer.pickle"))
+    daily_bow = {k:get_daily_bow_single(k,vect) for k in data}
+    cPickle.dump(daily_bow,open(DDIR+"daily_bow.pickle","wb"))
+
[…]

### 2016-06-15 manuscript/poltext.tex
Shortened text further.

@@
 Standard newspapers, but also new forms as twitter, facebook or blogs produce an ever growing amount of text influencing political debate.
-These texts can be of interest to political scientist in various analyses.
+These texts can be of interest to political scientists in various analyses.
 The question we will focus on in this analysis is the potential political bias that can be detected in these sources.
@@
 It gives us the possibility to get a more nuanced view of the spectrum of political positions favored in important news outlets in a country.
-In this paper we will test whether it is possible to automatically detect political bias in text and if such a model can be used on different kind of texts.
-The aim of this study is to provide some empirical evidence indicating that leveraging open data sources automated political bias prediction is possible with above chance accuracy.
-We will test our model on both in-domain and out-of-domain data; the different train and test data sets are summarized in \autoref{tab:classification_settings}.
-The classification tasks we considered as proxies for political bias prediction are
-
-\begin{enumerate}
-\item {\bf Classification of party affiliation}\\ (five class problem for 17th Bundestag)
-\item {\bf Classification of government membership} (binary problem)
-\item {\bf Classification of policy issues}\\ (56 class problem)
-\end{enumerate}
-
+In this paper we test whether it is possible to automatically detect political party affiliation in text and if such a model can be used on different kinds of texts that it was trained on.
 In order to validate and explain the predictions of the models three strategies that allow for better interpretations of the models are proposed.
@@
 
-%For the 18th period we also added the manifesto data of three additional parties that did not win seats but nonetheless played a very important role in the political discourse in the run-up to the election.
-Wollen wir die einfach rauslassen?
-Käme ja eh nur in einem kleinen Teil der Analyse vor und macht die Beschreibung an mehreren Stellen kürzer.
-
 \paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
@@
 
-%The party affiliation was used to evaluate the party evaluation classifiers trained on the parliament speeches.
-For this purpose the data acquisition was constrained to only those parties that were elected into the parliament.
-For the analyses based on policy issues all parties were considered, also those three that did not make it into parliament.
-
 The length of each annotated statement in the party manifestos is rather short.
@@
 In the Manifesto Project the human coders therefore do not code these statements separately but within context.
-In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains using the 56policy categories from the Manifesto Project.
+In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains using the 56 policy categories from the Manifesto Project.
 Within the Manifesto Project category scheme each of these belongs to one of seven policy domains.
@@
 The train and test splits are summarized in \autoref{tab:classification_settings}.
-% Falls wir die Medien nochreinnehmen, muss das bei out-of-domain auch noch erwähnt werden.
 % 
@@
 \end{table}
-
+%
 \subsection{Sentiment analysis}\label{sec:sentiment_analysis_methods}
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section.
-Because of space restrictions, we will only report results fromt he 17th legislative period here.
-Results are very similar for the 18th legislative period and can be accessed via github.
+%Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section. 
+Due to space restrictions we only report results from the 17th legislative period. Results are similar for the 18th legislative period.
 
@@
 This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
+\paragraph{Length of texts influences prediction accuracy}
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into the previously described party domains.
+In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics.
 The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
@@
 \hline
-avg / total &      0.64   &   0.63   &   0.62    &  2279 
+ total &      0.64   &   0.63   &   0.62    &  2279 
 %
@@
 }
+
 \begin{center}
 \begin{tabular}{lcccc}
-    &         precision    &recall &  f1-score  & N  \\
+    &         prec.    &recall &  f1-score  & N  \\
 \hline \hline
@@
 \hline
-avg / total    &   0.35  &    0.31  &    0.30   &   12075\\
+total    &   0.35  &    0.31  &    0.30   &   12075\\
 %
@@
     \hline
-    avg / total  &     0.88   &   0.86   &   0.86  &      36\\
+total  &     0.88   &   0.86   &   0.86  &      36\\
 \end{tabular}
@@
 
-\subsection{Validation}
-\paragraph{Differentiability of policy issues}
-One reason why some statements are difficult to classify could be that some issues are more difficult to differentiate than others.
-To test this a separate suite of experiments was run to train and test the prediction performance of the text classifiers models described in \autoref{sec:model}.
-As there was no out-of-domain evaluation set available in this setting only the evaluation error on in-domain data is reported.
-Note however that in this experiment too the evaluation data was never seen by any model during training time.
-In \autoref{tab:results_avg_political_view} results for the best and worst classes, in terms of predictability, are listed along with the average performance metrics on all classes.
-Precision and recall values of close to 0.5 on average can be considered rather high considering the large number of labels.
-\felix{Results are not surprising, but are equal to those by human coders.} \\
-
-\begin{table}
-\caption{
-\label{tab:results_avg_political_view}
-Classification of 56 political views, see \autoref{sec:data}.
-}
-\begin{center}
-\footnotesize
-\begin{tabular}{lcrrrr}
-code & meaning  &      prec    &recall & N\\
-\hline\hline
-%
-       501  & environmentalism + &      0.62   &   0.61 &      165\\
-        202 &   democracy + &   0.58  &    0.55   &      122\\
-        701    & labour +&  0.57  &    0.54   &   129\\
-                201    &human rights +  & 0.58   &   0.54   &      159\\
-         106   & peace + & 0.52&      0.57   &       21\\
-\dots\\
-        302     &centralism + & 0.25     & 0.20  &          10\\
-        401    &  free enterprise + &0.20    &  0.19    &       52\\
-        505    &welfare - &  0.13   &   0.14   &      14\\
-        409    & keyn. demand +&  0.14  &      0.13   &      8\\
-            0    & undefined &  0.09  &    0.12   &    17\\
-            \hline
- total  &  &  0.47    &  0.46 &    2946\\
-%
-\end{tabular}
-\end{center}
-\end{table}
-
 \paragraph{Policy change}
-Another explanation for missclassification could lie in the fact that parties change their policy positions.
-In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for sentence level predictions (see \autoref{tab:confusion}) as well as topic level predictions (see \autoref{tab:confusion_topic}).
+Another explanation for misclassification could lie in the fact that parties change their policy positions.
+In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for topic level predictions (see \autoref{tab:confusion_topic}).
 On the topic level results for the 17th legislative period are pretty good except for the SPD, which the classifier often predicts as CDU/CSU.
-One explanation for this missclassification could lie in policy changes, due to e.g.
-the coalition contract that forces the SPD to eventually accept compromises with the other governing parties, CDU/CSU.
+One explanation for this misclassification could be policy changes.
 \\
@@
 %
-%\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
-%We now know that party affiliation can be predicted using our model.
-But which words are most important for the classification performed here?
-To get a more precise idea of what differentiates the parties from each other the 10 highest and lowest correlations between individual words and the party affiliation label are shown for each party in \autoref{fig:party_word_correlations}.
-Correlations were computed on the data from the current, 18th, legislative period.
+\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
+To get a more precise idea of what word features are discriminative in terms of the party affiliation label correlations between words and party affilation labels are computed; these are the words that are preferentially used or avoided by each respective party.
 Some unspecific stopwords are excluded.
-%The following paragraphs highlight some examples of words that appear to be preferentially used or avoided by each respective party.
-Even though interpretations of these results are problematic in that they neglect the context in which these words were mentioned some interesting patterns can be found and related to the actual policies the parties are promoting.
-%% Vielleicht sollten wir hier dann auch Beispiel aus der 17 nehmen.
-%\paragraph{\bf Left party (linke)}
-%The Left party talks mainly about issues related to work ({\em Besch\"aftigten}) and unemployment ({\em Hartz IV, Erwerbslosen}). 
-%\paragraph{\bf Green party (gruene)}
-%The Green party talks about the formal procedures available to opposition parties to question government policy ({\em fragen, anfragen}) and, interestingly, about the reform of the Germany army ({\em R\"ustungsprojekte, Wehrbericht}).
-%\paragraph{\bf Social Democratic Party (SPD)}
-%The SPD often uses words related to rights of the working class ({\em International Labour Organisation, Arbeitnehmerrechte}) and the governing coalition ({\em bundestagsfraktion, koalitionspartner}).
-%
-%\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
-%The CDU/CSU party often uses words related to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit}).
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=2.8cm]{images/party_word_correlations-linke-18.pdf}
-%\includegraphics[width=2.9cm]{images/party_word_correlations-gruene-18.pdf}
-%\includegraphics[width=3cm]{images/party_word_correlations-spd-18.pdf}
-%\includegraphics[width=3cm]{images/party_word_correlations-cducsu-18.pdf}
-%%
-%\end{center}
-%\caption{
-%\label{fig:party_word_correlations}
-%Correlations between words and party affiliation label for parliament speeches can help interpreting the features used by a predictive model.
-Shown are the top 10 positively and negatively correlated text features for the current Bundestag.
-For interpretations see \autoref{sec:word_party_correlations}.}
-%\end{figure}
+We list for each party some of the top positively correlated words, sorted in descendent order.
+\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
+\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
+\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em Staatssekretaerin, Kanzlerin, Bundestagsfraktion}) and some words related to cutting of expenses ({\em Kuerzungen}).
+\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit, Stabilitaet}).
 
@@
 The authors report classification accuracies between 0.80 and 0.87, the accuracy in the 17th Bundestag was 0.85.
-While topic-level predictions were not performed in this binary setting, the party affiliation results in \autoref{tab:results_topic} suggest that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
+A pronounced drop in performance was reported when applying the model to texts from a different domain (e.g.
+older texts or texts from another chamber).
+While we did not perform topic-level predictions in this binary setting, the party affiliation results in \autoref{tab:results_topic} indicates that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
 \\
 
-\paragraph{Speech sentiment correlates with political power} 
-But what are the underlying features that give rise to the classifiers performance?
-In order to investigate this the bag-of-words features were analysed with respect to their sentiment.
-The average sentiment of each political party is shown in \autoref{fig:party_sentiments}.
-High values indicate more pronounced usage of positive words, whereas negative values indicate more pronounced usage of words associated with negative emotional content.
-
-The results show an interesting relationship between political power and sentiment.
-Political power was evaluated in terms of membership of the government.
+\subsection{Sentiment correlates with political power} 
+The pronounced drop in performance for the out-of-domain data suggests that there is a dependency between text features and labels in the training data that is not present in the out-of-domain data.
+One candidate feature for this dependency that can be easily examined is sentiment.
+Analysing the correlation between proxies of political power and text sentiment shows an interesting relationship between political power and sentiment.
+Political power was evaluated in terms of membership of the government as well as number of seats in the parliament.
 Correlating these indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
@@
 %\autoref{fig:party_sentiments} and 
-in \autoref{tab:sentiments}: In the current Bundestag, government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
-
-%Note that there is one party, the social democrats (SPD), which has many seats and switched from opposition to government with the 18th Bundestag: With its participation in the government the average sentiment of this party switched sign from negative to positive, suggesting that positive sentiment is a strong indicator of government membership.
-%
-%
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=6cm]{images/party-sentiments-17.pdf}  \hfill \includegraphics[width=5cm]{images/party-sentiments-18.pdf}
-%%
-%\end{center}
-%\caption{
-%\label{fig:party_sentiments}
-%Speech sentiments computed for speeches of each party; parties are ordered according to the number of seats in the parliament.
-There is a trend for more positive speech content with more political power.
-Note that the SPD (red) switched from opposition to government in the 18th Bundestag: their seats in the parliament increased and the average sentiment of their speeches switched sign from negative to overall positive sentiment.
-%}
-%\end{figure}
+in \autoref{tab:sentiments}: Government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
 
@@
 
-\subsection{An example web application}
-To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major German news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
+%\subsection{An example web application}
+%To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major German news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
 For each news article topic, headlines of articles are plotted along with the predictions of the policy issue of an article and two labels derived deterministically from the 56 class output, a left right index and the political domain of a text, see \cite{leftright}.
@@
 We would like to thank Friedrich Lindenberg for factoring out the \url{https://github.com/bundestag/plpr-scraper} from his Bundestag project.
-Some backend configurations for the web application were taken from an earlier collaboration with Daniel Kirsch.
 Michael Gaebler provided helpful feedback on an earlier version of the manuscript.
+\\
 %
+
+\newpage
 \small{

### 2016-06-16 bundestag.py
Added balanced words experiment for fb evaluation data.

@@ -209,6 +209,16 @@ def get_raw_text_fb(folder="data/parteien-auf-fb", suffix=".json.gz"):
     files = glob.glob(folder+"/*"+suffix)
     return zip(*chain(*filter(None,map(fbFileTuple,files))))
 
+def get_raw_text_fb_balanced_words(nWords=1000,folder="data/parteien-auf-fb", suffix=".json.gz"):
+    data,labels = get_raw_text_fb(folder,suffix)
+    newData, newLabels = [], []
+    for p in sp.unique(labels):
+        alltext = " ".join([x[0] for x in zip(data,labels) if x[1]==p])
+        newTexts = map(lambda x: " ".join(x),zip(*[iter(alltext.split(" "))]*nWords))
+        newData.extend(newTexts)
+        newLabels.extend([p]*len(newTexts))
+    return newData,newLabels
+
 def fbFileTuple(fbf):
     f = [ff[1] for ff in fbTraverserTop(fbf)]
     return zip(f, [fbf.split("/")[-1].split(".")[0]] * len(f))
@@ -334,8 +344,8 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     parameters = {'vect__ngram_range': [(1,2)],\
            #'vect__stop_words':(None,stops),\
            #'tfidf__use_idf': (True,False),\
-           'clf__C': (10.**sp.arange(-4,4,2.)).tolist(),
-           'vect__max_df':[.5, .75],
+           'clf__C': (10.**sp.arange(-2,3,1.)).tolist(),
+           'vect__max_df':[.5],
         }
     saveId = idstr+"-"+randid()
     X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
@@ -396,7 +406,7 @@ def classify_speeches_party_parliament(legislationPeriod = 18):
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="parliament-train-%d"%legislationPeriod)
 
 def classify_speeches_fb(legislationPeriod = 18):
-    evalData, evalLabels = get_raw_text_fb()
+    evalData, evalLabels = get_raw_text_fb_balanced_words()
     trainData, trainLabels = get_raw_text_bundestag(legislationPeriod=legislationPeriod)
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="fb-test-%d"%legislationPeriod)
 

### 2016-06-16 manuscript/poltext.tex
Added fb data subsampling. Shortened text.

@@
 \author{Felix Biessmann\\
-  {\tt felix.biessmann@gmail.com} \\\And
+  {\tt felix.biessmann@gmail.com} \\ \And
+  Sebastian Schelter\\
+  Technische Universit\"at Berlin\\
+  {\tt sebastian.schelter@tu-berlin.de}\\
+  \And
  Pola Lehmann \\
@@
 \paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
-The protocols of plenary debates are available through the website of the German Bundestag\footnote{\url{https://www.bundestag.de/protokolle}}; an open source API was used to query the data in a cleaned and structured format\footnote{\url{https://github.com/bundestag}}.
-In total 22784 speeches were extracted for the 17th legislative period.
+The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}.
+In total 22784 speeches were extracted for the 17th legislative period\footnote{Uninterrupted parts were treated as separate speech.}.
 
@@
 Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
-Each quasi-sentence has  two types of labels: the party affiliation and the manually assigned policy issue aimed at in each specific statement.
-
+Each quasi-sentence has  two types of labels: the party affiliation and the manually assigned policy issue aimed at in each quasi-sentence.
+The policy labels were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
 The length of each annotated statement in the party manifestos is rather short.
-The median length is 95 characters or 12 words.\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.} This can be considered as a very valuable property of the data set, because it allows a fine grained resolution of party manifestos and is a good test for media data, as some data like tweets, facebook texts etc.
-are often very short, too.
-However for a classifier (as well as for humans) such short sentences can be rather difficult to classify.
-Benoit et al.
-(2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
-In the Manifesto Project the human coders therefore do not code these statements separately but within context.
-In order to obtain less 'noisy' data points from each party -- for the party affiliation task only -- all statements were aggregated into eight policy domains using the 56 policy categories from the Manifesto Project.
-Within the Manifesto Project category scheme each of these belongs to one of seven policy domains.
-These were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
-An eight topic is generated for all statements not belonging to any one the 56 policy issues, for a complete list see also \cite{leftright}.
-Most party manifestos covered all eight topics, some party manifestos in the 17th Bundestag only covered seven.
+The median length is 95 characters or 12 words.\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}
 
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-All text data was segmented into semantic units, quasi-sentences for manifesto data and (uninterrupted) speech parts for parliament speech data.
-Strings of each semantic unit were tokenised and transformed into bag-of-word (BOW) vectors as implemented in scikit-learn \cite{scikit-learn}.
+All text data was tokenised and transformed into bag-of-word (BOW) vectors as implemented in scikit-learn \cite{scikit-learn}.
 Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
@@
 The model pipeline contained a number of  hyperparameters that were optimised using cross-validation.
-We first split the training data into a training data set that was used for optimisation of hyperparameters and an held-out test data set for evaluating how well the model performs on in-domain data.
-Hyperparameters were optimised using grid search and 3-fold cross-validation within the training set -- no data from the in-domain evaluation data or the out-of-domain evaluation data were used for hyperparameter optimisation.
-For the best setting of all hyperparameters the pipeline was trained again on all training data and evaluated on the evaluation data sets.
-For party affiliation prediction and government membership prediction the training and test set were 90\% and 10\%, respectively, of all data in a given legislative period.
-Wherever possible the generalisation performance of the models was also evaluated on out-of domain texts, for which we used the texts from party manifestos.
-For the policy issue prediction setting there was no out-of-domain evaluation data, so all labeled manifesto sentences in both legislative periods were split into a training and evaluation set of 90\% (train) and 10\% (evaluation).
-The train and test splits are summarized in \autoref{tab:classification_settings}.
-% 
-\begin{table}[t]\label{tab:classification_settings}
-\footnotesize
-\begin{tabular}{ccc}
-Training Set & \multicolumn{2}{c}{Test set} \\
-& In-domain&Out-of-domain\\
-\hline
-\hline
-Parliament ($90\%$) & Parliament ($10\%$)& Manifestos\\
-Manifestos ($90\%$)& Manifestos ($10\%$) & -\\
-\end{tabular}
-\end{table}
+The parliament speech data was split into training and validation (for hyperparameter optimization) in a 90\%/10\% ratio.
 %

### 2016-06-16 bundestag.py

@@ -212,7 +212,8 @@ def get_raw_text_fb(folder="data/parteien-auf-fb", suffix=".json.gz"):
 def get_raw_text_fb_balanced_words(nWords=1000,folder="data/parteien-auf-fb", suffix=".json.gz"):
     data,labels = get_raw_text_fb(folder,suffix)
     newData, newLabels = [], []
-    for p in sp.unique(labels):
+    uparties, upcounts = sp.unique(labels,return_counts=True)
+    for p in uparties:
         alltext = " ".join([x[0] for x in zip(data,labels) if x[1]==p])
         newTexts = map(lambda x: " ".join(x),zip(*[iter(alltext.split(" "))]*nWords))
         newData.extend(newTexts)
@@ -405,8 +406,13 @@ def classify_speeches_party_parliament(legislationPeriod = 18):
     evalData, evalLabels = get_raw_text(legislationPeriod=legislationPeriod)
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="parliament-train-%d"%legislationPeriod)
 
-def classify_speeches_fb(legislationPeriod = 18):
-    evalData, evalLabels = get_raw_text_fb_balanced_words()
+def classify_speeches_fb(legislationPeriod = 18, nsamples=50):
+    evalData, evalLabels = get_raw_text_fb_balanced_words(nWords=100)
+    uparties, upcounts = sp.unique(evalLabels,return_counts=True)
+    dd = zip(evalData,evalLabels)
+    sp.random.shuffle(dd)
+    hh = [[x for x in dd if x[1]==up] for up in uparties]
+    evalData,evalLabels = zip(*chain(*[h[:nsamples] for h in hh]))
     trainData, trainLabels = get_raw_text_bundestag(legislationPeriod=legislationPeriod)
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="fb-test-%d"%legislationPeriod)
 

### 2016-06-17 manuscript/poltext.tex
removed web app references, added authors

@@
 \author{Felix Biessmann\\
-  {\tt felix.biessmann@gmail.com} \\ \And
+%  {\tt felix.biessmann@gmail.com} \\ 
+  \And
   Sebastian Schelter\\
   Technische Universit\"at Berlin\\
-  {\tt sebastian.schelter@tu-berlin.de}\\
+%  {\tt sebastian.schelter@tu-berlin.de}\\
+\And 
+Daniel Kirsch
   \And
@@
   Wissenschaftszentrum Berlin \\
-  {\tt pola.lehmann@wzb.eu} \\}
+%  {\tt pola.lehmann@wzb.eu} \\
+  }
 
@@
 We propose methods to better interpret these results and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
-We present an example of a web application that combines topic modelling with automatic political affiliation analyses of news articles based on the proposed model.
+%We present an example of a web application that combines topic modelling with automatic political affiliation analyses of news articles based on the proposed model.
 \end{abstract}

### 2016-06-17 manuscript/poltext.tex
Ran more experiments on binary prediction, also with text aggregation Shortened more text passages and squeezed it into 5 pages Included fb data.

@@
 
-One area which produces large amounts of text which are of high interest for political scientists is modern media.
-Standard newspapers, but also new forms as twitter, facebook or blogs produce an ever growing amount of text influencing political debate.
-These texts can be of interest to political scientists in various analyses.
-The question we will focus on in this analysis is the potential political bias that can be detected in these sources.
-In many cases it is obvious which political bias an author has.
-In other cases some expertise is required to judge the political bias of a text.
-Keeping an unbiased view on what media report on requires to understand the political bias of texts.
-Different newspaper outlets are, for example, often ascribed as favoring different positions on the political left-right axis.
-Data from the European Media Systems Survey show that experts clearly associate different media outlets with specific parties (http://www.mediasystemsineurope.org).
-However such expert surveys on the political bias of newspapers can normally only give us a very general tendency of a newspaper, they do not distinguish between different journalists or different sections.
-Assistive technology can help in this context to try and obtain a more unbiased sample of information.
-It gives us the possibility to get a more nuanced view of the spectrum of political positions favored in important news outlets in a country.
-In this paper we test whether it is possible to automatically detect political party affiliation in text and if such a model can be used on different kinds of texts that it was trained on.
-In order to validate and explain the predictions of the models three strategies that allow for better interpretations of the models are proposed.
+A major problem with automated text analyses is that in contrast to human coders, who have read enough texts of different domains and are able to detect political bias in a variety of text domains, machine learning algorithms are naturally prone to poor generalization performance if the training data is biased towards one text domain.
+Yet, good training data is difficult to obtain.
+One of the best sources for automated political text analysis systems are plenary debates of the parliament: It is a large source of text data that can be clearly associated with a party, despite the intra-party heterogeneity of opinions.
+Thus many studies focus on the analysis of models and evaluate them on the same type of data.
+Here we examine to what extent these models generalize their performance to other text domains, such as party manifestos and texts from social media.
+We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.
+
+We investigate predictions of the models with three strategies.
 First the model misclassifications are related to changes in party policies.
-Second univariate measures of correlation between text features and party affiliation allow to relate the predictions to the kind of information that political experts use for interpreting texts.
-Third sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
+Second sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
+Third univariate measures of correlation between text features and party affiliation allow to relate the predictions to the kind of information that political experts use for interpreting texts.
 
@@
 \subsection{Data}
-Annotated political text data was obtained from two sources: a) the plenary debates held in the German parliament ({\em Bundestag}) and b) all manifesto texts of parties winning seats in the election to the German parliament in the the last, 17th, legislative period.
+Annotated political text data was obtained from three sources: a) the plenary debates held in the German parliament ({\em Bundestag}) b) all manifesto texts of parties winning seats in the election to the German parliament and c) facebook page posts of all parties.
+The texts from plenary debates were used to train a classifier and evaluate it on this in-domain data.
+The latter two data sources were used to test the generalization performance of the classifier on out-of-domain data.
 
@@
 The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}.
-In total 22784 speeches were extracted for the 17th legislative period\footnote{Uninterrupted parts were treated as separate speech.}.
+Uninterrupted parts were treated as separate speech.
 
@@
 Each quasi-sentence has  two types of labels: the party affiliation and the manually assigned policy issue aimed at in each quasi-sentence.
-The policy labels were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
 The length of each annotated statement in the party manifestos is rather short.
 The median length is 95 characters or 12 words.\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}
+In order to increase the length of the texts used for classification the policy labels were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
+So in this setting each party had just one data point for each of the topics.
+
+\paragraph{Facebook post data}
+For each party we crawled their facebook pages \cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post text, excluding all comments and other information.
+Like the manifesto data, also these texts were very short.
+As aggregation per topics was not possible for this data, we aggregated the texts by splitting all texts into parts of 100 words.
+In order to balance the classes we drew 50 random parts for each party.
+This procedure reduces the effect of text length and biased sampling of different parties when testing the classifier.
 
@@
 Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
-All of these hyperparameters were subjected to hyperparameter optimization as explained in \autoref{sec:crossvalidation}.
 
@@
 Bag-of-words feature vectors were used to train a multinomial logistic regression model.
-Let $y\in\{1,2,\dots,K\}$ be the true  label, where $K$ is the total number of labels and $\vec{W}=[\vec{w}_1,\dots,\vec{w}_K]\in\R^{d\times K}$ is the concatenation of the weight vectors $\vec{w}_k$ associated with the $k$th party then
+Let $y\in\{1,2,\dots,K\}$ be the true party affiliation and $\vec{w}_1,\dots,\vec{w}_K\in\R^{d}$ the weight vectors associated with the $k$th party then the party affiliation estimate is modelled as
 \begin{eqnarray}\label{eq:logreg_multiclass}
-p(y=k|\vec{x},\vec{W}) = \frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}}  \textrm{ with }  z_k=\vec{w}_k^{\top}\vec{x}.
+p(y=k|\vec{x}) = \frac{e^{z_k}}{\sum_{j=1}^K e^{z_j}}  \textrm{ with }  z_k=\vec{w}_k^{\top}\vec{x}.
 \end{eqnarray}
@@
 The model pipeline contained a number of  hyperparameters that were optimised using cross-validation.
-The parliament speech data was split into training and validation (for hyperparameter optimization) in a 90\%/10\% ratio.
+The parliament speech data was split into training and validation set (for hyperparameter optimization) in a 90\%/10\% ratio.
 %
@@
 A sentiment vector $\vec{s}\in\R^d$ was constructed from the sentiment polarity values in the sentiment dictionary.
-The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors and sentiment vectors.
+The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors and sentiment vector.
 
@@
 Interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
-Thus direct interpretation of the model coefficients $\vec{W}$ is problematic, see also \cite{Zien2009, Haufe2013}.
-In order to allow for better interpretation of the predictions and to assess which features are discriminative correlation coefficients between each word and the party affiliation label were computed.
-The words corresponding to the top positive and negative correlations are shown in \autoref{sec:word_party_correlations}.
+Thus direct interpretation of the model coefficients $\vec{w}_k$ is problematic, see also \cite{Zien2009, Haufe2013}.
+In order to allow for better interpretation of the predictions and to assess which features are discriminative, correlation coefficients between each word and the party affiliation label were computed.
 
@@
 \subsection{Predicting political party affiliation}
+When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
 The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
-
-When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
 These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.
-For out-of domain data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
-This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
-\paragraph{Length of texts influences prediction accuracy}
-A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics.
-The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
-This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
 
@@
 
+\paragraph{Validation on manifesto data}
+For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
+This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
+When testing the classifier on the facebook data (results not shown), this drop is similar.
+
+\paragraph{Length of texts influences prediction accuracy}
+A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
+In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics.
+The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
+This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
+As there were no topic labels for the facebook data, those texts were aggregated by sampling 50 parts of 100 words parts of the concatenated facebook posts of a party.
+The results are shown in \autoref{tab:results_fb} and indicate that also for these texts prediction accuracies comparable to the in-domain case can be achieved by increasing the length of the texts.
+
 \begin{table}[t]
@@
 \label{tab:results_out-of-domain}
-{\bf Out-of-domain classification performance} (quasi-sentence level) of a classifier trained on speeches of the 17th legislative period of the Bundestag.
+{\bf Out-of-domain} classification performance (quasi-sentence level) on {\bf manifesto data} of a classifier trained on speeches of the 17th legislative period of the Bundestag.
 }
@@
 \label{tab:results_topic}
-{\bf Out-of-domain classification performance} (topic level) on manifesto data of a classifier trained on the 17th legislative period of the Bundestag.
+{\bf Out-of-domain} classification performance (topic level) on {\bf manifesto data} of a classifier trained on the 17th legislative period of the Bundestag.
 Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}) the predictions made on topic level are more reliable.}
@@
 
+\begin{table}[t]
+\caption{
+\label{tab:results_fb}
+{\bf  Out-of-domain} classification performance (text length: 100 words) on texts of {\bf facebook posts} of respective party data using a classifier trained on the 18th legislative period of the Bundestag.
+The average prediction performance is comparable to that on in-domain test data.}
+\begin{center}
+\begin{tabular}{lcccc}
+    &         precision    &recall &  f1-score  & N  \\
+    \hline
+        \hline
+     cducsu &      0.66    &  0.90  &    0.76    &    50\\
+     gruene   &    0.43   &   0.20  &    0.27   &     50\\
+      linke     &  0.49    &  0.58   &   0.53   &     50\\
+        spd    &   0.82    &  0.82  &    0.82   &     50\\
+\hline
+avg / total     &  0.60   &   0.62   &   0.60  &     200\\
+
+\end{tabular}
+\end{center}
+
+\end{table}
+
 \paragraph{Policy change}
@@
 \vspace{0.5em}
-\footnotesize
-\begin{tabular}{lcccccc}
+\begin{tabular}{lc|ccccc}
 && \multicolumn{5}{c}{\bf Predicted}\\
 && cducsu & fdp& gruene& linke& spd\\
-\hline
 \hline
@@
 \label{tab:results_binary_17}
-Classification performance on the binary prediction problem in the 17th legislative period, categorizing speeches into government (FDP/CDU/CSU) and opposition (Linke, Gr\"une, SPD).
+Classification accuracy on the binary prediction problem, categorizing texts into government and opposition.
+Out-of-domain accuracy again drops close to chance performance for the manifesto data but remains higher for the facebook post texts.
 }
 \begin{center}
-\begin{tabular}{lcccccccc}
-& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
-    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
-\hline \hline
-government    &   0.83   &   0.84&      0.84     & 1037 & 0.49  &    0.59 &     0.54  &    4349\\
- opposition     &  0.86  &    0.86   &   0.86    &  1242 & 0.74 &     0.66  &    0.70   &   7726\\
+\begin{tabular}{lccc}
+& {\bf In-Domain} & \multicolumn{2}{c}{{\bf Out-of-Domain}}\\
+& Parliament & Manifestos & Facebook Posts\\
 \hline
-avg / total   &    0.85 &     0.85 &     0.85  &    2279 & 0.65  &     0.63  &     0.64  &    12075\\
-
+Accuracy    &   0.88   &   0.60&      0.71\\
 %
@@
 \end{table*}
-%%
-%%\begin{table*}[t]
-%%\caption{
-%%\label{tab:results_binary_18}
-%%Classification performance on the binary prediction problem in the 18th legislative period, categorizing speeches into government (SDP/CDU/CSU) and opposition (Linke, Gr\"une).
-%%}
-%%\begin{center}
-%%\begin{tabular}{lcccccccc}
-%%& \multicolumn{4}{c}{\bf Held-out parliament speeches} & \multicolumn{4}{c}{\bf Party Manifestos}\\
-%%    &         precision    &recall &  f1-score  & N    &         precision    &recall &  f1-score  & N\\
-%%\hline \hline
-%%government   &    0.88  &    0.95    &  0.92   &    786   &0.52   &   0.66 &     0.58  &    5972\\
-%% opposition    &   0.86    &  0.71   &   0.78   &    346 & 0.69  &    0.56    &  0.62   &   8229\\
-%%\hline
-%%avg / total    &   0.88    &  0.88     & 0.87  &    1132 &  0.62   &   0.60    &  0.60 &    14201\\
-%%%
-%%\end{tabular}
-%%\end{center}
-%%\end{table*}
-
-%
-\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
-To get a more precise idea of what word features are discriminative in terms of the party affiliation label correlations between words and party affilation labels are computed; these are the words that are preferentially used or avoided by each respective party.
-Some unspecific stopwords are excluded.
-We list for each party some of the top positively correlated words, sorted in descendent order.
-\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
-\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
-\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em Staatssekretaerin, Kanzlerin, Bundestagsfraktion}) and some words related to cutting of expenses ({\em Kuerzungen}).
-\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
-Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit, Stabilitaet}).
 
 \subsection{Predicting government status}\label{sec:sentiment_result}
-Next to the party affiliation labels also government membership labels were used to train models that predict whether or not a text is from a party that belonged to a governing coalition of the Bundestag.
+For a better comparison with other studies that predict party affiliation in a two party system we also trained a model on government membership labels.
 In \autoref{tab:results_binary_17} the results are shown for the 17th legislative period.
-While the in-domain evaluation precision and recall values reach values close to 0.9, the out-of-domain evaluation drops again to values between 0.6 and 0.7.
+While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on short quasi-sentences drops again close to chance performance.
 This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
@@
 older texts or texts from another chamber).
-While we did not perform topic-level predictions in this binary setting, the party affiliation results in \autoref{tab:results_topic} indicates that a similar increase in out-of-domain prediction accuracy could be achieved when aggregating texts to longer segments.
-\\
+Aggregation into topics did not increase the accuracy in this binary setting.
+The drop in accuracy when applying the binary classifier on facebook data (aggregated as in the party affiliation case) was less pronounced, accuracies were above 0.70.
+These results indicate that there might be some signal in the facebook data that is absent in the manifesto data.
+One such signal could be related to the fact that after the election, both the speakers in the parliament as well as the authors of the facebook posts  know which party is in the governmental coalition -- while the authors of the manifesto texts do not know that; these texts were written before the elections.
+A language feature that could capture this is sentiment.
 
 \subsection{Sentiment correlates with political power} 
-The pronounced drop in performance for the out-of-domain data suggests that there is a dependency between text features and labels in the training data that is not present in the out-of-domain data.
-One candidate feature for this dependency that can be easily examined is sentiment.
-Analysing the correlation between proxies of political power and text sentiment shows an interesting relationship between political power and sentiment.
-Political power was evaluated in terms of membership of the government as well as number of seats in the parliament.
-Correlating these indicators of political power with the mean sentiment of a party shows a strong positive correlation between speech sentiment and political power.
-This pattern is evident from the data in
-%\autoref{fig:party_sentiments} and 
-in \autoref{tab:sentiments}: Government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
+We find a strong positive correlation between speech sentiment and political power measured in parliament seats and government membership, see  \autoref{tab:sentiments}: Government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
 
@@
 
+%
+\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
+We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label. Some unspecific stopwords are excluded. 
+\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
+\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
+\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em Staatssekretaerin, Kanzlerin, Bundestagsfraktion}) and some words related to cutting of expenses ({\em Kuerzungen}).
+\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit, Stabilitaet}).
+
 %\subsection{An example web application}
@@
 
-\section{Conclusions, Limitations and Outlook}\label{sec:conclusion}
-This study presents a simple approach for automated political bias prediction.
-The results of these experiments show that automated political bias prediction is possible with above chance accuracy in some cases.
-It is worth noting that even if the accuracies are not perfect, they are above chance and comparable with results of comparable studies \cite{Yu2008, Hirst2014}.
-While these results do not allow for usage in production systems for classification, it is well possible to use such a system as assistive technology for human annotators in an active learning setting.
+\section{Conclusions and Limitations}\label{sec:conclusion}
+Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy in some cases even beyond the training text domain.
+These results encourage to use such systems as assistive technology for instance for human annotators in an active learning setting.
 
-One of the main limiting factors of an automated political bias prediction system is the availability of training data.
-Most training data sets that are publicly available have an inherent bias as they are sampled from a different domain.
-This study tried to quantify the impact of this effect.
-For the cases in which evaluation data from two domains was available there was a pronounced drop in prediction accuracy between the in domain evaluation set and the out of domain evaluation set.
-This effect was reported previously for similar data, see e.g.
-\cite{Yu2008}.
-Also the finding that shorter texts are more difficult to classify than longer texts is in line with previous studies \cite{Hirst2014}.
-When considering texts of sufficient length (for instance by aggregating all texts of a given political topic) classification performance improved and in some cases reliable predictions could be obtained even beyond the training text domain.
-
-Some aspects of these analyses could be interesting for social science researchers; three of these are highlighted here.
-First the misclassifications of a model can be related to the changes in policy of a party.
-Such analyses could be helpful to quantitatively investigate a change in policy.
-Second analysing the word-party correlations shows that some discriminative words can be related to the political views of a party; this allows for validation of the models by human experts.
-Third when correlating the sentiment of a speech with measures of political power there is a strong positive correlation between political power and positive sentiment.
-While such an insight in itself might seem not very surprising this quantifiable link between power and sentiment could be useful nonetheless: Sentiment analysis is a rather domain independent measure, it can be easily automated and scaled up to massive amounts of text data.
-Combining sentiment features with other measures of political bias could potentially help to alleviate some of the domain-adaptation problems encountered when applying models trained on parliament data to data from other domains.
+We find a strong effect of text length and text domain on the generalization performance of the classifier, as also shown in other studies on similar data \cite{Yu2008, Hirst2014}.
+The first effect, longer texts are easier to classify, makes intuitive sense.
+Also humans are worse at judging the political bias of shorter texts out of context: Benoit et al.
+(2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
+However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short.
+But political education would benefit from automatic analyses of these data streams; these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
 \\
-
-All data sets used in this study were publicly available, all code for experiments and the live web application can be found online \cite{fipi, fipidemo}.
+The second effect, drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments.
+In the case of party affiliation prediction, the out-of-domain classification is on par or even better than the prediction accuracy on in-domain data.
+However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification.
+We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
 
@@
 %
-
 \newpage

### 2016-06-17 bundestag.py

@@ -362,6 +362,7 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     labelsStr = [str(x) for x in test_clf.steps[-1][1].classes_]
     report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(y_test, predictedTest).tolist(): report += str(line)+"\n" 
+    report += "Accuracy: %0.2f\n"%metrics.accuracy_score(y_test,predictedTest)
     # train again on entire training set
     final_clf = text_clf.set_params(**best_clf.get_params()).fit(trainData, trainLabels)
     # dump classifier to pickle
@@ -372,6 +373,7 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     labelsStr = [str(x) for x in final_clf.steps[-1][1].classes_]
     report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(evalLabels, predictedEval).tolist(): report += str(line)+"\n" 
+    report += "Accuracy: %0.2f\n"%metrics.accuracy_score(evalLabels,predictedEval)
     # dump report
     open(OUT_DIR+'/report-'+saveId+'.txt','wb').write(report)
     return report
@@ -401,6 +403,15 @@ def classify_speeches_binary_parliament(legislationPeriod = 18):
     evalData, evalLabels = zip(*[(x[0],'government') if x[1] in gov else (x[0],'opposition') for x in zip(evalDataParty,evalLabelsParty)])
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="parliament-train-gov-%d"%legislationPeriod)
 
+
+def classify_speeches_binary_parliament_topics(legislationPeriod = 18):
+    trainDataParty, trainLabelsParty = get_raw_text_bundestag(legislationPeriod=legislationPeriod)
+    evalDataParty, evalLabelsParty = get_raw_text_topics(legislationPeriod=legislationPeriod)
+    gov = bundestagGovernment[legislationPeriod]['government']
+    trainData, trainLabels = zip(*[(x[0],'government') if x[1] in gov else (x[0],'opposition') for x in zip(trainDataParty,trainLabelsParty)])
[…]

### 2016-06-17 manuscript/poltext.tex
Shortened abstract, formatted authors.

@@
 
-\author{Felix Biessmann\\
-%  {\tt felix.biessmann@gmail.com} \\ 
+\author{Felix Biessmann\thanks{~\tt felix.biessmann@gmail.com}\\
   \And
-  Sebastian Schelter\\
-  Technische Universit\"at Berlin\\
-%  {\tt sebastian.schelter@tu-berlin.de}\\
+  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ Technische Universit\"at Berlin
 \And 
@@
   \And
- Pola Lehmann \\
-  Wissenschaftszentrum Berlin \\
-%  {\tt pola.lehmann@wzb.eu} \\
+ Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
+  Wissenschaftszentrum Berlin\\
 }
@@
 Every day large amounts of text are produced in public discourse.
-Some of this text is produced by actors whose political colour is very obvious, for example in the case of party elites.
-But there are also many actors raising their voices in the discourse, who cannot be clearly attributed to a political party, while their statements might actually by biased towards a specific party.
+Some of this text is produced by actors whose political colour is very obvious.
+But there are also many actors, who cannot be clearly attributed to a political party, while their statements might be biased towards a specific party.
 Identifying such biases is crucial for political research as well as media consumers, especially when analysing the influence of the media on political discourse and vice versa.
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-%Some interpretations of the results are highlighted and a web application of the models is presented at the end of the section. 
-Due to space restrictions we only report results from the 17th legislative period. Results are similar for the 18th legislative period.
+Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.
 
@@
 \label{tab:results_topic}
-{\bf Out-of-domain} classification performance (topic level) on {\bf manifesto data} of a classifier trained on the 17th legislative period of the Bundestag.
+{\bf Out-of-domain} classification performance (topic level) on {\bf manifesto data}.
 Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}) the predictions made on topic level are more reliable.}
@@
 \begin{table}[t]\label{tab:conf_mat_four_class}
-\caption{\label{tab:confusion_topic} {\bf Topic level confusion matrices} on evaluation data (manifesto texts), classifiers were trained on speeches of 17th Bundestag.}
+\caption{\label{tab:confusion_topic} {\bf Topic level confusion matrices} of manifesto texts.}
 \vspace{0.5em}
@@
 
-%\subsection{An example web application}
-%To show an example use case of the above models a web application was implemented that downloads regularly all articles from some major German news paper websites\footnote{\url{http://www.spiegel.de/politik}, \url{http://www.faz.net/aktuell/politik}, \url{http://www.welt.de/politik}, \url{http://www.sueddeutsche.de/politik}, \url{http://www.zeit.de/politik}} and applies some simple topic modelling to them.
-For each news article topic, headlines of articles are plotted along with the predictions of the policy issue of an article and two labels derived deterministically from the 56 class output, a left right index and the political domain of a text, see \cite{leftright}.
-Within each topic it is then possible to get an ordered (from left to right) overview of the articles on that topic.
-A preliminary demo is live at \cite{fipidemo} and the code is available on github\cite{fipi}.
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=10cm]{images/fipi-screenshot}
-%%
-%\end{center}
-%\caption{
-%\label{fig:fipi}
-%A screen shot of an example web application using the political view prediction combined with topic modelling to provide a heterogeneous overview of a topic. }
-%\end{figure}
-
 \section{Conclusions and Limitations}\label{sec:conclusion}
-Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy in some cases even beyond the training text domain.
+Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
 These results encourage to use such systems as assistive technology for instance for human annotators in an active learning setting.

### 2016-06-17 manuscript/poltext.tex
:x

@@
     \And
-  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ Technische Universit\"at Berlin
+ Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
+  Wissenschaftszentrum Berlin\\
 \And 
@@
  \And
- Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
-  Wissenschaftszentrum Berlin\\
+  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ TU Berlin
 }

### 2016-06-18 manuscript/poltext.tex
FB experiments on more data more fb results in text.

@@
 Some of this text is produced by actors whose political colour is very obvious.
-But there are also many actors, who cannot be clearly attributed to a political party, while their statements might be biased towards a specific party.
+But many actors cannot be clearly attributed to a political party, yet their statements might be biased towards a specific party.
 Identifying such biases is crucial for political research as well as media consumers, especially when analysing the influence of the media on political discourse and vice versa.
@@
 We propose methods to better interpret these results and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
-%We present an example of a web application that combines topic modelling with automatic political affiliation analyses of news articles based on the proposed model.
 \end{abstract}
@@
 
-A major problem with automated text analyses is that in contrast to human coders, who have read enough texts of different domains and are able to detect political bias in a variety of text domains, machine learning algorithms are naturally prone to poor generalization performance if the training data is biased towards one text domain.
-Yet, good training data is difficult to obtain.
-One of the best sources for automated political text analysis systems are plenary debates of the parliament: It is a large source of text data that can be clearly associated with a party, despite the intra-party heterogeneity of opinions.
-Thus many studies focus on the analysis of models and evaluate them on the same type of data.
-Here we examine to what extent these models generalize their performance to other text domains, such as party manifestos and texts from social media.
+A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained.
+Human coders have read enough texts of different domains to detect political bias in a variety of contexts and styles.
+Machine learning algorithms however are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
+Unfortunately good unbiased training data is difficult to obtain.
+One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data as it is a large source of text that can be clearly associated with a party.
+Here we examine to what extent these models generalise their to other text domains, such as party manifestos and texts from social media.
 We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.
@@
 
-In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results and future research directions.
+In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.
 
@@
 
-\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label, which we take here as a proxy for political bias.
+\paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label.
 The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}.
@@
 The data released in this project mainly comprises the complete manifestos of all parties that have won seats at a national election.
-Each quasi-sentence\footnote{A quasi-sentence has the length of an argument.
+Each statement or {\em quasi-sentence}\footnote{A quasi-sentence has the length of an argument.
 It is never longer than one sentence.} is annotated with one of 56 policy issue categories.
@@
 Like the manifesto data, also these texts were very short.
-As aggregation per topics was not possible for this data, we aggregated the texts by splitting all texts into parts of 100 words.
+As aggregation per topics was not possible for this data, we aggregated the texts by splitting all texts into parts of 1000 words.
 In order to balance the classes we drew 50 random parts for each party.
@@
 All text data was tokenised and transformed into bag-of-word (BOW) vectors as implemented in scikit-learn \cite{scikit-learn}.
-Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and several cutoffs for discarding too frequent and too infrequent words.
+Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and different cutoffs for discarding too frequent and too infrequent words.
 
@@
 \subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
-The model pipeline contained a number of  hyperparameters that were optimised using cross-validation.
-The parliament speech data was split into training and validation set (for hyperparameter optimization) in a 90\%/10\% ratio.
+The model pipeline contained a number of  hyperparameters that were optimised using gridsearch cross-validation.
+To this end the parliament speech data was split into training and validation set in a 90\%/10\% ratio; the pipeline was trained with each parameter setting on the training set and its performance validated on the validation set.
+The parameters of the best performing model were then used to train a model on the training and validation set data.
+None of the data in the separately held back in-domain test data nor the out-of-domain test data sets was used for this hyperparameter optimization.
 %
@@
 This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
-As there were no topic labels for the facebook data, those texts were aggregated by sampling 50 parts of 100 words parts of the concatenated facebook posts of a party.
-The results are shown in \autoref{tab:results_fb} and indicate that also for these texts prediction accuracies comparable to the in-domain case can be achieved by increasing the length of the texts.
+
+\paragraph{Validation on facebook data}
+For the facebook post data a similar effect was observed.
+Short texts resulted in poor prediction accuracies of 0.51 on average.
+In addition classes were highly unbalanced in this setting since some parties have an order of magnitude more posts than others.
+In order to reduce the impact of text length and unbiased sampling and as there were no topic labels for the facebook posts, those texts were aggregated by sampling 50 texts of 1000 words each of the concatenated facebook posts of a party.
+The results are shown in \autoref{tab:results_fb} and indicate that also for these texts prediction accuracies comparable to the in-domain case can be achieved.
 
@@
 \label{tab:results_fb}
-{\bf  Out-of-domain} classification performance (text length: 100 words) on texts of {\bf facebook posts} of respective party data using a classifier trained on the 18th legislative period of the Bundestag.
+{\bf  Out-of-domain} classification performance on 50 randomly selected {\bf facebook posts} of respective party (text length: 1000 words).
 The average prediction performance is comparable to that on in-domain test data.}
@@
         \hline
-     cducsu &      0.66    &  0.90  &    0.76    &    50\\
-     gruene   &    0.43   &   0.20  &    0.27   &     50\\
-      linke     &  0.49    &  0.58   &   0.53   &     50\\
-        spd    &   0.82    &  0.82  &    0.82   &     50\\
+ cducsu     &  0.65     & 1.00  &    0.79     &   50\\
+     gruene   &    0.67   &   0.12  &    0.20   &     50\\
+      linke       &0.60    &  0.82    &  0.69    &    50\\
+        spd       &1.00 &     0.92   &   0.96    &    50\\
 \hline
-avg / total     &  0.60   &   0.62   &   0.60  &     200\\
-
+avg / total    &   0.73   &   0.71  &    0.66   &    200\\
 \end{tabular}
@@
 In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for topic level predictions (see \autoref{tab:confusion_topic}).
-On the topic level results for the 17th legislative period are pretty good except for the SPD, which the classifier often predicts as CDU/CSU.
+On the topic level results for the 17th legislative period are good except for the SPD, which the classifier often predicts as CDU/CSU.
 One explanation for this misclassification could be policy changes.
+A similar effect can be observed in the 18th Bundestag when predicting facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
 \\
@@
 \hline
-Accuracy    &   0.88   &   0.60&      0.71\\
+Accuracy    &   0.88   &   0.60&      0.76\\
 %
@@
 A pronounced drop in performance was reported when applying the model to texts from a different domain (e.g.
-older texts or texts from another chamber).
-Aggregation into topics did not increase the accuracy in this binary setting.
+older texts or texts from another chamber), a similar effect as we find.
+Aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
 The drop in accuracy when applying the binary classifier on facebook data (aggregated as in the party affiliation case) was less pronounced, accuracies were above 0.70.
-These results indicate that there might be some signal in the facebook data that is absent in the manifesto data.
-One such signal could be related to the fact that after the election, both the speakers in the parliament as well as the authors of the facebook posts  know which party is in the governmental coalition -- while the authors of the manifesto texts do not know that; these texts were written before the elections.
+These results indicate that there might be some signal common to plenary debates and the facebook data that is absent in the manifesto data.
+One such signal could be related to the fact that after the election, both the speakers in the parliament as well as the authors of the facebook posts  know which party is in the governmental coalition -- while the authors of the manifesto texts do not know that as these texts were written before the elections.
 A language feature that could capture this is sentiment.
@@
 We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
-We find a strong positive correlation between speech sentiment and political power measured in parliament seats and government membership, see  \autoref{tab:sentiments}: Government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+Our results in  \autoref{tab:sentiments} show that government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
 

### 2016-06-18 bundestag.py

@@ -418,7 +418,8 @@ def classify_speeches_party_parliament(legislationPeriod = 18):
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="parliament-train-%d"%legislationPeriod)
 
 def classify_speeches_fb(legislationPeriod = 18, nsamples=50):
-    evalData, evalLabels = get_raw_text_fb_balanced_words(nWords=100)
+    evalData, evalLabels = get_raw_text_fb_balanced_words(nWords=1000)
+    #evalData, evalLabels = get_raw_text_fb()
     uparties, upcounts = sp.unique(evalLabels,return_counts=True)
     dd = zip(evalData,evalLabels)
     sp.random.shuffle(dd)
@@ -428,7 +429,8 @@ def classify_speeches_fb(legislationPeriod = 18, nsamples=50):
     optimize_hyperparams(trainData,trainLabels, evalData, evalLabels,idstr="fb-test-%d"%legislationPeriod)
 
 def classify_speeches_fb_binary(legislationPeriod = 18, nsamples=50):
-    evalData, evalLabels = get_raw_text_fb_balanced_words(nWords=100)
+    evalData, evalLabels = get_raw_text_fb_balanced_words(nWords=1000)
+    #evalData, evalLabels = get_raw_text_fb()
     uparties, upcounts = sp.unique(evalLabels,return_counts=True)
     dd = zip(evalData,evalLabels)
     sp.random.shuffle(dd)

### 2016-06-21 manuscript/poltext.tex
Kleine Änderungen im gesamten Text und einige Umstellungen in results section

@@
 \documentclass[11pt]{article}
-\usepackage{poltext2016}
+%\usepackage{poltext2016}
 \usepackage{times}
@@
  Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
-  Wissenschaftszentrum Berlin\\
+  Wissenschaftszentrum Berlin für Sozialforschung\\
 \And 
@@
 }
+% Sieht irgendwie komisch aus, mit nur zwei angegebenen affiliations, und meine ist ja auch so lang. Sollen wir die nicht in die Fussnote zur email Adresse packen?
 
@@
 %
-Analysis and classifications of political text is and has been a very important tool to generate political science data (Benoit et al.
-2015).
-Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project (Budge et al.
-2001; Klingemann et al.
-2006), the Comparative Agendas Project or Poltext (Pétry/Duval 2015).} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
+Analysis and classifications of political text is and has been a very important tool to generate political science data \cite{Benoit.Forthcoming}.
+Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
 The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists.
-Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve (Benoit et al.
-2009)
+Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve \cite{Benoit.2}.
 
-A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained.
+A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained \cite{Slapin.2014}.
 Human coders have read enough texts of different domains to detect political bias in a variety of contexts and styles.
+% Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme.
+Das ist wahrscheinlich wieder der Unterschied der Disziplinen, ihr müsst rechtfertigen, warum Computer es können, wenn Menschen es können, wir müssen in der Regel rechtgertigen warum Menschen es können, wenn Computer es besser machen.
 Machine learning algorithms however are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
@@
 One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data as it is a large source of text that can be clearly associated with a party.
-Here we examine to what extent these models generalise their to other text domains, such as party manifestos and texts from social media.
+Here we examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media.
 We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.
@@
 
-In the following \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.
+In the following, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.
 
@@
 The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}.
-Uninterrupted parts were treated as separate speech.
+Each uninterrupted part was treated as a separate speech.
 
@@
 For each party we crawled their facebook pages \cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post text, excluding all comments and other information.
-Like the manifesto data, also these texts were very short.
-As aggregation per topics was not possible for this data, we aggregated the texts by splitting all texts into parts of 1000 words.
-In order to balance the classes we drew 50 random parts for each party.
-This procedure reduces the effect of text length and biased sampling of different parties when testing the classifier.
+Like the manifesto data, these texts are very short.
+As aggregation per topic was not possible for this data, we aggregated the texts by splitting all texts into parts of 1000 words.
 
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.
+Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.\footnote{We used the speeches from the 17th legislative period for the first task, as this legislature is already completed and offers more data.
+We had to use the speeches from the 18th legislative period for the second task, because we only have recent facebook posts.}
+% Frage: Wollen wir nicht schreiben, dass wir das erste auf 17 und 18 gerechnet haben, aber nur 17.
+reporten?
+Stimmt ja und macht die Ergebnisse vielleicht vertrauensvoller.
+Einen entsprechenden Satz hatte ich ja schonmal formuliert.
 
 \subsection{Predicting political party affiliation}
+\paragraph{... on in-domain-data}
 When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
@@
 
-\paragraph{Validation on manifesto data}
+\paragraph{... on out-of-domain-data}
+% Validation on manifesto data hat mich hier irritiert, weil es in dem Absatz ja auch um die facebook Daten geht. 
 For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
@@
 
-\paragraph{Length of texts influences prediction accuracy}
+\subsection{Length of texts influences prediction accuracy}
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
@@
 
-\subsection{Sentiment correlates with political power} 
+\subsection{The discriminatory effect between the parties}
+\label{sec:discrim_effect}
+An important question is, of course, whether the difference between  the parties relies on different policies or if the differentiating effect between the parties is based on other aspects of the text.
+To get more insight into what aspects are responsible for discrimanting between the parties, we analysed, first, whether the words used most commonly by each party are related to policies in line with the parties' ideologies and second, we examined whether sentiment has a discriminatory effect between government and opposition parties, too.
+
+\paragraph{Correlations between words and parties}
+We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
+Some unspecific stopwords are excluded.
+We find clear differences between the parties and these difference are in line with the parties ideologies.
+\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
+\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
+\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em staatssekretaerin, kanzlerin, bundestagsfraktion}) and some words related to cutting of expenses ({\em kuerzungen}).
+\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em wettbewerbsfaehigkeit, entwicklung}) and words related to security ({\em sicherheit, stabilitaet}).
+
+\paragraph{Sentiment correlates with political power} 
 We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
@@
 %
-\subsection{Correlations between words and parties}\label{sec:word_party_correlations}
-We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label. Some unspecific stopwords are excluded. 
-\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
-\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
-\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em Staatssekretaerin, Kanzlerin, Bundestagsfraktion}) and some words related to cutting of expenses ({\em Kuerzungen}).
-\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
-Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em Wettbewerbsf\"ahigkeit, Entwicklung}) and words related to security ({\em Sicherheit, Stabilitaet}).
 
@@
 Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
-These results encourage to use such systems as assistive technology for instance for human annotators in an active learning setting.
+These are encouraging results with respect to the use of such systems as assistive technology, for instance for human annotators in an active learning setting.
 
-We find a strong effect of text length and text domain on the generalization performance of the classifier, as also shown in other studies on similar data \cite{Yu2008, Hirst2014}.
+In line with other findings \cite{Yu2008, Hirst2014}, we find a strong effect of text length and text domain on the generalization performance of the classifier.
 The first effect, longer texts are easier to classify, makes intuitive sense.
-Also humans are worse at judging the political bias of shorter texts out of context: Benoit et al.
-(2015), for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
+Also humans are worse at judging the political bias of shorter texts out of context: \cite{Benoit.Forthcoming}, for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
 However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short.
-But political education would benefit from automatic analyses of these data streams; these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
+But political scientists can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
 \\
-The second effect, drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments.
-In the case of party affiliation prediction, the out-of-domain classification is on par or even better than the prediction accuracy on in-domain data.
+The second effect, i.e.
+the drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments.
+In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data.
 However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification.

### 2016-06-21 manuscript/poltext.tex
restructured results. merged bibtex file.

@@
 \documentclass[11pt]{article}
-%\usepackage{poltext2016}
+\usepackage{poltext2016}
 \usepackage{times}
@@
  Pola Lehmann\thanks{ ~{\tt pola.lehmann@wzb.eu} }\\
-  Wissenschaftszentrum Berlin für Sozialforschung\\
+%  WZB f\"ur Sozialforschung\\
 \And 
@@
  \And
-  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ TU Berlin
+  Sebastian Schelter \thanks{~\tt sebastian.schelter@tu-berlin.de}\\ 
+%  TU Berlin
 }
@@
 
-A major problem with automated text analyses is generalisation to other text domains than that on which the system was trained \cite{Slapin.2014}.
-Human coders have read enough texts of different domains to detect political bias in a variety of contexts and styles.
+A major problem with automated text analyses is generalisation to text domains other than that on which the system was trained \cite{Slapin.2014}. 
+%The political experts who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
 % Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme.
 Das ist wahrscheinlich wieder der Unterschied der Disziplinen, ihr müsst rechtfertigen, warum Computer es können, wenn Menschen es können, wir müssen in der Regel rechtgertigen warum Menschen es können, wenn Computer es besser machen.
-Machine learning algorithms however are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
+% Ich glaube, dass Computer es besser machen, wuerde niemand behaupten.
+Computer machen das genau so gut oder schlecht, wie die qualitaet der daten (die ja von menschen gelabeled oder selektiert werden) ihnen erlaubt.
+Der Algorithmus spiel auch ne rolle, aber das hauptproblem ist Daten und die koennen nur von Menschen kommen.
+Computer sind hilfreich als assistive technology, weil sie das ganze skalierbar machen.
+Machine learning algorithms are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
 Unfortunately good unbiased training data is difficult to obtain.
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.\footnote{We used the speeches from the 17th legislative period for the first task, as this legislature is already completed and offers more data.
-We had to use the speeches from the 18th legislative period for the second task, because we only have recent facebook posts.}
+Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.\footnote{We used the speeches from the 17th legislative period for the first task as this legislature is already completed and offers more data.
+Results for the 18th Bundestag are similar but omitted for brevity.
+We used the speeches of the 18th legislative period for the facebook posts as the posts were more recent.}
 % Frage: Wollen wir nicht schreiben, dass wir das erste auf 17 und 18 gerechnet haben, aber nur 17.
@@
 
-\subsection{Predicting political party affiliation}
-\paragraph{... on in-domain-data}
-When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
+\subsection{In-domain predictions}
+
+When predicting party affiliation on text data from the same domain that was used for training the model average precision and recall values of above 0.6 are obtained.
 The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
@@
 
-\paragraph{... on out-of-domain-data}
+\subsection{Out-of-domain predictions}
 % Validation on manifesto data hat mich hier irritiert, weil es in dem Absatz ja auch um die facebook Daten geht. 
-For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4 \autoref{tab:results_out-of-domain}.
-This drop in out of domain prediction accuracy is in line with previous findings \cite{Yu2008}.
-When testing the classifier on the facebook data (results not shown), this drop is similar.
-
-\subsection{Length of texts influences prediction accuracy}
-A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the strings to be classified, see also \autoref{sec:data}.
-In order to investigate whether this low out-of-domain prediction performance was due to the domain difference (parliament speech vs manifesto data) or due to the short length of the data points, the manifesto data was aggregated into political topics.
-The topic level results are shown in \autoref{tab:results_topic} and demonstrate that when the texts to be classified are sufficiently long and the word count statistics are sufficiently dense the classification performance on out-of-domain data can - at least in some cases - achieve reliable precision and recall values close to 1.0.
-This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
-
-\paragraph{Validation on facebook data}
+% Sorry, das war schlecht strukturiert - wollte validation als ueberschrift drinlassen, weil das von Dir war, Pola.
+For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4, see \autoref{tab:results_out-of-domain}.
 For the facebook post data a similar effect was observed.
-Short texts resulted in poor prediction accuracies of 0.51 on average.
-In addition classes were highly unbalanced in this setting since some parties have an order of magnitude more posts than others.
-In order to reduce the impact of text length and unbiased sampling and as there were no topic labels for the facebook posts, those texts were aggregated by sampling 50 texts of 1000 words each of the concatenated facebook posts of a party.
-The results are shown in \autoref{tab:results_fb} and indicate that also for these texts prediction accuracies comparable to the in-domain case can be achieved.
+The short texts resulted in poor prediction accuracies of 0.51 on average.
+In addition in this setting classes were highly unbalanced since some parties have an order of magnitude more posts than others.
 
@@
 \end{center}
+\end{table}
 
-\end{table}
+\subsection{Influence of text length on accuracy}
+A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to be classified, see also \autoref{sec:data}.
+In order to investigate the effect of text length the data was aggregated into longer texts.
+Manifesto data was aggregated into political topics.
+The topic level prediction results are shown in \autoref{tab:results_topic}.
+For all parties except for the SPD F1 scores of above 0.8 are obtained.
+As there were no topic labels for the facebook posts, those texts were aggregated by sampling 50 texts of 1000 words each of the concatenated facebook posts of a party.
+The results are shown in \autoref{tab:results_fb}.
+Also for these texts prediction accuracies comparable to the in-domain case can be achieved.
+This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
 
@@
 
-\paragraph{Policy change}
-Another explanation for misclassification could lie in the fact that parties change their policy positions.
-In order to investigate this confusion matrices were extracted for the predictions on the out-of-domain evaluation data for topic level predictions (see \autoref{tab:confusion_topic}).
-On the topic level results for the 17th legislative period are good except for the SPD, which the classifier often predicts as CDU/CSU.
+\paragraph{Misclassifications and policy change}
+An explanation for misclassifications could be that parties change their policy positions.
+This requires inspecting the specific data points that were misclassified.
+The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that on the topic level the SPD manifesto texts are often predicted as belonging to the CDU/CSU.
+A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
 One explanation for this misclassification could be policy changes.
-A similar effect can be observed in the 18th Bundestag when predicting facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
 \\
@@
 In \autoref{tab:results_binary_17} the results are shown for the 17th legislative period.
-While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on short quasi-sentences drops again close to chance performance.
+While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on manifesto data drops again close to chance performance.
 This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
-The authors report classification accuracies between 0.80 and 0.87, the accuracy in the 17th Bundestag was 0.85.
-A pronounced drop in performance was reported when applying the model to texts from a different domain (e.g.
-older texts or texts from another chamber), a similar effect as we find.
-Aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
+The authors report classification accuracies between 0.80 and 0.87 and find a pronounced drop in performance on texts from a different domain (e.g.
+older texts or texts from another chamber).
+In our results aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
 The drop in accuracy when applying the binary classifier on facebook data (aggregated as in the party affiliation case) was less pronounced, accuracies were above 0.70.
-These results indicate that there might be some signal common to plenary debates and the facebook data that is absent in the manifesto data.
-One such signal could be related to the fact that after the election, both the speakers in the parliament as well as the authors of the facebook posts  know which party is in the governmental coalition -- while the authors of the manifesto texts do not know that as these texts were written before the elections.
+
+\subsection{Discriminative features}
+\label{sec:discrim_effect}
+An important question is whether the difference between the features of each party relies on different policies or on on other aspects of the text.
+To get more insight into what aspects are discriminative for government membership and for parties we analysed, first, the effect of text sentiment and the text features most correlated with each party.
+
+% ich glaub der uebergang von dem sentiment in der letzten subsection passt ganz gut zu den sentiment analysen im naechsten
+% auch, weils beim sentiment noch um die binary classification geht, die gerade kam
+\paragraph{Sentiment correlates with political power}
+The drop in prediction accuracy in the government prediction task was more pronounced for manifesto texts than for facebook posts.
+What do facebook posts and plenary debates have in common?
+Both the speakers in the parliament as well as the authors of the facebook posts know which party is in the government -- while the authors of the manifesto texts do not know that as these texts were written before the elections.
 A language feature that could capture this is sentiment.
-
-\subsection{The discriminatory effect between the parties}
-\label{sec:discrim_effect}
-An important question is, of course, whether the difference between  the parties relies on different policies or if the differentiating effect between the parties is based on other aspects of the text.
-To get more insight into what aspects are responsible for discrimanting between the parties, we analysed, first, whether the words used most commonly by each party are related to policies in line with the parties' ideologies and second, we examined whether sentiment has a discriminatory effect between government and opposition parties, too.
-
-\paragraph{Correlations between words and parties}
-We quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
-Some unspecific stopwords are excluded.
-We find clear differences between the parties and these difference are in line with the parties ideologies.
-\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
-\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
-\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em staatssekretaerin, kanzlerin, bundestagsfraktion}) and some words related to cutting of expenses ({\em kuerzungen}).
-\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
-Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em wettbewerbsfaehigkeit, entwicklung}) and words related to security ({\em sicherheit, stabilitaet}).
-
-\paragraph{Sentiment correlates with political power} 
-We quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
+In order to assess this effect we quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
 Our results in  \autoref{tab:sentiments} show that government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
+Also previous studies find that text features discriminative in a two party system are not necessarily related to policies but more to language of defence and attack \cite{Hirst2014}.
 
@@
 %
+\paragraph{Correlations between words and parties}
+To determine other discriminative features we quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
+Some unspecific stopwords are excluded.
+We find clear differences between the parties and these difference are in line with the parties ideologies.
+\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
+\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
+\paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em staatssekretaerin, kanzlerin, bundestagsfraktion}) and some words related to cutting of expenses ({\em kuerzungen}).
+\paragraph{\bf Christian Democratic Union/Christian Social Union (CDU/CSU)}
+Often used words relate to a pro-economy attitude, such as competitiveness or (economic) development ({\em wettbewerbsfaehigkeit, entwicklung}) and words related to security ({\em sicherheit, stabilitaet}).
 
@@
 Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
-These are encouraging results with respect to the use of such systems as assistive technology, for instance for human annotators in an active learning setting.
+These results suggest that such systems could be helpful as assistive technology for instance for human annotators in an active learning setting.
 
@@
 However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short.
-But political scientists can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
+But political education can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
 \\
@@
 %
-\newpage
-\small{
 \bibliographystyle{plain}
 \bibliography{political_bias_prediction}
-}
 

### 2016-06-22 manuscript/poltext.tex
Added just very view minor changes

@@
 A major problem with automated text analyses is generalisation to text domains other than that on which the system was trained \cite{Slapin.2014}. 
-%The political experts who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
+Traditionally text analysis is done by political experts, who are annotating political text having in mind many context factors.
+There has, however, always been discussion about the reliability of such annotations.
+Automated analysis can solve this problem, but they are more fragile in regard to validity questions.
+%who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
 % Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme.
@@
 Computer sind hilfreich als assistive technology, weil sie das ganze skalierbar machen.
+% Ich versteh deinen Kommentar, aber wir mussten uns tatsächlich am Anfang des Projekts viel mit so Fragen rumschlagen, warum wir denn noch mit Menschen kodieren, wenn Computer das doch besser können.
+Eigentlich geht es da aber, glaube ich, um das Spannugnsgeflecht zwischen Reliabilität und Validität.
+Habe versucht das jetzt nochmal so umzuformulieren.
 Machine learning algorithms are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
@@
 In addition in this setting classes were highly unbalanced since some parties have an order of magnitude more posts than others.
+% Kann es sein, dass hier irgendwie zwei Worte zu viel drin sind? irgendwie verstehe ich den Satz jedenfalls gerade nicht ganz.
 
@@
 
-\paragraph{Misclassifications and policy change}
+\subsection{Misclassifications and policy change}
 An explanation for misclassifications could be that parties change their policy positions.
@@
 \label{sec:discrim_effect}
-An important question is whether the difference between the features of each party relies on different policies or on on other aspects of the text.
+An important question is whether the difference between the features of each party relies on different policies or on other aspects of the text.
 To get more insight into what aspects are discriminative for government membership and for parties we analysed, first, the effect of text sentiment and the text features most correlated with each party.

### 2016-06-22 manuscript/poltext.tex
paper polishing and some comments

@@
 \usepackage{pbox}
-\usepackage{color}
+\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\felix}[1]{\textcolor{green}{#1}}
+\newcommand{\ssc}[1]{\textcolor{magenta}{#1}}
 
@@
 Identifying such biases is crucial for political research as well as media consumers, especially when analysing the influence of the media on political discourse and vice versa.
-In this study we investigate to what extent political party affiliation can be predicted from textual content.
-Results indicate that automatic classification of political affiliation is possible with above chance accuracy also across text domains.
+In this study, we investigate to what extent political party affiliation can be predicted from textual content.
+Results indicate that automatic classification of political affiliation is possible with above chance accuracy, even across text domains.
 We propose methods to better interpret these results and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
@@
 %
-Analysis and classifications of political text is and has been a very important tool to generate political science data \cite{Benoit.Forthcoming}.
-Traditionally such classifications are done by experts, who read and label the text of interest.\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.} This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
-The growing field of automated text analysis, that allows the analysis of much more text in less time, is therefore of great interest to political scientists.
-Additionally automated text analyses allow for a more objective and replicable analysis of political text then human coders could achieve \cite{Benoit.2}.
+Analysis and classifications of political text is and has been a very important tool to generate political science data~\cite{Benoit.Forthcoming}.
+Traditionally, experts conduct such classifications by reading and labeling the text of interest\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.}.
+This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
+The growing field of automated text analysis, which allows for the analysis of much more text in less time, is therefore of great interest to political scientists.
+Additionally, automated text analyses allow for a more objective and replicable analysis of political text than the one which human coders could achieve~\cite{Benoit.2}.
 
-A major problem with automated text analyses is generalisation to text domains other than that on which the system was trained \cite{Slapin.2014}. 
-Traditionally text analysis is done by political experts, who are annotating political text having in mind many context factors.
+A major problem with automated text analyses is generalisation to text domains other than that on which the system has been trained~\cite{Slapin.2014}. 
+Traditionally, experts conduct text analysis, who consider many contextual factors during the annotation of political text.
 There has, however, always been discussion about the reliability of such annotations.
-Automated analysis can solve this problem, but they are more fragile in regard to validity questions.
+Automated analysis has the potential to solve this problem, but is more fragile with regard to questions of validity.
 %who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
@@
 Machine learning algorithms are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
-Unfortunately good unbiased training data is difficult to obtain.
-One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data as it is a large source of text that can be clearly associated with a party.
-Here we examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media.
-We discuss the effects of text length and domain shift of text data and investigate some potential reasons for the differences in classification performance.
+Unfortunately, good unbiased training data is difficult to obtain.
+One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party.
+We examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media.
+We discuss the effects of text length and domain shift of text data, and investigate some potential reasons for the differences in classification performance.
 
-We investigate predictions of the models with three strategies.
-First the model misclassifications are related to changes in party policies.
-Second sentiment analysis is used to investigate whether this aspect of language has discriminatory power.
-Third univariate measures of correlation between text features and party affiliation allow to relate the predictions to the kind of information that political experts use for interpreting texts.
+We investigate predictions of the models with three strategies: First, the model misclassifications are related to changes in party policies.
+Second, we use sentiment analysis to investigate whether this aspect of language has discriminatory power.
+Third, univariate measures of correlation between text features and party affiliation allow us to relate the predictions to the kind of information that political experts use for interpreting texts.
 
-In the following, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results} the results are discussed and \autoref{sec:conclusion} concludes with some interpretations of the results.
+In the following, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results}, we discuss results and \autoref{sec:conclusion} concludes with interpretations of the results.
 
@@
 %
-All experiments were run on publicly available data sets of German political texts and standard libraries for processing the text.
+We ran all experiments on publicly available data sets of German political texts, and applied standard libraries for processing the text.
 The following sections describe the details of data acquisition and feature extraction.
@@
 The texts from plenary debates were used to train a classifier and evaluate it on this in-domain data.
-The latter two data sources were used to test the generalization performance of the classifier on out-of-domain data.
+We employed the latter two data sources to test the generalization performance of the classifier on out-of-domain data.
 
 \paragraph{Parliament discussion data} Parliament texts are annotated with the respective party label.
-The protocols of plenary debates are available through the website of the German Bundestag \cite{bundestag}; an open source API was used to query the data in a cleaned and structured format \cite{bundestag-github}.
+The protocols of plenary debates are available through the website of the German Bundestag~\cite{bundestag}; we leveraged an open source API to query the data in a cleaned and structured format~\cite{bundestag-github}.
 Each uninterrupted part was treated as a separate speech.
@@
 \paragraph{Party manifesto data}
-The party manifesto text was taken from the Manifesto Corpus  \cite{manifesto}.
-The data released in this project mainly comprises the complete manifestos of all parties that have won seats at a national election.
+The party manifesto text originates from the Manifesto Corpus~\cite{manifesto}.
+The data released in this project mainly comprises the complete manifestos of all parties that have won seats in a national election.
 Each statement or {\em quasi-sentence}\footnote{A quasi-sentence has the length of an argument.
 It is never longer than one sentence.} is annotated with one of 56 policy issue categories.
-Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see \cite{leftright}.
+Examples for the policy categories are {\em welfare state expansion, welfare state limitation, democracy, equality}; for a complete list and detailed explanations on how the annotators were instructed see~\cite{leftright}.
 Each quasi-sentence has two types of labels: the party affiliation and the manually assigned policy issue aimed at in each quasi-sentence.
 The length of each annotated statement in the party manifestos is rather short.
-The median length is 95 characters or 12 words.\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}
-In order to increase the length of the texts used for classification the policy labels were used to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
-So in this setting each party had just one data point for each of the topics.
+The median length is 95 characters or 12 words\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}.
+In order to increase the length of the texts for classification, we employed the policy labels to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
+In this setting, each party had just one data point for each of the topics.
 
 \paragraph{Facebook post data}
-For each party we crawled their facebook pages \cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post text, excluding all comments and other information.
+We crawled the facebook page of each party~\cite{gruene-fb, spd-fb, cducsu-fb, linke-fb} and extracted the post texts, excluding all comments and other information.
 Like the manifesto data, these texts are very short.
@@
 \subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-All text data was tokenised and transformed into bag-of-word (BOW) vectors as implemented in scikit-learn \cite{scikit-learn}.
-Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and different cutoffs for discarding too frequent and too infrequent words.
+We tokenised all text data tokenised and transformed it into bag-of-word (BOW) vectors as implemented in scikit-learn~\cite{scikit-learn}.
+Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and different cut-offs for discarding too frequent and too infrequent words.
 
 \section{Classification Model and Training}\label{sec:model}
-Bag-of-words feature vectors were used to train a multinomial logistic regression model.
-Let $y\in\{1,2,\dots,K\}$ be the true party affiliation and $\vec{w}_1,\dots,\vec{w}_K\in\R^{d}$ the weight vectors associated with the $k$th party then the party affiliation estimate is modelled as
+We leveraged bag-of-words feature vectors to train a multinomial logistic regression model.
+Let $y\in\{1,2,\dots,K\}$ be the true party affiliation and $\vec{w}_1,\dots,\vec{w}_K\in\R^{d}$ the weight vectors associated with the $k$th party.
+Then the party affiliation estimate is modelled as
 \begin{eqnarray}\label{eq:logreg_multiclass}
@@
 \subsection{Optimisation of Model Parameters}\label{sec:crossvalidation}
-The model pipeline contained a number of  hyperparameters that were optimised using gridsearch cross-validation.
-To this end the parliament speech data was split into training and validation set in a 90\%/10\% ratio; the pipeline was trained with each parameter setting on the training set and its performance validated on the validation set.
-The parameters of the best performing model were then used to train a model on the training and validation set data.
+The model pipeline contained a number of hyperparameters that we optimised using gridsearch cross-validation.
+To this end, we spli the parliament speech data into training and validation set in a 90\%/10\% ratio; we trained the pipeline with each parameter setting on the training set and validated its performance on the validation set.
+We chose the parameters of the best performing model to train a model on the training and validation set data.
 None of the data in the separately held back in-domain test data nor the out-of-domain test data sets was used for this hyperparameter optimization.
@@
 \subsection{Sentiment analysis}\label{sec:sentiment_analysis_methods}
-A publicly available key word list was used to extract sentiments \cite{remquahey2010}.
+We extracted sentiments via a publicly available key word list~\cite{remquahey2010}.
 A sentiment vector $\vec{s}\in\R^d$ was constructed from the sentiment polarity values in the sentiment dictionary.
-The sentiment index used for attributing positive or negative sentiment to a text was computed as the cosine similarity between BOW vectors and sentiment vector.
+We compute the sentiment index for attributing positive or negative sentiment to a text as the cosine similarity between BOW vectors and sentiment vector.
 
@@
 Interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
-Thus direct interpretation of the model coefficients $\vec{w}_k$ is problematic, see also \cite{Zien2009, Haufe2013}.
-In order to allow for better interpretation of the predictions and to assess which features are discriminative, correlation coefficients between each word and the party affiliation label were computed.
+Thus direct interpretation of the model coefficients $\vec{w}_k$ is problematic, see also~\cite{Zien2009, Haufe2013}.
+In order to allow for better interpretation of the predictions and to assess which features are discriminative, we computed correlation coefficients between each word and the party affiliation label.
 
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-Predictions compared with the manifesto data were done using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag.\footnote{We used the speeches from the 17th legislative period for the first task as this legislature is already completed and offers more data.
+Predictions compared with the manifesto data were computed using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag\footnote{We leverage the speeches from the 17th legislative period for the first task as this legislature is already completed and offers more data.
 Results for the 18th Bundestag are similar but omitted for brevity.
-We used the speeches of the 18th legislative period for the facebook posts as the posts were more recent.}
+We employ the speeches of the 18th legislative period for the facebook posts as the posts were more recent.}.
 % Frage: Wollen wir nicht schreiben, dass wir das erste auf 17 und 18 gerechnet haben, aber nur 17.
@@
 
-When predicting party affiliation on text data from the same domain that was used for training the model average precision and recall values of above 0.6 are obtained.
-The evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag are listed in \autoref{tab:results_in-domain}.
-These results are comparable to those of \cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.
+When predicting party affiliation on text data from the same domain that was used for training the model, average precision and recall values of above 0.6 are obtained.
+We list the evaluation results for the political party affiliation prediction on in-domain data (held-out parliamentary speech text) for the 17th Bundestag in~\autoref{tab:results_in-domain}.
+These results are comparable to those of~\cite{Hirst2014} who report a classification accuracy of 0.61 on a five class problem predicting party affiliation in the European parliament.
 
@@
 % Sorry, das war schlecht strukturiert - wollte validation als ueberschrift drinlassen, weil das von Dir war, Pola.
-For out-of domain data obtained from manifesto data the models yield significantly lower precision and recall values between 0.3 and 0.4, see \autoref{tab:results_out-of-domain}.
-For the facebook post data a similar effect was observed.
+For out-of domain data obtained from manifesto data, the models yield significantly lower precision and recall values between 0.3 and 0.4, see~\autoref{tab:results_out-of-domain}.
+We observe a similar effect for the facebook post data.
 The short texts resulted in poor prediction accuracies of 0.51 on average.
-In addition in this setting classes were highly unbalanced since some parties have an order of magnitude more posts than others.
+Additionally, classes were highly unbalanced in this setting, since some parties have an order of magnitude more posts than others.
 % Kann es sein, dass hier irgendwie zwei Worte zu viel drin sind? irgendwie verstehe ich den Satz jedenfalls gerade nicht ganz.
@@
 {\bf Out-of-domain} classification performance (topic level) on {\bf manifesto data}.
-Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}) the predictions made on topic level are more reliable.}
+Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}), the predictions made on topic level are more reliable.}
 \begin{center}
@@
 \subsection{Influence of text length on accuracy}
-A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to be classified, see also \autoref{sec:data}.
-In order to investigate the effect of text length the data was aggregated into longer texts.
-Manifesto data was aggregated into political topics.
-The topic level prediction results are shown in \autoref{tab:results_topic}.
-For all parties except for the SPD F1 scores of above 0.8 are obtained.
-As there were no topic labels for the facebook posts, those texts were aggregated by sampling 50 texts of 1000 words each of the concatenated facebook posts of a party.
-The results are shown in \autoref{tab:results_fb}.
-Also for these texts prediction accuracies comparable to the in-domain case can be achieved.
-This increase is in line with previous findings on the influence of text length on political bias prediction accuracy \cite{Hirst2014}.
+A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to classify, see also~\autoref{sec:data}.
+In order to investigate the effect of text length, we aggregated the data into longer texts, and group Manifesto data into political topics.
+\autoref{tab:results_topic} shows the topic level prediction results.
+We obtain F1 scores of above 0.8 for all parties except for the SPD .
+As the facebook posts lacked topic labels, we conducted the aggregation of these texts by sampling 50 texts of 1000 words from each of the concatenated facebook posts of a party.
+\ssc{this is a bit unclear, maybe another sentence with details would be good} The results are shown in \autoref{tab:results_fb}.
+Prediction accuracies comparable to the in-domain case can also be achieved for these texts.
+This increase is in line with previous findings on the influence of text length on political bias prediction accuracy~\cite{Hirst2014}.
 
@@
 \subsection{Misclassifications and policy change}
-An explanation for misclassifications could be that parties change their policy positions.
-This requires inspecting the specific data points that were misclassified.
-The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that on the topic level the SPD manifesto texts are often predicted as belonging to the CDU/CSU.
-A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
+An explanation for misclassifications could be that parties change their policy positions over time.
+This requires us to inspect the specific data points that were misclassified.
+The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level.
+A similar effect can be observed in the 18th Bundestag, when we predict for both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
 One explanation for this misclassification could be policy changes.
-\\
+\ssc{That sounds too much like a speculation, can we maybe give one example, e.g.
+that the conservatives strengthened environmental issues?} \\
 
@@
 \subsection{Predicting government status}\label{sec:sentiment_result}
-For a better comparison with other studies that predict party affiliation in a two party system we also trained a model on government membership labels.
-In \autoref{tab:results_binary_17} the results are shown for the 17th legislative period.
+We also trained a model on government membership labels, in order to a better compare against other studies that predict party affiliation in a two party system.
+We show the results for the 17th legislative period in \autoref{tab:results_binary_17}.
 While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on manifesto data drops again close to chance performance.
-This is in line with the results on binary classification of political bias in the Canadian parliament \cite{Yu2008}.
-The authors report classification accuracies between 0.80 and 0.87 and find a pronounced drop in performance on texts from a different domain (e.g.
+This is in line with results on binary classification of political bias in the Canadian parliament~\cite{Yu2008}.
+The authors report classification accuracies between 0.80 and 0.87, and find a pronounced drop in performance on texts from a different domain (e.g.
 older texts or texts from another chamber).
-In our results aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
-The drop in accuracy when applying the binary classifier on facebook data (aggregated as in the party affiliation case) was less pronounced, accuracies were above 0.70.
+In our results, the aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
+The drop in accuracy of the binary classifier on facebook data (aggregated analogous to the party affiliation case) was less pronounced, accuracies were above 0.70.
 
@@
 \label{sec:discrim_effect}
-An important question is whether the difference between the features of each party relies on different policies or on other aspects of the text.
-To get more insight into what aspects are discriminative for government membership and for parties we analysed, first, the effect of text sentiment and the text features most correlated with each party.
+An important question is whether the difference between the features of each party stems from different policies or from other aspects of the text.
+In order to get more insight into what aspects are discriminative for government membership and for parties, we first analysed the effect of text sentiment and the textual features most correlated with each party.
+\ssc{this sentence is too long, but I don't have a good idea how to split it :D}
 
@@
 What do facebook posts and plenary debates have in common?
-Both the speakers in the parliament as well as the authors of the facebook posts know which party is in the government -- while the authors of the manifesto texts do not know that as these texts were written before the elections.
+Both the speakers in the parliament as well as the authors of the facebook posts are aware of which party is in the government -- while the authors of the manifesto texts lack this knowledge, as these texts were written before the elections.
 A language feature that could capture this is sentiment.
-In order to assess this effect we quantified the correlation between government membership as well as number of seats in the parliament with text sentiment.
-Our results in  \autoref{tab:sentiments} show that government membership correlates with positive sentiment with a correlation coefficient of 0.98 and the number of seats correlates with 0.89.
-Also previous studies find that text features discriminative in a two party system are not necessarily related to policies but more to language of defence and attack \cite{Hirst2014}.
+In order to assess this effect, we quantified the correlation between government membership as well as the number of seats in the parliament with text sentiment.
+Our results in \autoref{tab:sentiments} show that government membership correlates nearly perfect with positive sentiment, giving a correlation coefficient of 0.98, and the number of seats correlates with 0.89.
+Previous studies also find that text features discriminative in a two party system are not necessarily related to policies but more to language of defense and attack~\cite{Hirst2014}.
 
@@
 \label{tab:sentiments}
-Correlation coefficient between average sentiment of political speeches of a party in the german Bundestag with two indicators of political power, a) membership in the government and b) the number of seats a party occupies in the parliament.
+Correlation coefficient between the average sentiment of political speeches of a party in the german Bundestag with two indicators of political power: a) membership in the government and b) the number of seats a party occupies in the parliament.
 }
@@
 \paragraph{Correlations between words and parties}
-To determine other discriminative features we quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
-Some unspecific stopwords are excluded.
-We find clear differences between the parties and these difference are in line with the parties ideologies.
-\paragraph{\bf Left party (linke)} Often used words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
-\paragraph{\bf Green party (gruene)} Use words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
+In order to determine further discriminative features, we quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
+Unspecific stopwords are excluded.
+We find clear differences between the parties, which are in line with the parties ideologies.
+\paragraph{\bf Left party (linke)} Frequent words include referrals to big companies ({\em konzerne}) and their profits ({\em profite}), the working class {\em beschaeftigte}, the social welfare program {\em hartz iv} as well as war ({\em krieg}).
+\paragraph{\bf Green party (gruene)} Uses words related to environmental damage ({\em klimaschaedlichen}), exploited low wage employees ({\em leiharbeitskraefte}) and pensions ({\em garantierente}).
 \paragraph{\bf Social Democratic Party (SPD)} Uses mostly unspecific words related to the parliament and governmental processes ({\em staatssekretaerin, kanzlerin, bundestagsfraktion}) and some words related to cutting of expenses ({\em kuerzungen}).
@@
 \section{Conclusions and Limitations}\label{sec:conclusion}
-Evaluating classifiers trained on parliament speeches we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
+During our evaluation of classifiers trained on parliament speeches, we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
 These results suggest that such systems could be helpful as assistive technology for instance for human annotators in an active learning setting.
 
-In line with other findings \cite{Yu2008, Hirst2014}, we find a strong effect of text length and text domain on the generalization performance of the classifier.
-The first effect, longer texts are easier to classify, makes intuitive sense.
-Also humans are worse at judging the political bias of shorter texts out of context: \cite{Benoit.Forthcoming}, for example, found in an experiment where experts where differentiating whether specific sentences were dealing with economic or social policy only about 35\% agreement between all expert coders.
-However short texts are a realistic challenge for automated political bias prediction systems: Often political texts from social media data and other web media are indeed very short.
-But political education can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions yet are difficult to analyse with human annotators due to the large volume of data.
+In line with other findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier.
+The first effect, that longer texts are easier to classify, intuitively makes sense.
+Humans are similarly worse at judging the political bias of shorter texts out of context.
+\cite{Benoit.Forthcoming}, for example, conducted an experiment where experts were differentiating whether specific sentences were dealing with economic or social policy.
+They observed only about 35\% agreement between all expert coders.
+However, short texts are a realistic \ssc{I don't understand this word here} challenge for automated political bias prediction systems: Political texts from social media data and other web media are indeed very short on many occasions.
+But political education can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions, even though they are difficult to analyse with human annotators due to the large volume of data.
 \\
@@
 However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification.
-We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
+\ssc{sentence too long} We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
 

### 2016-06-22 manuscript/poltext.tex
polished. slowly converging, seemingly.

@@
 A major problem with automated text analyses is generalisation to text domains other than that on which the system has been trained~\cite{Slapin.2014}. 
-Traditionally, experts conduct text analysis, who consider many contextual factors during the annotation of political text.
+%Traditionally, experts conduct text analysis, who consider many contextual factors during the annotation of political text.
 There has, however, always been discussion about the reliability of such annotations.
 Automated analysis has the potential to solve this problem, but is more fragile with regard to questions of validity.
-%who judge and label political texts have read enough content of different domains to detect political bias in a variety of contexts and styles. 
-% Den vorangegangenen Satz würde ich streichen, weil ich mit der Aussage nicht übereinstimme.
-Das ist wahrscheinlich wieder der Unterschied der Disziplinen, ihr müsst rechtfertigen, warum Computer es können, wenn Menschen es können, wir müssen in der Regel rechtgertigen warum Menschen es können, wenn Computer es besser machen.
-% Ich glaube, dass Computer es besser machen, wuerde niemand behaupten.
-Computer machen das genau so gut oder schlecht, wie die qualitaet der daten (die ja von menschen gelabeled oder selektiert werden) ihnen erlaubt.
-Der Algorithmus spiel auch ne rolle, aber das hauptproblem ist Daten und die koennen nur von Menschen kommen.
-Computer sind hilfreich als assistive technology, weil sie das ganze skalierbar machen.
 % Ich versteh deinen Kommentar, aber wir mussten uns tatsächlich am Anfang des Projekts viel mit so Fragen rumschlagen, warum wir denn noch mit Menschen kodieren, wenn Computer das doch besser können.
@@
 Habe versucht das jetzt nochmal so umzuformulieren.
-Machine learning algorithms are naturally prone to poor generalisation performance if the training data is biased towards one text domain.
+% hm.
+also validitaet is ja schon im abschnitt davor angesprochen - ich war ehrlich gesagt auch erstaunt, wie das labeling gemacht wurde.
+um validitaet abschaetzen zu koennen braucht man ja mindestens 3 labels pro datenpunkt.
+den paragraph mit generalisierung einzuleiten und dann die validitaet nochmal nur so anzureissen ohne darauf einzugehn wieso fand ich nich so fluessig.
+der eigentliche punkt hier war ja generalisierung, und da haben menschen den vorteil, dass manche menschen heterogenere samples gelesen haben im lauf ihres lebens als die algorithmen von uns kriegen.
+ich hab versucht das wieder klarer rauszuarbeiten.
+While political experts have read texts of different domains and are able to detect political bias in a variety of contexts and styles, machine learning algorithms are prone to poor generalisation performance across text domains if the training data is biased towards one domain only.
 Unfortunately, good unbiased training data is difficult to obtain.
 One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party.
-We examine to what extent models trained on this data can be generalised to other text domains, such as party manifestos and texts from social media.
+We examine to what extent models trained on this data can generalise their predictions to other text domains, such as party manifestos and texts from social media.
 We discuss the effects of text length and domain shift of text data, and investigate some potential reasons for the differences in classification performance.
@@
 We obtain F1 scores of above 0.8 for all parties except for the SPD .
-As the facebook posts lacked topic labels, we conducted the aggregation of these texts by sampling 50 texts of 1000 words from each of the concatenated facebook posts of a party.
-\ssc{this is a bit unclear, maybe another sentence with details would be good} The results are shown in \autoref{tab:results_fb}.
+As the facebook posts lacked topic labels, we conducted the aggregation of these texts by first concatenating all facebook posts of a party into one long text; this text was then partitioned into segments of 1000 words each.
+For each party 50 random segments were selected for classification.
+The results are shown in \autoref{tab:results_fb}.
 Prediction accuracies comparable to the in-domain case can also be achieved for these texts.
@@
 The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level.
-A similar effect can be observed in the 18th Bundestag, when we predict for both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
-One explanation for this misclassification could be policy changes.
-\ssc{That sounds too much like a speculation, can we maybe give one example, e.g.
-that the conservatives strengthened environmental issues?} \\
+A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
+While we cannot list examples due to space limitations, we emphasize that these misclassifications can be related to policy changes of parties.
+\\
+% there were concrete examples (17th bundestag social reforms -> SPD mixed up with CDU/CSU, 18th bundestag fukushima -> CDU/CSU mixed up with green party), which can not be included due to space constraints.
+we might leave this out then?
 
@@
 An important question is whether the difference between the features of each party stems from different policies or from other aspects of the text.
-In order to get more insight into what aspects are discriminative for government membership and for parties, we first analysed the effect of text sentiment and the textual features most correlated with each party.
-\ssc{this sentence is too long, but I don't have a good idea how to split it :D}
+To address this point we analysed features that are discriminative for government membership and for parties, respectively.
 
@@
 What do facebook posts and plenary debates have in common?
-Both the speakers in the parliament as well as the authors of the facebook posts are aware of which party is in the government -- while the authors of the manifesto texts lack this knowledge, as these texts were written before the elections.
+In contrast to the authors of manifestos both the speakers in the parliament as well as the authors of facebook posts know which party is in the government.
 A language feature that could capture this is sentiment.
-In order to assess this effect, we quantified the correlation between government membership as well as the number of seats in the parliament with text sentiment.
-Our results in \autoref{tab:sentiments} show that government membership correlates nearly perfect with positive sentiment, giving a correlation coefficient of 0.98, and the number of seats correlates with 0.89.
+Indeed our results in \autoref{tab:sentiments} show that positive sentiment strongly correlates with government membership and the number of seats in the parliament.
 Previous studies also find that text features discriminative in a two party system are not necessarily related to policies but more to language of defense and attack~\cite{Hirst2014}.
@@
 
-In line with other findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier.
+In line with previous findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier.
 The first effect, that longer texts are easier to classify, intuitively makes sense.
-Humans are similarly worse at judging the political bias of shorter texts out of context.
-\cite{Benoit.Forthcoming}, for example, conducted an experiment where experts were differentiating whether specific sentences were dealing with economic or social policy.
+Also humans are challenged when judging the political bias of shorter texts out of context \cite{Benoit.Forthcoming}.
+%, for example, conducted an experiment where experts were differentiating whether specific sentences were dealing with economic or social policy.
 They observed only about 35\% agreement between all expert coders.
-However, short texts are a realistic \ssc{I don't understand this word here} challenge for automated political bias prediction systems: Political texts from social media data and other web media are indeed very short on many occasions.
-But political education can benefit from automatic analyses of these data streams, as these media have strong influence on public opinions, even though they are difficult to analyse with human annotators due to the large volume of data.
-\\
+% zu wenig platz fuer den satz, aber glaube, der punkt kommt auch so rueber: kurzer text is schwierig auf fuer menschen. 
+% re challenge: das soll hier heissen, dass das zwar schwierig aber realistisch ist, weil mans ja machen muss in vielen kontexten. 
+% ich hab versucht, das klarer zu machen, indem ich das in dem satz oben auch erwaehne. 
+However, short texts are a realistic challenge for automated political bias prediction systems: Political texts from social media data and other web sources are indeed often very short and hence difficult to analyse for both human annotators and algorithms.
+Both political education and science can benefit from automatic analyses of these very data streams, as these media have strong influence on public opinions yet cannot be analysed by humans alone, due to the volume of data.
+
 The second effect, i.e.
-the drop in generalization performance on out-of-domain data, can be alleviated by aggregating texts into longer segments.
+the drop in generalization performance on out-of-domain data, appears to be correlated to the first one: it can be alleviated in some cases by aggregating texts into longer segments.
 In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data.
-However in the binary classification setting (government membership), aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain classification.
-\ssc{sentence too long} We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
+However in the binary classification setting (government membership prediction), text aggregation does not help as much: Aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain accuracy drop.
+We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.
 

### 2016-06-24 manuscript/poltext.tex
added the policy change section

@@
 A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to classify, see also~\autoref{sec:data}.
-In order to investigate the effect of text length, we aggregated the data into longer texts, and group Manifesto data into political topics.
+In order to investigate the effect of text length, we aggregated the data into longer texts, and group manifesto data into political topics.
 \autoref{tab:results_topic} shows the topic level prediction results.
@@
 An explanation for misclassifications could be that parties change their policy positions over time.
-This requires us to inspect the specific data points that were misclassified.
+For this task we are predicting the manifestos which have been written prior to the debates on the basis of the debates.
+If a party adapts positions from another party during the legislative period this can, thus, leads to missclassifications of the manifesto text.
+The results seem to point in this direction.
 The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level.
-A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
-While we cannot list examples due to space limitations, we emphasize that these misclassifications can be related to policy changes of parties.
-\\
+This was the the legistlative period when the CDU under chancellor Merkel was making a strong left move with respect to socioeconomic questions.\\
 % there were concrete examples (17th bundestag social reforms -> SPD mixed up with CDU/CSU, 18th bundestag fukushima -> CDU/CSU mixed up with green party), which can not be included due to space constraints.

### 2016-06-25 htmlText.py
Proofread by native speaker.

@@ -1,0 +1,22 @@
+import urllib
+from bs4 import BeautifulSoup
+
+def get_text_from_url(url)
+    html = urllib.urlopen(url).read()
+    soup = BeautifulSoup(html)
+
+    # kill all script and style elements
+    for script in soup(["script", "style"]):
+        script.extract()    # rip it out
+
+    # get text
+    text = soup.get_text()
+
+    # break into lines and remove leading and trailing space on each
+    lines = (line.strip() for line in text.splitlines())
+    # break multi-headlines into a line each
+    chunks = (phrase.strip() for line in lines for phrase in line.split("  "))
+    # drop blank lines
+    text = '\n'.join(chunk for chunk in chunks if chunk)
+
+    return text.encode('utf-8')

### 2016-06-25 manuscript/poltext.tex
Proofread by native speaker.

--- 2016-06-22 manuscript/poltext.tex
@@
 \usepackage{pbox}
-\usepackage[dvipsnames]{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 
@@
 \begin{abstract}
-Every day large amounts of text are produced in public discourse.
+Every day a large amount of text is produced during public discourse.
 Some of this text is produced by actors whose political colour is very obvious.
-But many actors cannot be clearly attributed to a political party, yet their statements might be biased towards a specific party.
-Identifying such biases is crucial for political research as well as media consumers, especially when analysing the influence of the media on political discourse and vice versa.
-In this study, we investigate to what extent political party affiliation can be predicted from textual content.
-Results indicate that automatic classification of political affiliation is possible with above chance accuracy, even across text domains.
-We propose methods to better interpret these results and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
+However, though many actors cannot clearly be associated with a political party, their statements may be biased towards a specific party.
+Identifying such biases is crucial for political research as well as for media consumers, especially when analysing the influence of the media on political discourse and vice versa.
+In this study, we investigate the extent to which political party affiliation can be predicted from textual content.
+Results indicate that automated classification of political affiliation is possible with an accuracy better than chance, even across different text domains.
+We propose methods to better interpret these results, and find that features not related to political policies, such as speech sentiment, can be discriminative and thus exploited by text analysis models.
 \end{abstract}
@@
 Analysis and classifications of political text is and has been a very important tool to generate political science data~\cite{Benoit.Forthcoming}.
-Traditionally, experts conduct such classifications by reading and labeling the text of interest\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.}.
-This is, however, a very time consuming task and thus sets various limits to the possible amount of data that a few experts can analyze.
+Traditionally, experts conduct such classifications by reading and labelling the text of interest\footnote{See for example the Manifesto Project, the Comparative Agendas Project or Poltext.}.
+This is, however, a very time consuming task and thus sets various limits on the possible amount of data that a few experts can analyse.
 The growing field of automated text analysis, which allows for the analysis of much more text in less time, is therefore of great interest to political scientists.
-Additionally, automated text analyses allow for a more objective and replicable analysis of political text than the one which human coders could achieve~\cite{Benoit.2}.
+Additionally, automated text analyses allow for a more objective and replicable analysis of political text than human coders can achieve~\cite{Benoit.2}.
 
@@
 ich hab versucht das wieder klarer rauszuarbeiten.
-While political experts have read texts of different domains and are able to detect political bias in a variety of contexts and styles, machine learning algorithms are prone to poor generalisation performance across text domains if the training data is biased towards one domain only.
+% Ja,sorry, hatte nicht nochmal gelesen was davor stand. Finde es jetzt gut so.
+While political experts can read texts from different domains and are able to detect political bias appearing in a variety of contexts and styles, machine learning algorithms are prone to poor performance generalisation across text domains if the training data is biased towards one domain only.
 Unfortunately, good unbiased training data is difficult to obtain.
-One of the best sources for automated political text analysis systems are plenary debates of the parliament: Many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party.
+One of the best sources for automated political text analysis systems are plenary debates of the parliament: many studies are based on this type of data, as it consists of a large source of text that can be clearly associated with a party.
 We examine to what extent models trained on this data can generalise their predictions to other text domains, such as party manifestos and texts from social media.
-We discuss the effects of text length and domain shift of text data, and investigate some potential reasons for the differences in classification performance.
+We discuss the effects of text length and domain shifts of text data, and investigate some possible reasons for the differences in classification performance.
 
-We investigate predictions of the models with three strategies: First, the model misclassifications are related to changes in party policies.
+We investigate the predictions of the models with three strategies: first, we test the influence of text length on the prediction accuracy.
 Second, we use sentiment analysis to investigate whether this aspect of language has discriminatory power.
@@
 
-In the following, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures; in \autoref{sec:results}, we discuss results and \autoref{sec:conclusion} concludes with interpretations of the results.
+In this article, \autoref{sec:data} gives an overview of the data acquisition and preprocessing methods, \autoref{sec:model} presents the model, training and evaluation procedures, in \autoref{sec:results} we discuss results and \autoref{sec:conclusion} concludes with interpretations of the results.
 
@@
 %
-We ran all experiments on publicly available data sets of German political texts, and applied standard libraries for processing the text.
+We ran all experiments using publicly available data sets of German political texts, and applied standard libraries for processing the text.
 The following sections describe the details of data acquisition and feature extraction.
@@
 \subsection{Data}
-Annotated political text data was obtained from three sources: a) the plenary debates held in the German parliament ({\em Bundestag}) b) all manifesto texts of parties winning seats in the election to the German parliament and c) facebook page posts of all parties.
+Annotated political text data was obtained from three sources: a) the plenary debates held in the German parliament ({\em Bundestag}) b) all manifestos of parties winning seats in the election to the German parliament and c) facebook posts from all parties.
 The texts from plenary debates were used to train a classifier and evaluate it on this in-domain data.
-We employed the latter two data sources to test the generalization performance of the classifier on out-of-domain data.
+We employed the latter two data sources to test the generalisation performance of the classifier on out-of-domain data.
 
@@
 The length of each annotated statement in the party manifestos is rather short.
-The median length is 95 characters or 12 words\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}.
-In order to increase the length of the texts for classification, we employed the policy labels to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
+The median length is 95 characters, or 12 words\footnote{The longest statement is 522 characters (65 words) long, the 25\%/50\%/75\% percentiles are 63/95/135 characters or 8/12/17 words, respectively.}.
+In order to increase the length of the texts for classification, we used the policy labels to aggregate the data into the following topics: {\em External Relations, Freedom and Democracy, Political System, Economy, Welfare and Quality of Life, Fabric of Society, Social Groups}.
 In this setting, each party had just one data point for each of the topics.
@@
 
-\subsection{Bag-of-Words Vectorization}\label{sec:bow-vectorization}
-We tokenised all text data tokenised and transformed it into bag-of-word (BOW) vectors as implemented in scikit-learn~\cite{scikit-learn}.
-Several options for BOW vectorizations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and different cut-offs for discarding too frequent and too infrequent words.
+\subsection{Bag-of-Words Vectorisation}\label{sec:bow-vectorization}
+We tokenised all text data and transformed it into bag-of-word (BOW) vectors as implemented in scikit-learn~\cite{scikit-learn}.
+Several options for BOW vectorisations were tried, including term-frequency-inverse-document-frequency normalisation, n-gram patterns up to size $n=3$ and different cut-offs for discarding words which were too frequent or infrequent.
 
@@
 The model pipeline contained a number of hyperparameters that we optimised using gridsearch cross-validation.
-To this end, we spli the parliament speech data into training and validation set in a 90\%/10\% ratio; we trained the pipeline with each parameter setting on the training set and validated its performance on the validation set.
+To this end, we split the parliament speech data into training and validation sets in a 90\%/10\% ratio; we trained the pipeline with each parameter setting on the training set and validated its performance on the validation set.
 We chose the parameters of the best performing model to train a model on the training and validation set data.
-None of the data in the separately held back in-domain test data nor the out-of-domain test data sets was used for this hyperparameter optimization.
+None of the data in the separately held back in-domain test data nor the out-of-domain test data sets was used for this hyperparameter optimisation.
 %
@@
 \subsection{Interpreting bag-of-words models}\label{sec:correlations_methods}
-Interpreting coefficients of linear models (independent of the regularizer used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
+Interpreting coefficients of linear models (independent of the regulariser used) implicitly assumes uncorrelated features; this assumption is violated by the text data used in this study.
 Thus direct interpretation of the model coefficients $\vec{w}_k$ is problematic, see also~\cite{Zien2009, Haufe2013}.
@@
 The following section gives an overview of the results for all political bias prediction tasks. 
-Predictions compared with the manifesto data were computed using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag\footnote{We leverage the speeches from the 17th legislative period for the first task as this legislature is already completed and offers more data.
+Predictions compared with the manifesto data were computed using models trained on texts from the 17th Bundestag, predictions obtained for facebook post texts were computed with models trained on the 18th Bundestag\footnote{We leveraged the speeches from the 17th legislative period for the first task as this legislature is already completed and offers more data.
 Results for the 18th Bundestag are similar but omitted for brevity.
@@
 % Sorry, das war schlecht strukturiert - wollte validation als ueberschrift drinlassen, weil das von Dir war, Pola.
-For out-of domain data obtained from manifesto data, the models yield significantly lower precision and recall values between 0.3 and 0.4, see~\autoref{tab:results_out-of-domain}.
+For out-of-domain data obtained from manifesto data, the models yield significantly lower precision and recall values between 0.3 and 0.4, see~\autoref{tab:results_out-of-domain}.
 We observe a similar effect for the facebook post data.
@@
 {\bf Out-of-domain} classification performance (topic level) on {\bf manifesto data}.
-Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}), the predictions made on topic level are more reliable.}
+Compared to quasi-sentence level predictions (\autoref{tab:results_out-of-domain}), the predictions made on the topic level are more reliable.}
 \begin{center}
@@
 \subsection{Influence of text length on accuracy}
-A main factor that made the prediction on the out-of-domain prediction task particularly difficult is the short length of the texts to classify, see also~\autoref{sec:data}.
-In order to investigate the effect of text length, we aggregated the data into longer texts, and group Manifesto data into political topics.
+A key factor that made the prediction in the out-of-domain prediction task particularly difficult was the short length of the texts to classify, see also~\autoref{sec:data}.
+In order to investigate the effect of text length, we aggregated the data into longer texts, and grouped manifesto data into political topics.
 \autoref{tab:results_topic} shows the topic level prediction results.
@@
 
-\subsection{Misclassifications and policy change}
-An explanation for misclassifications could be that parties change their policy positions over time.
-This requires us to inspect the specific data points that were misclassified.
+\subsection{Misclassification and policy change}
+Automatic political text analysis requires a profound understanding of the models used.
+One way to better understand these models is to inspect the misclassifications of a model: a potential explanation for the misclassifications could be that parties change their policy positions over time.
+%If a party adapts positions from another party during the legislative period this can lead to misclassification of the previously written respective manifesto text. 
 The confusion matrix for the 17th Bundestag in \autoref{tab:confusion_topic} shows that the SPD manifesto texts are often predicted as belonging to the CDU/CSU on the topic level.
-A similar effect can be observed in the 18th Bundestag when predicting both manifesto texts as well as facebook posts (results not shown): texts of the Green party are often predicted to be from the CDU/CSU.
-While we cannot list examples due to space limitations, we emphasize that these misclassifications can be related to policy changes of parties.
+This was the the legislative period when the CDU under chancellor Merkel was making a strong left move with respect to socioeconomic issues.
 \\
@@
 we might leave this out then?
+% Ich habe bei der section immer noch etwas Bauchschmerzen, habe jetzt nochmal umformuliert und versucht doch ein example unterzubringen.
 
@@
 \label{tab:results_binary_17}
-Classification accuracy on the binary prediction problem, categorizing texts into government and opposition.
+Classification accuracy on the binary prediction problem, categorising texts into government and opposition.
 Out-of-domain accuracy again drops close to chance performance for the manifesto data but remains higher for the facebook post texts.
@@
 We also trained a model on government membership labels, in order to a better compare against other studies that predict party affiliation in a two party system.
-We show the results for the 17th legislative period in \autoref{tab:results_binary_17}.
-While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on manifesto data drops again close to chance performance.
+\autoref{tab:results_binary_17} shows the results for the 17th legislative period.
+While the in-domain prediction accuracy is close to 0.9, the out-of-domain evaluation on manifesto data drops again to a performance close to chance.
 This is in line with results on binary classification of political bias in the Canadian parliament~\cite{Yu2008}.
@@
 In our results, the aggregation into topics did not increase the accuracy in this binary setting when classifying manifesto texts.
-The drop in accuracy of the binary classifier on facebook data (aggregated analogous to the party affiliation case) was less pronounced, accuracies were above 0.70.
+The drop in accuracy of the binary classifier on facebook data (aggregated analogous to the party affiliation case) was less pronounced: accuracies were above 0.70.
 
@@
 \label{sec:discrim_effect}
-An important question is whether the difference between the features of each party stems from different policies or from other aspects of the text.
-To address this point we analysed features that are discriminative for government membership and for parties, respectively.
+Another important question when analysing automatic text classification models is whether the difference between the features of each party stems from different policies or from other aspects of the text.
+To address this point we analysed features that are discriminative for government membership and for parties.
 
@@
 What do facebook posts and plenary debates have in common?
-In contrast to the authors of manifestos both the speakers in the parliament as well as the authors of facebook posts know which party is in the government.
-A language feature that could capture this is sentiment.
+In contrast with the authors of manifestos, both the speakers in the parliament as well as the authors of facebook posts know which party is in government.
+A language feature that might capture this is sentiment.
 Indeed our results in \autoref{tab:sentiments} show that positive sentiment strongly correlates with government membership and the number of seats in the parliament.
-Previous studies also find that text features discriminative in a two party system are not necessarily related to policies but more to language of defense and attack~\cite{Hirst2014}.
+Previous studies also find that text features which are discriminative in a two party system are not necessarily related to policies but more to language of defence and attack~\cite{Hirst2014}.
 
@@
 \label{tab:sentiments}
-Correlation coefficient between the average sentiment of political speeches of a party in the german Bundestag with two indicators of political power: a) membership in the government and b) the number of seats a party occupies in the parliament.
+Correlation coefficient between the average sentiment of political speeches of a party in the German Bundestag with two indicators of political power: a) membership in the government and b) the number of seats a party occupies in the parliament.
 }
@@
 In order to determine further discriminative features, we quantified which words were preferentially used by each party by measuring the correlation of single words with the party label.
-Unspecific stopwords are excluded.
+Unspecific stopwords were excluded.
 We find clear differences between the parties, which are in line with the parties ideologies.
@@
 \section{Conclusions and Limitations}\label{sec:conclusion}
-During our evaluation of classifiers trained on parliament speeches, we find that automated political bias prediction is possible with above chance accuracy even beyond the training text domain.
-These results suggest that such systems could be helpful as assistive technology for instance for human annotators in an active learning setting.
+%During our evaluation of classifiers trained on parliament speeches, 
+We find that automated political bias prediction is possible with an accuracy better than chance, even beyond the training text domain.
+These results suggest that such systems could be helpful as assistive technology, for example for human annotators in an active learning setting.
 
-In line with previous findings \cite{Yu2008, Hirst2014}, we encounter a strong effect of text length and text domain on the generalization performance of the classifier.
+In line with previous findings \cite{Yu2008, Hirst2014}, we find a large effect of text length and text domain on the generalisation performance of the classifier.
 The first effect, that longer texts are easier to classify, intuitively makes sense.
@@
 % ich hab versucht, das klarer zu machen, indem ich das in dem satz oben auch erwaehne. 
-However, short texts are a realistic challenge for automated political bias prediction systems: Political texts from social media data and other web sources are indeed often very short and hence difficult to analyse for both human annotators and algorithms.
-Both political education and science can benefit from automatic analyses of these very data streams, as these media have strong influence on public opinions yet cannot be analysed by humans alone, due to the volume of data.
+However, short texts are a realistic challenge for automated political bias prediction systems: political texts from social media data and other web sources are often very short and hence difficult to analyse for both human annotators and algorithms.
+Both political education and science can benefit from automatic analyses of these very data streams, as these fields have a strong influence on public opinion and yet cannot be analysed by humans alone, due to the volume of data.
 
 The second effect, i.e.
-the drop in generalization performance on out-of-domain data, appears to be correlated to the first one: it can be alleviated in some cases by aggregating texts into longer segments.
+the drop in generalisation performance on out-of-domain data, appears to be correlated to the first one: it can be alleviated in some cases by aggregating texts into longer segments.
 In the case of party affiliation prediction, the out-of-domain classification is on a par or even better than the prediction accuracy on in-domain data.
-However in the binary classification setting (government membership prediction), text aggregation does not help as much: Aggregating manifesto data, which was written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain accuracy drop.
+However in the binary classification setting (government membership prediction), text aggregation does not help as much: aggregating manifesto data, written without the knowledge of which party would be member of the government, into longer texts does not counteract the effect of out-of-domain accuracy drop.
 We attribute this effect in part to the fact that sentiment appears to be a discriminative feature for government membership.

### 2016-06-25 bundestag.py

@@ -298,10 +298,10 @@ def news_experiment(trainData,trainLabels,evalData,evalLabels, folds=2, idstr='d
                             ('tfidf', TfidfTransformer()),
                             ('clf',LogisticRegression(class_weight='balanced'))])
     parameters = {'vect__ngram_range': [(1,2)],\
-           'vect__stop_words':(None,stops),\
-           'tfidf__use_idf': (True,False),\
-           'clf__C': (10.**sp.arange(1,5,1.)).tolist(),
-           'vect__max_df':[.3, .5, .75],
+           #'vect__stop_words':(None,stops),\
+           #'tfidf__use_idf': (True,False),\
+           'clf__C': (10.**sp.arange(-2,5,1.)).tolist(),
+           'vect__max_df':[.1, .5, .75],
         }
     saveId = idstr+"-"+randid()
     X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
@@ -345,8 +345,8 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     parameters = {'vect__ngram_range': [(1,2)],\
            #'vect__stop_words':(None,stops),\
            #'tfidf__use_idf': (True,False),\
-           'clf__C': (10.**sp.arange(-2,3,1.)).tolist(),
-           'vect__max_df':[.5],
+           'clf__C': (10.**sp.arange(-2,5,.5)).tolist(),
+           'vect__max_df':[.1,.2,.5,7],
         }
     saveId = idstr+"-"+randid()
     X_train, X_test, y_train, y_test = cross_validation.train_test_split(trainData, trainLabels, test_size=0.1, random_state=0)
@@ -374,6 +374,7 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(evalLabels, predictedEval).tolist(): report += str(line)+"\n" 
     report += "Accuracy: %0.2f\n"%metrics.accuracy_score(evalLabels,predictedEval)
+    import pdb;pdb.set_trace()
     # dump report
     open(OUT_DIR+'/report-'+saveId+'.txt','wb').write(report)
     return report
@@ -404,13 +405,18 @@ def classify_speeches_binary_parliament(legislationPeriod = 18):
[…]

### 2016-06-25 fb.py

@@ -9,11 +9,25 @@ from sklearn import cross_validation,metrics
 import scipy as sp
 import re
 
-DDIR = '/Users/felix/Data/fb-politics/'
+DDIR = "/Users/felix/Code/Python/fipi/data/parteien-auf-fb"
 
 dat = [DDIR + x + ".json.gz" for x in ['afd','npd','pegida']]
 
 urlPat = r'(http://.*\.html)'
+
+def getUrlsAndUsers(post): 
+    urls = []
+    if post['message']: 
+        foundUrl = re.search(urlPat,post['message'])
+        if foundUrl:
+            name = ''
+            if post.has_key('name'):
+                name += post['name']
+            import pdb;pdb.set_trace()
+            urls.append((post['created'],post['id']+": "+name,foundUrl.groups(0)[0]))
+    if len(post['comments'])>0: 
+        urls += list(chain(*map(getUrls,post['comments'])))
+    return filter(lambda x: len(x)>0,urls)
 
 def getUrls(post): 
     urls = []
@@ -21,7 +35,6 @@ def getUrls(post):
         foundUrl = re.search(urlPat,post['message'])
         if foundUrl:
             name = ''
-            import pdb;pdb.set_trace()
             if post.has_key('name'):
                 name += post['name']
             urls.append((post['created'],post['id']+": "+name,foundUrl.groups(0)[0]))
@@ -121,5 +134,16 @@ class fbTraverserUrl(object):
             for line in fh:
                 yield getUrls(json.loads(line, strict=False))
                     
+
+class fbTraverserUrlUsers(object):
+    def __init__(self, fn):
+        self.fn = fn
+
+    def __iter__(self):
+        with gzip.open(self.fn) as fh:
+            for line in fh:
+                yield getUrlsAndUsers(json.loads(line, strict=False))
+                    
+
 
 

### 2016-06-26 manuscript/poltext.tex
fb parser with more functionality - added news download and text cleaning - flattened likes per post

@@
 Automatic political text analysis requires a profound understanding of the models used.
-One way to better understand these models is to inspect the misclassifications of a model: a potential explanation for the misclassifications could be that parties change their policy positions over time.
+One way to better understand these models is to inspect the misclassifications of a model.
+A potential explanation for the misclassifications could be that parties change their policy positions over time.
 %If a party adapts positions from another party during the legislative period this can lead to misclassification of the previously written respective manifesto text. 

### 2016-06-26 fb.py

@@ -8,6 +8,8 @@ from sklearn.grid_search import GridSearchCV
 from sklearn import cross_validation,metrics
 import scipy as sp
 import re
+import urllib
+from bs4 import BeautifulSoup
 
 DDIR = "/Users/felix/Code/Python/fipi/data/parteien-auf-fb"
 
@@ -15,7 +17,57 @@ dat = [DDIR + x + ".json.gz" for x in ['afd','npd','pegida']]
 
 urlPat = r'(http://.*\.html)'
 
-def getUrlsAndUsers(post): 
+def get_text_from_url(url):
+    html = urllib.urlopen(url).read()
+    soup = BeautifulSoup(html,"lxml")
+    # kill all script and style elements
+    for script in soup(["script", "style"]):
+        script.extract()    # rip it out
+    # get text
+    text = soup.get_text()
+    # break into lines and remove leading and trailing space on each
+    lines = (line.strip() for line in text.splitlines())
+    # break multi-headlines into a line each
+    chunks = (phrase.strip() for line in lines for phrase in line.split("  "))
+    # drop blank lines
+    text = '\n'.join(chunk for chunk in chunks if len(chunk.split(" "))>1)
+    return text.encode('utf-8')
+
+def getLikes(post):
+    likes = []
+    if post.has_key("likedBy"):
+        likes = [post['likedBy']]   
+    if len(post['comments'])>0: 
+        likes += map(getLikes,post['comments'])
+    return chain(*likes)
+    
+
+def getContent(post):
+    text,likes,urlText,name,url,content = "",[],"","","",{}
+    if post['message']: 
+        text = post['message']
+        foundUrl = re.search(urlPat,post['message'])
+        if foundUrl:
+            url = foundUrl.groups(0)[0]
+            urlText = get_text_from_url(url)
+        if post.has_key('name'): 
+            name = post['name']
+        if post.has_key("likedBy"):
+            likes = post['likedBy']    
+        content = {
+            "url":url,
+            "urlText":urlText,
+            "text":text,
+            "likes":likes,
+            "date":post['created'],
+            "id":post['id'] + ": "+name 
+            }
+    if len(post['comments'])>0: 
+        content['likes'] = list(getLikes(post)) + likes
+    return content
+
+
+def getContentFlatOld(post): 
     urls = []
     if post['message']: 
         foundUrl = re.search(urlPat,post['message'])
@@ -23,11 +75,30 @@ def getUrlsAndUsers(post):
             name = ''
             if post.has_key('name'):
[…]

### 2016-07-06 cointeraction.py
Added Kang et al's approach for scaling up random walk graph kernels.

@@ -1,7 +1,7 @@
 import json,gzip
 import pandas as pd
 from scipy.sparse import csr_matrix,diags
-from scipy.sparse.linalg import eigs
+from scipy.sparse.linalg import eigs,svds
 from itertools import chain
 from sklearn.feature_extraction.text import CountVectorizer
 from sklearn.pipeline import Pipeline
@@ -23,7 +23,7 @@ dat = [DDIR + x + ".json.gz" for x in ['afd','npd','pegida']]
 
 urlPat = r'(http://.*\.html)'
 
-def fit_kcca(Ks,ncomp=1,gamma=1e-3):
+def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     Fits kernel CCA model
     INPUT:
@@ -52,6 +52,27 @@ def fit_kcca(Ks,ncomp=1,gamma=1e-3):
         alphas.append(Vs[ik*N:(ik+1)*N,:ncomp])
     return alphas
 
+def normAdjMat(A):
+    '''
+    Normalizes columns of adjecency matrix
+    '''
+    norm = adj_mat.sum(axis=0)
+    norm[norm == 0] = 1
+    return adj_mat / norm
+
+def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
+    '''
+    Computes approximate random walk graph kernel 
+    as in algorithm 2 of Kang et al, "Fast Random Walk Graph Kernel", SIAM
+    Instead of adjeciency matrix A the algorithm starts with an eigendecomposition
+    of A for each graph
+    '''
+    # compute inverse in eigenspace of product graph
+    Lambda = 1./(1./sp.hstack([LB * la for la in LA]) - c)
+    L = sp.kron(qA.dot(UA),qB.dot(UB))
+    R = sp.kron(UA.dot(pA),UB.dot(pB))
+    return qA.T.dot(pA) * qB.T.dot(pB) + c*L.dot(sp.diag(Lambda)).dot(R)
+
 def readPostLine(line):
     c = line.decode('utf-8').split("\t")
     postId, postType, usrLikes = c[0], c[1], [int(i) for i in c[2:]]
@@ -68,14 +89,14 @@ def readPostWeek(fn,maxUsers,numComp=6):
     rows,cols = zip(*chain(*map(enumerate,likes)))
[…]

### 2016-07-07 cointeraction.py
Added missing matrix transpose in random walk kernel approximation.

@@ -70,7 +70,7 @@ def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
     # compute inverse in eigenspace of product graph
     Lambda = 1./(1./sp.hstack([LB * la for la in LA]) - c)
     L = sp.kron(qA.dot(UA),qB.dot(UB))
-    R = sp.kron(UA.dot(pA),UB.dot(pB))
+    R = sp.kron(UA.T.dot(pA),UB.T.dot(pB))
     return qA.T.dot(pA) * qB.T.dot(pB) + c*L.dot(sp.diag(Lambda)).dot(R)
 
 def readPostLine(line):
@@ -112,15 +112,16 @@ def getPartyKernel(party,fns,maxUser,numComp, years=['2014','2015','2016']):
     p = Pool(4)
     cigs = p.map(getCointeractionGraphTuple,tpls)
     N = len(cigs)
-    K = sp.zeros((N,N))
     print("Found %d weeks"%N)
+    K = sp.zeros((N,N))
     p = sp.ones(maxUser)/maxUser
     for x in range(N):
         for y in range(x+1,N):
             K[x,y] = randomWalkGraphKernelApprox(cigs[x][1],cigs[x][0],cigs[y][1],cigs[y][0],p,p,p,p,c=0.5) 
     #X = sp.sparse.vstack([sp.sparse.hstack([*c[1]]) for c in cigs])
     #K = X.dot(X.T)
-    return sp.array(sp.real(K.todense()))
+    #return sp.array(sp.real(K.todense()))
+    return K
 
 def getPartyKernelTupel(tpl):return getPartyKernel(*tpl)
 

### 2016-07-07 cointeraction.py
Random walk kernel approx now returns full kernel matrix

@@ -114,14 +114,14 @@ def getPartyKernel(party,fns,maxUser,numComp, years=['2014','2015','2016']):
     N = len(cigs)
     print("Found %d weeks"%N)
     K = sp.zeros((N,N))
-    p = sp.ones(maxUser)/maxUser
+    prob = sp.ones(maxUser)/maxUser
     for x in range(N):
-        for y in range(x+1,N):
-            K[x,y] = randomWalkGraphKernelApprox(cigs[x][1],cigs[x][0],cigs[y][1],cigs[y][0],p,p,p,p,c=0.5) 
+        for y in range(x,N):
+            K[x,y] = randomWalkGraphKernelApprox(cigs[x][1],cigs[x][0],cigs[y][1],cigs[y][0],prob,prob,prob,prob,c=0.5) 
     #X = sp.sparse.vstack([sp.sparse.hstack([*c[1]]) for c in cigs])
     #K = X.dot(X.T)
     #return sp.array(sp.real(K.todense()))
-    return K
+    return K + K.T
 
 def getPartyKernelTupel(tpl):return getPartyKernel(*tpl)
 

### 2016-07-08 cointeraction.py
initial test for random walk kernel approximation.

@@ -56,9 +56,29 @@ def normAdjMat(A):
     '''
     Normalizes columns of adjecency matrix
     '''
-    norm = adj_mat.sum(axis=0)
-    norm[norm == 0] = 1
-    return adj_mat / norm
+    norm = A.sum(axis=0)
+    norm[norm == 0] = 1.0
+    return A / sp.double(norm)
+
+def testRandomWalk():
+    A = sp.array([[0,1,1],[1,0,1],[1,1,0]])
+    B = sp.array([[0,1,0],[1,0,1],[0,1,0]])
+    q = sp.ones(A.shape[0])/A.shape[0]
+    k = randomWalkGraphKernel(A,B,q,q,q,q)    
+    LA,UA = eig(A)
+    LB,UB = eig(B)
+    q = csr_matrix(q)
+    r = 2
+    khat = randomWalkGraphKernelApprox(csr_matrix(UA[:,:r]),csr_matrix(LA[:r]),csr_matrix(UB[:,:r]),csr_matrix(LB[:r]),q,q,q,q)
+
+def randomWalkGraphKernel(UA,UB,qA,qB,pA,pB,c=0.5):
+    '''
+    Computes random walk graph kernel 
+    '''
+    W = sp.kron(normAdjMat(UA),normAdjMat(UB))
+    q = sp.kron(qA,qB)
+    p = sp.kron(pB,pB)
+    return q.dot(sp.linalg.inv(sp.eye(W.shape[0]) - c*W)).dot(p)
 
 def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
     '''
@@ -68,10 +88,12 @@ def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
     of A for each graph
     '''
     # compute inverse in eigenspace of product graph
-    Lambda = 1./(1./sp.hstack([LB * la for la in LA]) - c)
-    L = sp.kron(qA.dot(UA),qB.dot(UB))
-    R = sp.kron(UA.T.dot(pA),UB.T.dot(pB))
-    return qA.T.dot(pA) * qB.T.dot(pB) + c*L.dot(sp.diag(Lambda)).dot(R)
+    o = csr_matrix(sp.ones(UA.shape[1]*UB.shape[1]))
+    Lambda = csr_matrix(o/((o/sp.sparse.kron(LA,LB)) - c))
+    L = sp.sparse.kron(UA.T.dot(qA),UB.T.dot(qB))
[…]

### 2016-07-08 cointeraction.py
fixed test?

@@ -61,15 +61,17 @@ def normAdjMat(A):
     return A / sp.double(norm)
 
 def testRandomWalk():
-    A = sp.array([[0,1,1],[1,0,1],[1,1,0]])
-    B = sp.array([[0,1,0],[1,0,1],[0,1,0]])
+    A = sp.array([[0,1,1,0],[1,0,1,1],[1,1,0,1],[0,1,1,0]])
+    B = sp.array([[0,1,0,0],[1,0,1,1],[0,1,0,1],[0,1,1,0]])
     q = sp.ones(A.shape[0])/A.shape[0]
     k = randomWalkGraphKernel(A,B,q,q,q,q)    
-    LA,UA = eig(A)
-    LB,UB = eig(B)
-    q = csr_matrix(q)
+    LA,UA = sp.linalg.eig(A)
+    LB,UB = sp.linalg.eig(B)
+    q = csr_matrix(q).T
     r = 2
     khat = randomWalkGraphKernelApprox(csr_matrix(UA[:,:r]),csr_matrix(LA[:r]),csr_matrix(UB[:,:r]),csr_matrix(LB[:r]),q,q,q,q)
+    print("k: %f"%k)
+    print("khat: %f"%khat)
 
 def randomWalkGraphKernel(UA,UB,qA,qB,pA,pB,c=0.5):
     '''

### 2016-07-10 cointeraction.py
More graph kernel experiments.

@@ -33,7 +33,7 @@ def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     N = Ks[0].shape[0]
     m = len(Ks)
-    #Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
+    Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
     # Generate Left-hand side of eigenvalue equation
     VK = sp.vstack(Ks)
     LH = VK.dot(VK.T)
@@ -61,13 +61,13 @@ def normAdjMat(A):
     return A / sp.double(norm)
 
 def testRandomWalk():
-    A = sp.array([[0,1,1,0],[1,0,1,1],[1,1,0,1],[0,1,1,0]])
-    B = sp.array([[0,1,0,0],[1,0,1,1],[0,1,0,1],[0,1,1,0]])
-    q = sp.ones(A.shape[0])/A.shape[0]
-    k = randomWalkGraphKernel(A,B,q,q,q,q)    
+    A = normAdjMat(sp.array([[0,1,1,0],[1,0,1,1],[1,1,0,1],[0,1,1,0]]))
+    B = normAdjMat(sp.array([[0,1,0,0],[1,0,1,1],[0,1,0,1],[0,1,1,0]]))
+    qd = sp.ones(A.shape[0])/A.shape[0]
+    k = randomWalkGraphKernel(A,B,qd,qd,qd,qd)    
     LA,UA = sp.linalg.eig(A)
     LB,UB = sp.linalg.eig(B)
-    q = csr_matrix(q).T
+    q = csr_matrix(qd).T
     r = 2
     khat = randomWalkGraphKernelApprox(csr_matrix(UA[:,:r]),csr_matrix(LA[:r]),csr_matrix(UB[:,:r]),csr_matrix(LB[:r]),q,q,q,q)
     print("k: %f"%k)
@@ -91,11 +91,15 @@ def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
     '''
     # compute inverse in eigenspace of product graph
     o = csr_matrix(sp.ones(UA.shape[1]*UB.shape[1]))
-    Lambda = csr_matrix(o/((o/sp.sparse.kron(LA,LB)) - c))
+    Lambda = csr_matrix(o/((o/sp.kron(LA.flatten(),LB.flatten())) - c))
[…]

### 2016-07-13 api.py
Added temporal embedding to linear co-interaction kernels.

@@ -6,7 +6,6 @@ from classifier import Classifier
 from retrying import retry
 import urllib2
 from bs4 import BeautifulSoup
-from readability.readability import Document
 from newsreader import get_news
 from apscheduler.schedulers.background import BackgroundScheduler
 import json
@@ -26,12 +25,20 @@ def fetch_url(url):
     '''
     get url with readability
     '''
-    html = urllib2.urlopen(url).read()
-    readable_article = Document(html).summary()
-    title = Document(html).short_title()
-    text = BeautifulSoup(readable_article).get_text()
-
-    return title,text
+    html = urllib.request.urlopen(url).read()
+    soup = BeautifulSoup(html,"lxml")
+    # kill all script and style elements
+    for script in soup(["script", "style"]):
+        script.extract()    # rip it out
+    # get text
+    text = soup.get_text()
+    # break into lines and remove leading and trailing space on each
+    lines = (line.strip() for line in text.splitlines())
+    # break multi-headlines into a line each
+    chunks = (phrase.strip() for line in lines for phrase in line.split("  "))
+    # drop blank lines
+    text = '\n'.join(chunk for chunk in chunks if len(chunk.split(" "))>1)
+    return text.encode('utf-8')
 
 ### API
 @app.route("/api/newstopics")

### 2016-07-13 cointeraction.py

@@ -23,6 +23,17 @@ dat = [DDIR + x + ".json.gz" for x in ['afd','npd','pegida']]
 
 urlPat = r'(http://.*\.html)'
 
+def embed(X,tau):
+    '''
+    Temporal embedding for scipy.sparse matrices
+    '''
+    T,D = X.shape
+    startInd = max(0,-tau.min())
+    stopInd = min(T,T-tau.max())
+    Xt = sp.sparse.hstack([X[startInd+t:stopInd+t,:] for t in tau])
+    return Xt
+    
+
 def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     Fits kernel CCA model
@@ -33,7 +44,7 @@ def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     N = Ks[0].shape[0]
     m = len(Ks)
-    Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
+    #Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
     # Generate Left-hand side of eigenvalue equation
     VK = sp.vstack(Ks)
     LH = VK.dot(VK.T)
@@ -99,7 +110,7 @@ def randomWalkGraphKernelApprox(UA,LA,UB,LB,qA,qB,pA,pB,c=0.5):
 
 def randomWalkGraphKernelApproxTuple(tpl):
     p = csr_matrix(sp.ones(tpl[0].shape[0])/tpl[0].shape[0]).T
-    return (tpl[-2],tpl[-1],randomWalkGraphKernelApprox(tpl[0],tpl[1],tpl[2],tpl[3],p,p,p,p,0.99))
+    return (tpl[-2],tpl[-1],randomWalkGraphKernelApprox(tpl[0],tpl[1],tpl[2],tpl[3],p,p,p,p,0.00001))
 
 def readPostLine(line):
     c = line.decode('utf-8').split("\t")
@@ -130,7 +141,7 @@ def getCointeractionGraph(fn,maxUsers,numComp,k=3):
         #N = A.T.dot(csr_matrix(A.sum(axis=1)))
         #A.dot(diags(sp.array(( N /).todense()).flatten()))
         U,S,V = svds(A,numComp)
-        return (S**2,csr_matrix(V.T))
+        return (S**2,csr_matrix(V).T)
     except:
         return (sp.ones(numComp),A[:numComp,:].T)
 
@@ -141,7 +152,7 @@ def graphKernelDummy(A,B):
 
[…]

### 2016-07-13 bundestag.py
Added topic classification experiment (c.f. Zirn et al) Made scripts compatible with python 3.

@@ -18,7 +18,7 @@ from itertools import chain
 from sklearn.cross_validation import StratifiedKFold
 import pandas as pd
 from scipy import random,unique,vstack,hstack,arange,array,mean,corrcoef
-import cPickle
+import pickle
 from sentiments import SentimentClassifier
 from fb import fbTraverserTop
 
@@ -134,21 +134,21 @@ def get_sentiments(legis=17):
     # correlation between sentiment and government membership
     member = hstack([vstack([sentiments[p],[1.0] * len(sentiments[p])]) if p in gov else vstack([sentiments[p],[-1.0] * len(sentiments[p])]) for p in sentiments.keys()])
     corMember = corrcoef(member[0,:],member[1,:])
-    print "Corr sentiment vs government membership: %0.2f"%corMember[0,1]
+    print("Corr sentiment vs government membership: %0.2f"%corMember[0,1])
     meanSentiment,govMember =zip(*[(mean(sentiments[p]),1.0) if p in gov else (mean(sentiments[p]), -1.0) for p in sentiments.keys()])
-    print "Corr Mean Sentiment vs gov membership: %0.2f"%corrcoef(meanSentiment,govMember)[0,1]
+    print("Corr Mean Sentiment vs gov membership: %0.2f"%corrcoef(meanSentiment,govMember)[0,1])
     # correlation between sentiment and party seats
     seats = hstack([vstack([sentiments[p],[bundestagSeats[legis][p]] * len(sentiments[p])]) for p in sentiments.keys()])
     corSeats = corrcoef(seats[0,:],seats[1,:])
-    print "Corr sentiment vs seats: %0.2f"%corSeats[0,1]
+    print("Corr sentiment vs seats: %0.2f"%corSeats[0,1])
     # mean of parties
     meanSentiment,seats=zip(*[(mean(sentiments[p]),bundestagSeats[legis][p]) for p in sentiments.keys()])
-    print "Corr Mean Sentiment vs seats: %0.2f"%corrcoef(meanSentiment,seats)[0,1]
+    print("Corr Mean Sentiment vs seats: %0.2f"%corrcoef(meanSentiment,seats)[0,1])
 
 def get_stops(legislationPeriod=17,includenames=True):
     # generic stopwords
     stopwords = codecs.open("data/stopwords.txt", "r", "utf-8").readlines()[7:]
-    stops = map(lambda x:x.lower().strip(),stopwords)
+    stops = [x.lower().strip() for x in stopwords]
     names = []
     if includenames:
         # names of abgeordnete
[…]

### 2016-07-13 classifier.py

@@ -1,5 +1,5 @@
 # -*- coding: utf-8 -*-
-import cPickle
+import pickle
 from scipy import ones,hstack,arange,reshape,zeros,setdiff1d
 import json
 import os
@@ -133,10 +133,10 @@ class Classifier:
         '''
         # if there is no classifier file or training is invoked
         if (not os.path.isfile('classifier.pickle')) or train:
-            print 'Training classifier'
+            print('Training classifier')
             self.train(text_clf=text_clf,parameters=parameters)
-        print 'Loading classifier'
-        self.clf = cPickle.load(open('classifier.pickle'))['clf']
+        print('Loading classifier')
+        self.clf = pickle.load(open('classifier.pickle'))['clf']
 
     def predict(self,text):
         '''
@@ -191,5 +191,5 @@ class Classifier:
         gs_clf = GridSearchCV(text_clf, parameters, cv=folds, n_jobs=-1,verbose=4)
         gs_clf.fit(data,labels)
         # dump classifier to pickle
-        cPickle.dump({'clf':gs_clf.best_estimator_,'score':gs_clf.best_score_},open('classifier.pickle','wb'),-1)
+        pickle.dump({'clf':gs_clf.best_estimator_,'score':gs_clf.best_score_},open('classifier.pickle','wb'),-1)
 

### 2016-07-13 cointeraction.py

@@ -11,6 +11,7 @@ from sklearn.externals.joblib import Parallel,delayed
 import glob
 from multiprocessing import Pool
 from sklearn import cross_validation,metrics
+import pylab as pl
 import scipy as sp
 import os  
 import re
@@ -178,7 +179,7 @@ def getPartyKernel(party,fns,maxUser,numComp, years=['2014','2015','2016'], kern
 
 def getPartyKernelTupel(tpl):return getPartyKernel(*tpl)
 
-def readAll(folder=DDIR,numComp=6,years=['2014','2015','2016']):
+def readAll(folder=DDIR,numComp=3,years=['2014','2015','2016']):
     fs = [(d,os.listdir(DDIR+"/"+d)) for d in os.listdir(DDIR) if os.path.isdir(DDIR+"/"+d)]
     print("Found %d parties in %s"%(len(fs),folder))
     maxUser = 1+max([max([readMaxUser(os.path.join(DDIR,fss[0],ff)) for ff in fss[1]]) for fss in fs])
@@ -193,7 +194,7 @@ def evaluateKCCA(Ks,trainIdx,testIdx,numComp,gamma):
     yhatTest = [a[0].T.dot(a[1]) for a in zip(alphas,[k[trainIdx,:][:,testIdx] for k in Ks.values()])]
     return yhatTrain,yhatTest
 
-def run_cointeraction(folder=DDIR,numComp=6,years=['2014','2015','2016'],testRatio=.5):
+def run_cointeraction(folder=DDIR,numComp=2,years=['2014','2015','2016'],testRatio=.5):
     Ks = readAll(folder,numComp,years)
     N = Ks[list(Ks.keys())[0]].shape[0]    
     trainIdx = range(int(N * (1-testRatio)))
@@ -224,7 +225,6 @@ def getCorrs(yhat,M,ic):
     return cors
 
 def plotTrends(Ks,yhat,numComp,teststr):   
-    import pylab as pl
     for ic in range(numComp):
         cors = sp.zeros((len(Ks),len(Ks)))
         for x in range(len(Ks)):

### 2016-07-13 party_classifier.py

@@ -1,5 +1,5 @@
 # -*- coding: utf-8 -*-
-import cPickle
+import pickle
 from scipy import ones,hstack,arange,reshape,zeros,random
 import json
 import os
@@ -96,10 +96,10 @@ class PartyClassifier:
         '''
         # if there is no classifier file or training is invoked
         if (not os.path.isfile('party_classifier.pickle')) or train:
-            print 'Training party classifier'
+            print('Training party classifier')
             self.train(text, data, text_clf=text_clf, parameters=parameters)
-        print 'Loading party classifier'
-        self.clf = cPickle.load(open('party_classifier.pickle'))['clf']
+        print('Loading party classifier')
+        self.clf = pickle.load(open('party_classifier.pickle'))['clf']
 
     def predict(self,text):
         '''
@@ -146,6 +146,6 @@ class PartyClassifier:
         gs_clf = GridSearchCV(text_clf, parameters, cv=StratifiedKFold(labels, folds), n_jobs=-1,verbose=4)
         gs_clf.fit(data,labels)
         # dump classifier to pickle
-        cPickle.dump({'clf':gs_clf.best_estimator_,'score':gs_clf.best_score_},\
+        pickle.dump({'clf':gs_clf.best_estimator_,'score':gs_clf.best_score_},\
             open('party_classifier.pickle','wb'),-1)
 

### 2016-07-13 bundestag.py
Topic data reading without undefined topics.

@@ -300,7 +300,7 @@ def csv2DataTupleTopicsAllParties(f):
     df = pd.read_csv(f)
     df['content'] = df['content'].astype('str')
     df['topic'] = ((df['cmp_code'].dropna()) / 100).apply(lambda x: int(x))
-    df = df.dropna()
+    df = df.ix[df['topic'] > 0].dropna()
     return zip(df['content'].tolist(), df['topic'].tolist())
 
 def csv2DataTupleTopics(f):
@@ -397,7 +397,6 @@ def optimize_hyperparams(trainData,trainLabels,evalData,evalLabels, folds=2, ids
     report += '\nConfusion Matrix (rows=true, cols=predicted)\n'+', '.join(labelsStr)+'\n'
     for line in metrics.confusion_matrix(evalLabels, predictedEval).tolist(): report += str(line)+"\n" 
     report += "Accuracy: %0.2f\n"%metrics.accuracy_score(evalLabels,predictedEval)
-    import pdb;pdb.set_trace()
     # dump report
     open(OUT_DIR+'/report-'+saveId+'.txt','wb').write(report.encode('utf-8'))
     return report

### 2016-07-14 cointeraction.py
A simpler graph representation.

@@ -33,7 +33,16 @@ def embed(X,tau):
     stopInd = min(T,T-tau.max())
     Xt = sp.sparse.hstack([X[startInd+t:stopInd+t,:] for t in tau])
     return Xt
-    
+
+def centerKernel(K):
+    # K is of dimension N x N
+    N = K.shape[0]
+    # D is a row vector storing the column averages of K
+    D = K.sum(axis=0)/N
+    # E is the average of all the entries of K
+    E = K.sum() / N
+    J = sp.outer(sp.ones(N),D)
+    return K - J - J.T + E * sp.ones((N, N))
 
 def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
@@ -45,6 +54,7 @@ def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     N = Ks[0].shape[0]
     m = len(Ks)
+    Ks = [centerKernel(k) for k in Ks]
     #Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
     # Generate Left-hand side of eigenvalue equation
     VK = sp.vstack(Ks)
@@ -141,8 +151,11 @@ def getCointeractionGraph(fn,maxUsers,numComp,k=3):
         # normalizing diagonal matrix
         #N = A.T.dot(csr_matrix(A.sum(axis=1)))
         #A.dot(diags(sp.array(( N /).todense()).flatten()))
-        U,S,V = svds(A,numComp)
-        return (S**2,csr_matrix(V).T)
+        ####### 
+        #U,S,V = svds(A,numComp)
+        #return (S**2,csr_matrix(V).T)
+        a = csr_matrix(A.sum(axis=0))
+        return csr_matrix(a/a)
     except:
         return (sp.ones(numComp),A[:numComp,:].T)
 
@@ -170,7 +183,7 @@ def getPartyKernel(party,fns,maxUser,numComp, years=['2014','2015','2016'], kern
             K[k[0],k[1]] = k[2] 
             K[k[1],k[0]] = k[2]
     elif kernelType == 'linear':
-        tau = sp.array([-2,-1,0])
+        tau = sp.array([-1,0])
         X = csr_matrix(sp.sparse.vstack([sp.sparse.hstack([*c[1].T]) for c in cigs]))
         X = embed(X,tau)
         K = sp.array(sp.real(X.dot(X.T).todense()))
[…]

### 2016-07-14 cointeraction.py
Finalized poltext presentation Simpler graph analysis.

@@ -2,6 +2,7 @@ import json,gzip
 import pandas as pd
 from scipy.sparse import csr_matrix,diags
 from scipy.sparse.linalg import eigs,svds
+from scipy.stats import zscore
 from itertools import chain
 from sklearn.feature_extraction.text import CountVectorizer
 from sklearn.pipeline import Pipeline
@@ -54,7 +55,7 @@ def fitKcca(Ks,ncomp=1,gamma=1e-3):
     """
     N = Ks[0].shape[0]
     m = len(Ks)
-    Ks = [centerKernel(k) for k in Ks]
+    #Ks = [centerKernel(k) for k in Ks]
     #Ks = [k/sp.linalg.eigh(k)[0].max() for k in Ks]
     # Generate Left-hand side of eigenvalue equation
     VK = sp.vstack(Ks)
@@ -141,24 +142,10 @@ def readPostWeek(fn,maxUsers,numComp=6):
 
 def getCointeractionGraph(fn,maxUsers,numComp,k=3):
     A = readPostWeek(fn,maxUsers)
-    timestamp = fn.split("/")[-1].split(".")[0]
-    try:
-        #AAt = A.T.dot(A)
-        #AAtsum = sp.array(AAt.sum(axis=0)).flatten()
-        #AAtsum[AAtsum == 0] = 1.
-        #AAt = AAt.dot(diags(1./AAtsum))
-        #L,U = eigs(AAt,numComp)
-        # normalizing diagonal matrix
-        #N = A.T.dot(csr_matrix(A.sum(axis=1)))
-        #A.dot(diags(sp.array(( N /).todense()).flatten()))
-        ####### 
-        #U,S,V = svds(A,numComp)
-        #return (S**2,csr_matrix(V).T)
-        a = csr_matrix(A.sum(axis=0))
-        return csr_matrix(a/a)
-    except:
-        return (sp.ones(numComp),A[:numComp,:].T)
-
+    a = csr_matrix(A.sum(axis=0))
+    a.data = a.data/a.data
+    return a
+    
 def getCointeractionGraphTuple(x): return getCointeractionGraph(*x)
 
 def graphKernelDummy(A,B):
@@ -183,11 +170,11 @@ def getPartyKernel(party,fns,maxUser,numComp, years=['2014','2015','2016'], kern
             K[k[0],k[1]] = k[2] 
             K[k[1],k[0]] = k[2]
[…]
